### 2024-09-12 paper.qmd
.

---
title: "paper"
author: "Ali Youssef"
format: docx
editor: visual
---

##

# UPS1 Case Study 0

The dataset consists of DIA proteomics data from a UPS1-spiked E. coli protein mixture. Eight different concentrations of UPS1 were used (0.1 to 50 fmol), grouped into two categories: low concentrations (0.1–2.5 fmol, labeled as Conc. 2) and high concentrations (5–50 fmol, labeled as Conc. 1). In this dataset 48 HUMAN UPS1 proteins were detected. The data is provided in a `SummarizedExperiment` format. This example illustrates the application of LimROTS in a simple case where two groups are compared, considering only the concentrations and no other parameters. 48 UPS1-spiked human proteins were detected in this dataset.

## Run LimROTS

```{r}

# Load necessary packages
library(LimROTS)
library(parallel)

# Load the dataset
data("UPS1.Case0")
print(UPS1.Case0)

# Set random seed for reproducibility
set.seed(1234, kind = "default")

# Set metadata and formula for LimROTS analysis
meta.info <- "Conc."
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case0)[1]/4  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc."  # Formula for group comparison

# Run LimROTS analysis with trend and robust settings enabled
limrots.result <- LimROTS(data.exp = UPS1.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE)
```

## Run ROTS

```{r}

# Load necessary packages
library(LimROTS)
library(parallel)

# Load the dataset
data("UPS1.Case0")
print(UPS1.Case0)

# Set random seed for reproducibility
set.seed(1234, kind = "default")

# Set metadata and formula for LimROTS analysis
meta.info <- "Conc."
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case0)[1]/4  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc."  # Formula for group comparison

# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = UPS1.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE)
```

### 2024-09-12 R/LimROTS.r

@@ -63,7 +63,8 @@ LimROTS <- function (data.exp, B = 1000, K = NULL, a1 = NULL, a2 = NULL,
                      log = TRUE, progress = FALSE,
                                    verbose = TRUE, meta.info = NULL, cluster = NULL ,
                                   group.name = NULL , formula.str = NULL, trend = TRUE, robust = TRUE,
-                                  time = NULL, event = NULL, paired = FALSE)
+                                  time = NULL, event = NULL, paired = FALSE,
+                     n.ROTS = FALSE)
 {
 
 
@@ -217,7 +218,9 @@ LimROTS <- function (data.exp, B = 1000, K = NULL, a1 = NULL, a2 = NULL,
     message("Bootstrapping samples")
 
 
-
+  if(n.ROTS == TRUE){
+    meta.info = NULL
+  }
 
   if(!is.null(meta.info))
     {

### 2024-09-12 paper.qmd
.

@@
 group.name <- "Conc."
-formula.str <- "~ 0 + Conc."  # Formula for group comparison
+formula.str <- "Conc."
 
@@
                           cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
 ```
+
+## Run limma
+
+```{r}
+
+library(limma)
+
+data("UPS1.Case0")
+
+data <- assay(UPS1.Case0)
+
+formula.str <- "~ 0 + Conc."
+
+meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
+
+meta.info$sample.id <- NULL
+
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+
+fit <- lmFit(data, design.matrix)
+
+cont_matrix <- makeContrasts("Conc.1-Conc.2",  levels=design.matrix)
+
+fit2 <- contrasts.fit(fit, cont_matrix)
+
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+
+limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
+
+```

### 2024-09-13 paper.qmd
.

--- 2024-09-12 paper.qmd
@@
 ```{r}
-
 # Load necessary packages
@@
 library(parallel)
-
 # Load the dataset
@@
 print(UPS1.Case0)
-
 # Set random seed for reproducibility
 set.seed(1234, kind = "default")
-
 # Set metadata and formula for LimROTS analysis
@@
 K <- floor(K) 
-num_cores <- 10  # Number of cores for parallel processing
+num_cores <- 30  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
 formula.str <- "~ 0 + Conc."  # Formula for group comparison
-
 # Run LimROTS analysis with trend and robust settings enabled
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+saveRDS(limrots.result, "../LimROTS paper/UPS1.Case0.LimROTS.rds")
 ```
@@
 ## Run ROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+# Load the dataset
+data("UPS1.Case0")
+print(UPS1.Case0)
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+# Set metadata and formula for LimROTS analysis
+meta.info <- "Conc."
+B <- 1000  # Number of bootstrap samples
+K <- dim(UPS1.Case0)[1]/4  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "Conc."
+formula.str <- "Conc."
+# Run LimROTS analysis with trend and robust settings enabled
+rots.result <- LimROTS(data.exp = UPS1.Case0, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
+saveRDS(rots.result, "../LimROTS paper/UPS1.Case0.ROTS.rds")
+```
+
+## Run limma
+
+```{r}
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+data("UPS1.Case0")
+data <- assay(UPS1.Case0)
+formula.str <- "~ 0 + Conc."
+meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("Conc.1-Conc.2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
+saveRDS(limma.result, "../LimROTS paper/UPS1.Case0.limma.rds")
+```
+
+## Comparison
+
+```{r}
+library(caret)
+
+TP <- UPS1.Case0@elementMetadata@listData[["GeneID"]]
+TP <- TP[grepl("HUMAN" , TP)]
+
+predictions_limrots <- limrots.result$FDR < 0.01
+predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
+true_labels_limrots <- ifelse(row.names(limrots.result$data) %in% TP , TRUE, FALSE)
+true_labels_limrots <- factor(true_labels_limrots, levels = c(TRUE, FALSE))
+conf_matrix_limrots <- confusionMatrix(predictions_limrots, true_labels_limrots)
+
+predictions_rots <- rots.result$FDR < 0.01
+predictions_rots <- factor(predictions_rots, levels = c(TRUE, FALSE))
+true_labels_rots <- ifelse(row.names(rots.result$data) %in% TP , TRUE, FALSE)
+true_labels_rots <- factor(true_labels_rots, levels = c(TRUE, FALSE))
+conf_matrix_rots <- confusionMatrix(predictions_rots, true_labels_rots)
+
+predictions_limma <- limma.result$adj.P.Val < 0.01
+predictions_limma <- factor(predictions_limma, levels = c(TRUE, FALSE))
+true_labels_limma <- ifelse(row.names(limma.result) %in% TP , TRUE, FALSE)
+true_labels_limma <- factor(true_labels_limma, levels = c(TRUE, FALSE))
+conf_matrix_limma <- confusionMatrix(predictions_limma, true_labels_limma)
+
+```
+
+```{r}
+library(dplyr, quietly = T, warn.conflicts = T)  
+library(ggplot2)
+
+extract_metrics <- function(conf_matrix, method_name) {
+  metrics <- c(
+    conf_matrix$byClass["Sensitivity"],
+    conf_matrix$byClass["Specificity"],
+    conf_matrix$byClass["Pos Pred Value"],
+    conf_matrix$byClass["Neg Pred Value"],
+    conf_matrix$byClass["F1"],
+    conf_matrix$byClass["Balanced Accuracy"]
+  )
+
+  data.frame(Method = method_name, Metric = names(metrics), Value = as.numeric(metrics))
+}
+
+metrics_limrots <- extract_metrics(conf_matrix_limrots, "limROTS")
+metrics_rots <- extract_metrics(conf_matrix_rots, "ROTS")
+metrics_limma <- extract_metrics(conf_matrix_limma, "limma")
+
+all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma)
+
+ggplot(all_metrics, aes(x = Metric, y = Value, fill = Method)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  theme_bw() +
+  labs(title = "Comparison of Performance Metrics",
+       y = "Value", x = "Metric") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+-   **ROTS** and **limROTS** perform almost equally well across all metrics, with very high sensitivity, specificity, PPV, NPV, and balanced accuracy.
+However, ROTS edges slightly ahead in F1 score and PPV.
+
+-   **limma** performs significantly worse than both **limROTS** and **ROTS**, especially in PPV and F1 score, making it less reliable for this particular dataset and analysis task.
+The lower sensitivity and specificity suggest that limma struggles with both true positive and true negative identification compared to the other two methods.
+
+-   Previous studies have demonstrated that ROTS outperforms limma in similar scenarios across various types of omics data.
+This case study is intended to illustrate that the new LimROTS implementation can achieve comparable performance to ROTS when applied to this type of data, particularly in a straightforward experimental design where only two groups are being compared and no covariates are considered.
+
+-   In the upcoming case studies, we will demonstrate how LimROTS can outperform ROTS, limma, and ANCOVA in the analysis of IDA proteomics datasets.
+
+# UPS1 Case Study 1
+
+## Run LimROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+# Load the dataset
+data("UPS1.Case1")
+print(UPS1.Case1)
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+# Set metadata and formula for LimROTS analysis
+meta.info <- c("Conc." , "inj")
+B <- 1000  # Number of bootstrap samples
+K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 10  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "Conc."
+formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
+# Run LimROTS analysis with trend and robust settings enabled
+limrots.result <- LimROTS(data.exp = UPS1.Case1, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+saveRDS(limrots.result, "../LimROTS paper/UPS1.Case1.LimROTS.rds")
+```
+
+## Run ROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+# Load the dataset
+data("UPS1.Case1")
+print(UPS1.Case1)
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+# Set metadata and formula for LimROTS analysis
+meta.info <- c("Conc." , "inj")
+B <- 1000  # Number of bootstrap samples
+K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "Conc."
+formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
+# Run LimROTS analysis with trend and robust settings enabled
+rots.result <- LimROTS(data.exp = UPS1.Case1, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
+saveRDS(rots.result, "../LimROTS paper/UPS1.Case1.ROTS.rds")
+```
+
+## Run limma
+
+```{r}
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+data("UPS1.Case1")
+data <- assay(UPS1.Case1)
+formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"
+meta.info <- data.frame(colData(UPS1.Case1), check.names = FALSE, row.names = row.names(colData(UPS1.Case1)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("Conc.1-Conc.2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
+saveRDS(limma.result, "../LimROTS paper/UPS1.Case1.limma.rds")
+```
+
+## Comparison
+
+```{r}
+library(tidyverse, quietly = T, warn.conflicts = T)
+
+alpha_thresholds <- seq(0, 0.2, by = 0.001)
+significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$FDR < alpha))
+significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
+significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+
+plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma)
+
+plot_data.melt <-  plot_data %>%
+  pivot_longer(cols = starts_with("significant"), names_to = "Method", values_to = "FDR")
+
+# Plot using ggplot2
+ggplot(plot_data.melt, aes(x = Alpha, y = FDR, color = Method)) +
+  geom_line( size = 1) +
+  theme_minimal() +
+  labs(title = "Significant Proteins vs FDR",
+       x = "FDR",
+       y = "Number of Significant Proteins") +
+   geom_vline(xintercept = 0.01, linetype="dotted", 
+                color = "red", size=1) + 
+    geom_vline(xintercept = 0.05, linetype="dotted", 
+                color = "red", size=1)
+```
+
+Case Study 1 presents a more complex situation where UPS1 human proteins are removed, so no significant differences are expected among the E.
+coli proteins.
+The model has been adjusted to account for analytical replicates from different injections.
+The figure demonstrates that LimROTS and limma handle the data similarly in terms of the number of significant proteins.
+It is important to note that comparing ROTS in this case is not entirely fair, as ROTS was not designed to account for covariates.
+However, it is included in this case study to highlight the importance of considering covariates in differential protein analysis, were LimROTS will be more suitable to be used.
+
+# UPS1 Case Study 2
+
+In the final case study for IDA proteomics data, we will explore a more intricate experimental design, the data will be sourced from two software programs, Spectronaut and ScaffoldDIA.
+Then a synthetic, unbalanced fake batches assigned to the samples.
+The assignment follows the ratio: `c(rep("M", 9), rep("F", 3), rep("M", 3), rep("F", 9), rep("M", 9), rep("F", 3), rep("M", 3), rep("F", 9))`.
+Additionally, 100 E.
+coli proteins were randomly selected, and an effect size of 10 was added to each.
+The model was then run using these fake batches.
+The expected outcome is that only the UPS1 human proteins will be identified as truly significant, while none of the remaining proteins should show significant differences between the concentration groups.\
+\
+This scenario resembles a real-world case where the experiment involves unbalanced batch assignments or, for instance, an uneven gender ratio among the samples.
+LimROTS, limma and ANCOVA will be compared in that case study.
+
+## Run LimROTS
 
@@
 # Load the dataset
-data("UPS1.Case0")
-print(UPS1.Case0)
+data("UPS1.Case4")
+print(UPS1.Case4)
 
@@
 # Set metadata and formula for LimROTS analysis
-meta.info <- "Conc."
+meta.info <- c("Conc." , "tool" , "fake.batch")
 B <- 1000  # Number of bootstrap samples
-K <- dim(UPS1.Case0)[1]/4  # Set the value for K based on the data size
+K <- dim(UPS1.Case4)[1]/4 # Set the value for K based on the data size
 K <- floor(K) 
-num_cores <- 10  # Number of cores for parallel processing
+num_cores <- 30  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
 group.name <- "Conc."
-formula.str <- "~ 0 + Conc."  # Formula for group comparison
+formula.str <- "~ 0 + Conc. + tool + fake.batch"  # Formula for group comparison
 
 # Run LimROTS analysis with trend and robust settings enabled
-rots.result <- LimROTS(data.exp = UPS1.Case0, 
+limrots.result <- LimROTS(data.exp = UPS1.Case4, 
                           B = B, K = K, meta.info = meta.info, 
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+saveRDS(limrots.result, "../LimROTS paper/UPS1.Case2.LimROTS.rds")
 ```
+
+## Run limma
+
+```{r}
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+data("UPS1.Case4")
+data <- assay(UPS1.Case4)
+formula.str <- "~ 0 + Conc. + tool + fake.batch"
+meta.info <- data.frame(colData(UPS1.Case4), check.names = FALSE, row.names = row.names(colData(UPS1.Case4)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("Conc.1-Conc.2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
+saveRDS(limma.result, "../LimROTS paper/UPS1.Case2.limma.rds")
+```
+
+Run ANCOVA
+
+```{r}
+formula.str <- "~ Conc. + tool + fake.batch"
+data("UPS1.Case4")
+data <- assay(UPS1.Case4)
+meta.info <- data.frame(colData(UPS1.Case4), check.names = FALSE, row.names = row.names(colData(UPS1.Case4)))
+meta.info$sample.id <- NULL
+ancova.result <- data.frame()
+for(i in 1:nrow(data)){
+  df.temp <- data[i,]
+  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
+                     tool = meta.info$tool, fake.batch = meta.info$fake.batch )
+  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
+  coef_table <- anova(ancova_model)
+  coef_table <- data.frame(coef_table)
+  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
+  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
+}
+ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
+
+```
+
+## Comparison
+
+```{r}
+library(caret)
+
+TP <- UPS1.Case4@elementMetadata@listData[["GeneID"]]
+TP <- TP[grepl("HUMAN" , TP)]
+
+predictions_limrots <- limrots.result$FDR < 0.01
+predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
+true_labels_limrots <- ifelse(row.names(limrots.result$data) %in% TP , TRUE, FALSE)
+true_labels_limrots <- factor(true_labels_limrots, levels = c(TRUE, FALSE))
+conf_matrix_limrots <- confusionMatrix(predictions_limrots, true_labels_limrots)
+
+predictions_limma <- limma.result$adj.P.Val < 0.01
+predictions_limma <- factor(predictions_limma, levels = c(TRUE, FALSE))
+true_labels_limma <- ifelse(row.names(limma.result) %in% TP , TRUE, FALSE)
+true_labels_limma <- factor(true_labels_limma, levels = c(TRUE, FALSE))
+true_labels_limma <- confusionMatrix(predictions_limma, true_labels_limma)
+
+predictions_ancova <- ancova.result$FDR < 0.01
+predictions_ancova <- factor(predictions_ancova, levels = c(TRUE, FALSE))
+true_labels_ancova <- ifelse(ancova.result$protein %in% TP , TRUE, FALSE)
+true_labels_ancova <- factor(true_labels_ancova, levels = c(TRUE, FALSE))
+conf_matrix_ancova <- confusionMatrix(predictions_ancova, true_labels_ancova)
+```
+
+```{r}
+library(dplyr, quietly = T, warn.conflicts = T)  
+library(ggplot2)
+
+extract_metrics <- function(conf_matrix, method_name) {
+  metrics <- c(
+    conf_matrix$byClass["Sensitivity"],
+    conf_matrix$byClass["Specificity"],
+    conf_matrix$byClass["Pos Pred Value"],
+    conf_matrix$byClass["Neg Pred Value"],
+    conf_matrix$byClass["F1"],
+    conf_matrix$byClass["Balanced Accuracy"]
+  )
+
+  data.frame(Method = method_name, Metric = names(metrics), Value = as.numeric(metrics))
+}
+
+metrics_limrots <- extract_metrics(conf_matrix_limrots, "limROTS")
+metrics_limma <- extract_metrics(conf_matrix_limma, "limma")
+metrics_ancova <- extract_metrics(conf_matrix_ancova, "ANCOVA")
+
+all_metrics <- bind_rows(metrics_limrots, metrics_limma, metrics_ancova)
+
+ggplot(all_metrics, aes(x = Metric, y = Value, fill = Method)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  theme_bw() +
+  labs(title = "Comparison of Performance Metrics",
+       y = "Value", x = "Metric") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+-   **limROTS** is clearly the best performer, with strong results across sensitivity, specificity, PPV, and F1 score, indicating a well-balanced method with accurate classification of significant and non-significant proteins.
+
+-   **limma** and **ANCOVA** both achieve perfect sensitivity and NPV, but their extremely poor specificity and PPV suggest they misclassify many proteins as significant, leading to a high number of false positives.
+This results in very low F1 scores and reduced balanced accuracy, making them unreliable for this particular case study.
+
+In biomarker and drug discovery, particularly during the discovery phase using omics approaches, accurate identification of true significant features is crucial to streamline subsequent steps, making them faster, more precise, and reliable.
+The selection of an appropriate statistical tool is a critical decision, given the multitude of available methods.
+In the context of DIA proteomics, which is an ideal experimental setup for early biomarker and drug discovery phases, many previous studies have relied on tools such as ANCOVA and limma, which were developed for broader omics applications.
+While these tools may be acceptable, more specialized methods should be employed for greater accuracy.
+LimROTS offers a robust statistical approach for identifying true significant proteins in DIA experiments, minimizing false positive discoveries.
+LimROTS excels by providing high balanced accuracy, strong sensitivity, and excellent specificity, making it a reliable tool for these critical phases of discovery.\
+In theory, LimROTS is applicable to other types of omics, such as transcriptomics and metabolomics. However, further case studies are needed to validate its effectiveness in these areas.

### 2024-09-13 tomakeapackae.r

@@ -15,7 +15,7 @@ library(devtools)
 
 devtools::check()
 devtools::build(vignettes = FALSE)
- devtools::build_vignettes()
+devtools::build_vignettes()
 
 library(BiocCheck)
 BiocCheck()

### 2024-09-13 paper.qmd
.

@@
 
-```{r}
+```{r results = "hide" , message = FALSE, warning = FALSE}
 # Load necessary packages
@@
 
-```{r}
+```{r results = "hide" , message = FALSE, warning = FALSE}
 library(limma)
@@
 
-```{r}
-library(caret)
+```{r results = "hide" , message = FALSE, warning = FALSE}
+library(caret, quietly = T, warn.conflicts = T)
 
@@
 
-```{r}
+```{r results = "hide" , message = FALSE, warning = FALSE}
 
@@
 
-Run ANCOVA
+## Run ANCOVA
 
-```{r}
+```{r results = "hide" , message = FALSE, warning = FALSE}
 formula.str <- "~ Conc. + tool + fake.batch"
@@
 
-```{r}
+```{r results = "hide" , message = FALSE, warning = FALSE}
 library(caret)

### 2024-09-14 paper.qmd
.

@@
 
+## Run t-test
+
+```{r}
+library(SummarizedExperiment, quietly = T, warn.conflicts = F)
+data("UPS1.Case0")
+data <- assay(UPS1.Case0)
+meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
+t.test.result <- data.frame()
+
+for(i in 1:nrow(data)){
+  p <- t.test(data[i,1:12], data[i,13:24])
+  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
+}
+
+t.test.result$fdr <- p.adjust(t.test.result$p.value)
+saveRDS(t.test.result, "../LimROTS paper/UPS1.Case0.t.test.rds")
+```
+
 ## Comparison
@@
 library(caret, quietly = T, warn.conflicts = T)
+
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case0.LimROTS.rds")
+rots.result <- readRDS("../LimROTS paper/UPS1.Case0.ROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case0.limma.rds")
+t.test.result <- readRDS("../LimROTS paper/UPS1.Case0.t.test.rds")
 
@@
 
+predictions_t.test <- t.test.result$fdr < 0.01
+predictions_t.test <- factor(predictions_t.test, levels = c(TRUE, FALSE))
+true_labels_t.test <- ifelse(t.test.result$protein %in% TP , TRUE, FALSE)
+true_labels_t.test <- factor(true_labels_t.test, levels = c(TRUE, FALSE))
+conf_matrix_t.test <- confusionMatrix(predictions_t.test, true_labels_t.test)
+
 ```
@@
 library(ggplot2)
+library(ggsci)
 
@@
 metrics_limma <- extract_metrics(conf_matrix_limma, "limma")
+metrics_t.test <- extract_metrics(conf_matrix_t.test, "t.test")
 
-all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma)
+all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma,metrics_t.test)
 
@@
   theme_bw() +
-  labs(title = "Comparison of Performance Metrics",
+  labs(title = "Comparison of Performance Case Study0",
        y = "Value", x = "Metric") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_npg() + 
+    theme(
+    plot.title = element_text(face = "bold", color = "black"),
+    axis.title = element_text(face = "bold", color = "black"),
+    axis.text = element_text(face = "bold", color = "black"),
+    axis.ticks = element_line(color = "black")
+  )
 

### 2024-09-14 paper.qmd
.

@@
 
-```{r results = "hide" , message = FALSE, warning = FALSE}
+```{r results = "hide" , message = FALSE, warning = FALSE, eval = TRUE}
 # Load necessary packages
@@
 
-```{r results = "hide" , message = FALSE, warning = FALSE}
+```{r results = "hide" , message = FALSE, warning = FALSE, eval=FALSE}
 # Load necessary packages
@@
 
-```{r results = "hide" , message = FALSE, warning = FALSE}
+```{r results = "hide" , message = FALSE, warning = FALSE, eval=FALSE}
 library(limma)
@@
 
-```{r}
+```{r eval=FALSE}
 library(SummarizedExperiment, quietly = T, warn.conflicts = F)
@@
 
+## Run ANCOVA
+
+```{r results = "hide" , message = FALSE, warning = FALSE, eval=FLASE}
+
+formula.str <- "~Conc.+ inj + Conc.*inj"
+data("UPS1.Case1")
+data <- assay(UPS1.Case1)
+meta.info <- data.frame(colData(UPS1.Case1), check.names = FALSE, row.names = row.names(colData(UPS1.Case1)))
+meta.info$sample.id <- NULL
+ancova.result <- data.frame()
+for(i in 1:nrow(data)){
+  df.temp <- data[i,]
+  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
+                     inj = meta.info$inj)
+  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
+  coef_table <- anova(ancova_model)
+  coef_table <- data.frame(coef_table)
+  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
+  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
+}
+ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
+saveRDS(ancova.result, "../LimROTS paper/UPS1.Case1.ancova.rds")
+```
+
 ## Comparison
@@
 library(tidyverse, quietly = T, warn.conflicts = T)
+library(ggsci)
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case1.LimROTS.rds")
+rots.result <- readRDS("../LimROTS paper/UPS1.Case1.ROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case1.limma.rds")
+ancova.result <- readRDS("../LimROTS paper/UPS1.Case1.ancova.rds")
 
@@
 significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+significant_ancova <- sapply(alpha_thresholds, function(alpha) sum(ancova.result$FDR < alpha))
 
 plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
-                        significant_ROTS = significant_ROTS , significant_limma = significant_limma)
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma  ,significant_ancova = significant_ancova)
 
@@
     geom_vline(xintercept = 0.05, linetype="dotted", 
-                color = "red", size=1)
+                color = "red", size=1)+
+    scale_color_npg() + 
+    theme(
+    plot.title = element_text(face = "bold", color = "black"),
+    axis.title = element_text(face = "bold", color = "black"),
+    axis.text = element_text(face = "bold", color = "black"),
+    axis.ticks = element_line(color = "black")
+  )
 ```

### 2024-09-14 paper.qmd
.

@@
   theme_minimal() +
-  labs(title = "Significant Proteins vs FDR",
+  labs(title = "Significant Proteins vs FDR Case study 1",
        x = "FDR",
@@
 
-```{r results = "hide" , message = FALSE, warning = FALSE}
+```{r results = "hide" , message = FALSE, warning = FALSE, eval=FALSE}
 
@@
 
-```{r results = "hide" , message = FALSE, warning = FALSE}
+```{r results = "hide" , message = FALSE, warning = FALSE, eval=FALSE}
 formula.str <- "~ Conc. + tool + fake.batch"
@@
 ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
-
+saveRDS(ancova.result, "../LimROTS paper/UPS1.Case2.ancova.rds")
 ```
@@
 library(caret)
+
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case2.LimROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case2.limma.rds")
+ancova.result <- readRDS("../LimROTS paper/UPS1.Case2.ancova.rds")
 
@@
 true_labels_limma <- factor(true_labels_limma, levels = c(TRUE, FALSE))
-true_labels_limma <- confusionMatrix(predictions_limma, true_labels_limma)
+conf_matrix_limma <- confusionMatrix(predictions_limma, true_labels_limma)
 
@@
   theme_bw() +
-  labs(title = "Comparison of Performance Metrics",
+  labs(title = "Comparison of Performance case study 2",
        y = "Value", x = "Metric") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
+      scale_fill_npg() + 
+    theme(
+    plot.title = element_text(face = "bold", color = "black"),
+    axis.title = element_text(face = "bold", color = "black"),
+    axis.text = element_text(face = "bold", color = "black"),
+    axis.ticks = element_line(color = "black")
+  )
 ```

### 2024-09-16 paper.qmd
.

--- 2024-09-14 paper.qmd
@@
 
-## Run t-test
-
-```{r}
-library(SummarizedExperiment, quietly = T, warn.conflicts = F)
-data("UPS1.Case0")
-data <- assay(UPS1.Case0)
-meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
-t.test.result <- data.frame()
-
-for(i in 1:nrow(data)){
-  p <- t.test(data[i,1:12], data[i,13:24])
-  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
-}
-
-t.test.result$fdr <- p.adjust(t.test.result$p.value)
-saveRDS(t.test.result, "../LimROTS paper/UPS1.Case0.t.test.rds")
-```
-
 ## Comparison
@@
 library(caret, quietly = T, warn.conflicts = T)
-
-limrots.result <- readRDS("../LimROTS paper/UPS1.Case0.LimROTS.rds")
-rots.result <- readRDS("../LimROTS paper/UPS1.Case0.ROTS.rds")
-limma.result <- readRDS("../LimROTS paper/UPS1.Case0.limma.rds")
-t.test.result <- readRDS("../LimROTS paper/UPS1.Case0.t.test.rds")
 
@@
 
-predictions_t.test <- t.test.result$fdr < 0.01
-predictions_t.test <- factor(predictions_t.test, levels = c(TRUE, FALSE))
-true_labels_t.test <- ifelse(t.test.result$protein %in% TP , TRUE, FALSE)
-true_labels_t.test <- factor(true_labels_t.test, levels = c(TRUE, FALSE))
-conf_matrix_t.test <- confusionMatrix(predictions_t.test, true_labels_t.test)
-
 ```
@@
 library(ggplot2)
-library(ggsci)
 
@@
 metrics_limma <- extract_metrics(conf_matrix_limma, "limma")
-metrics_t.test <- extract_metrics(conf_matrix_t.test, "t.test")
 
-all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma,metrics_t.test)
+all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma)
 
@@
   theme_bw() +
-  labs(title = "Comparison of Performance Case Study0",
+  labs(title = "Comparison of Performance Metrics",
        y = "Value", x = "Metric") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_npg() + 
-    theme(
-    plot.title = element_text(face = "bold", color = "black"),
-    axis.title = element_text(face = "bold", color = "black"),
-    axis.text = element_text(face = "bold", color = "black"),
-    axis.ticks = element_line(color = "black")
-  )
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 In theory, LimROTS is applicable to other types of omics, such as transcriptomics and metabolomics. However, further case studies are needed to validate its effectiveness in these areas.
+
+# BM21 Case study 3
+
+## Run LimROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+
+# Load the dataset
+load("../LimROTS paper/metabolomics/BM21.Case0.rda")
+
+print(BM21.Case0)
+
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- "mixing.ratio"
+B <- 1000  # Number of bootstrap samples
+K <- dim(BM21.Case0)[1]/2  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 10  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "mixing.ratio"
+formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison
+
+# Run LimROTS analysis with trend and robust settings enabled
+limrots.result <- LimROTS(data.exp = BM21.Case0, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+```
+
+```{r}
+length(which(limrots.result$FDR < 0.05))
+summary(limrots.result$q_values)
+plot(limrots.result$q_values)
+hist(limrots.result$FDR)
+
+limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
+                                LimROTS.FC = limrots.result$corrected.logf,
+                                FDR = limrots.result$FDR)
+```
+
+## Run ROTS
+
+```{r}
+# Load the dataset
+load("../LimROTS paper/metabolomics/BM21.Case0.rda")
+  library(SummarizedExperiment)
+library(doParallel)
+library(LimROTS)
+print(BM21.Case0)
+
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- "mixing.ratio"
+B <- 1000  # Number of bootstrap samples
+K <- dim(BM21.Case0)[1]/2 # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 10  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "mixing.ratio"
+formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison
+
+# Run LimROTS analysis with trend and robust settings enabled
+rots.result <- LimROTS(data.exp = BM21.Case0, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = T)
+```
+
+```{r}
+length(which(rots.result$FDR < 0.01))
+summary(rots.result$q_values)
+plot(rots.result$q_values)
+hist(rots.result$q_values)
+hist(rots.result$p)
+hist(rots.result$FDR)
+```
+
+## Run Limma
+
+```{r}
+
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+
+load("../LimROTS paper/metabolomics/BM21.Case0.rda")
+
+print("BM21.Case0")
+data <- assay(BM21.Case0)
+formula.str <- "~ 0 + mixing.ratio"
+meta.info <- data.frame(colData(BM21.Case0), check.names = FALSE, row.names = row.names(colData(BM21.Case0)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=T, robust=T)
+limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
+
+length(which(limma.result$adj.P.Val < 0.01))
+
+saveRDS(limma.result, "../LimROTS paper/BM21.case0.limma.rds")
+```

### 2024-09-18 paper.qmd
.

@@
 
+## Run t-test
+
+```{r}
+
+library(SummarizedExperiment, quietly = T, warn.conflicts = F)
+data("UPS1.Case0")
+data <- assay(UPS1.Case0)
+meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
+t.test.result <- data.frame()
+
+for(i in 1:nrow(data)){
+  p <- t.test(data[i,1:12], data[i,13:24])
+  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
+}
+
+t.test.result$fdr <- p.adjust(t.test.result$p.value)
+saveRDS(t.test.result, "../LimROTS paper/UPS1.Case0.t.test.rds")
+
+```
+
 ## Comparison
@@
 
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case0.LimROTS.rds")
+rots.result <- readRDS("../LimROTS paper/UPS1.Case0.ROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case0.limma.rds")
+t.test.result <- readRDS("../LimROTS paper/UPS1.Case0.t.test.rds")
+
 TP <- UPS1.Case0@elementMetadata@listData[["GeneID"]]
@@
 
-predictions_limrots <- limrots.result$FDR < 0.01
+predictions_limrots <- limrots.result$q_values$qvalues < 0.01
 predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
@@
 K <- floor(K) 
-num_cores <- 10  # Number of cores for parallel processing
+num_cores <- 50  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
 K <- floor(K) 
-num_cores <- 30  # Number of cores for parallel processing
+num_cores <- 50  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
 
+## Run ANCOVA
+
+```{r}
+
+formula.str <- "~Conc.+ inj + Conc.*inj"
+data("UPS1.Case1")
+data <- assay(UPS1.Case1)
+meta.info <- data.frame(colData(UPS1.Case1), check.names = FALSE, row.names = row.names(colData(UPS1.Case1)))
+meta.info$sample.id <- NULL
+ancova.result <- data.frame()
+for(i in 1:nrow(data)){
+  df.temp <- data[i,]
+  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
+                     inj = meta.info$inj)
+  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
+  coef_table <- anova(ancova_model)
+  coef_table <- data.frame(coef_table)
+  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
+  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
+}
+ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
+saveRDS(ancova.result, "../LimROTS paper/UPS1.Case1.ancova.rds")
+```
+
 ## Comparison
@@
 library(tidyverse, quietly = T, warn.conflicts = T)
+library(ggsci)
+
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case1.LimROTS.rds")
+rots.result <- readRDS("../LimROTS paper/UPS1.Case1.ROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case1.limma.rds")
+ancova.result <- readRDS("../LimROTS paper/UPS1.Case1.ancova.rds")
 
 alpha_thresholds <- seq(0, 0.2, by = 0.001)
-significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$FDR < alpha))
+significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$q_values$qvalues < alpha))
 significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
 significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+significant_ancova <- sapply(alpha_thresholds, function(alpha) sum(ancova.result$FDR < alpha))
 
 plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
-                        significant_ROTS = significant_ROTS , significant_limma = significant_limma)
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma  ,significant_ancova = significant_ancova)
 
@@
 ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
+ancova.result <- saveRDS("../LimROTS paper/UPS1.Case2.ancova.rds")
 
@@
 
+limrots.result <- readRDS("../LimROTS paper/UPS1.Case2.LimROTS.rds")
+limma.result <- readRDS("../LimROTS paper/UPS1.Case2.limma.rds")
+ancova.result <- readRDS("../LimROTS paper/UPS1.Case2.ancova.rds")
+
 TP <- UPS1.Case4@elementMetadata@listData[["GeneID"]]
@@
 K <- floor(K) 
-num_cores <- 10  # Number of cores for parallel processing
+num_cores <- 30   # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
 formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison
+
+trace(LimROTS::LimROTS , edit = T)
+
+trace(LimROTS::calculateFDR , edit = T)
 
@@
 ```{r}
-length(which(limrots.result$FDR < 0.05))
+length(which(limrots.result$FDR < 0.01))
+
+length(which(limrots.result$BH.pvalue < 0.01))
+
 summary(limrots.result$q_values)
@@
 
-limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
+limrots.result.df <- data.frame(row.names  = row.names(limrots.result$data),
                                 LimROTS.FC = limrots.result$corrected.logf,
-                                FDR = limrots.result$FDR)
+                                FDR = limrots.result$FDR, p.value = limrots.result$pvalue, bh = limrots.result$BH.pvalue)
 ```
@@
 
+trace(LimROTS::calculateFDR , edit = T)
+
+#trace(LimROTS , edit = T)
 # Run LimROTS analysis with trend and robust settings enabled

### 2024-09-18 paper.qmd
.

@@
 
-predictions_limrots <- limrots.result$q_values$qvalues < 0.01
+predictions_limrots <- limrots.result$q_values$qvalues < 0.05
 predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
@@
 
-predictions_rots <- rots.result$FDR < 0.01
+predictions_rots <- rots.result$FDR < 0.05
 predictions_rots <- factor(predictions_rots, levels = c(TRUE, FALSE))
@@
 
-predictions_limma <- limma.result$adj.P.Val < 0.01
+predictions_limma <- limma.result$adj.P.Val < 0.05
 predictions_limma <- factor(predictions_limma, levels = c(TRUE, FALSE))
@@
 
-predictions_t.test <- t.test.result$fdr < 0.01
+predictions_t.test <- t.test.result$fdr < 0.05
 predictions_t.test <- factor(predictions_t.test, levels = c(TRUE, FALSE))
@@
 -   **ROTS** and **limROTS** perform almost equally well across all metrics, with very high sensitivity, specificity, PPV, NPV, and balanced accuracy.
-However, ROTS edges slightly ahead in F1 score and PPV.
 
@@
 ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
-ancova.result <- saveRDS("../LimROTS paper/UPS1.Case2.ancova.rds")
+saveRDS(ancova.result, "../LimROTS paper/UPS1.Case2.ancova.rds")
 
@@
 
-predictions_ancova <- ancova.result$FDR < 0.01
+predictions_ancova <- ancova.result$FDR < 0.05
 predictions_ancova <- factor(predictions_ancova, levels = c(TRUE, FALSE))
@@
 
-trace(LimROTS::LimROTS , edit = T)
-
-trace(LimROTS::calculateFDR , edit = T)
-
 # Run LimROTS analysis with trend and robust settings enabled
@@
 ```{r}
-length(which(limrots.result$FDR < 0.01))
+length(which(limrots.result$q_values$qvalues < 0.05))
 
-length(which(limrots.result$BH.pvalue < 0.01))
+length(which(limrots.result$BH.pvalue < 0.05))
 
@@
 plot(limrots.result$q_values)
-hist(limrots.result$FDR)
+hist(limrots.result$q_values)
 
@@
 
-trace(LimROTS::calculateFDR , edit = T)
-
 #trace(LimROTS , edit = T)
@@
 ```{r}
-length(which(rots.result$FDR < 0.01))
+length(which(rots.result$FDR < 0.05))
 summary(rots.result$q_values)
@@
 
-length(which(limma.result$adj.P.Val < 0.01))
+length(which(limma.result$adj.P.Val < 0.05))
 
@@
 ```
+
+```{r}
+
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+
+# Load the dataset
+load("../LimROTS paper/metabolomics/BM21.Case1.rda")
+
+print(BM21.Case1)
+
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- "mixing.ratio"
+B <- 1000  # Number of bootstrap samples
+K <- dim(BM21.Case1)[1]/2  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30   # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "mixing.ratio"
+formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison
+
+trace(LimROTS::testStatistic_with_covariates_Fit, edit = T)
+trace(LimROTS::testStatistic_with_covariates_permutating, edit = T)
+trace(LimROTS::testStatistic_with_covariates, edit = T)
+
+# Run LimROTS analysis with trend and robust settings enabled
+limrots.result <- LimROTS(data.exp = BM21.Case1, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+```

### 2024-09-18 R/LM_with_limma.fit.r

@@ -43,7 +43,7 @@
 testStatistic_with_covariates_Fit <- function(data, group.name, meta.info , formula.str,
                                                       trend, robust) {
 
-  combined_data <- cbind(data[[1]], data[[2]])
+  combined_data <- do.call(cbind, data)
   design.matrix <- model.matrix(formula(formula.str), data = meta.info)
   colnames(design.matrix) <- make.names(colnames(design.matrix) )
   fit <- limma::lmFit(combined_data, design.matrix)

### 2024-09-18 R/LM_with_limma.r

@@ -38,7 +38,7 @@
 
 testStatistic_with_covariates <- function(data, group.name, meta.info, formula.str ,
                                           trend, robust) {
-  combined_data <- cbind(data[[1]], data[[2]])
+  combined_data <- do.call(cbind, data)
   colnames(combined_data) <- paste0(colnames(combined_data), "." , seq(1,ncol(combined_data)))
     covariates.p <- data.frame()
     meta.info.temp <- meta.info

### 2024-09-18 R/Bootstrap_functions.r
.

@@ -17,7 +17,8 @@ bootstrapSamples <- function (data, B, labels, paired)
   for (i in 1:B) {
     for (label in unique(labels)) {
       pos <- which(labels == label)
-      samples[i, pos] <- sample(pos, length(pos), replace = TRUE)
+      pos.names <- colnames(data)[pos]
+      samples[i, pos] <- sample(pos.names, length(pos), replace = TRUE)
     }
   }
   if (paired) {

### 2024-09-18 R/LM_with_limma.r

@@ -32,13 +32,14 @@
 #'
 #' @importFrom stats model.matrix formula
 #' @import utils
+#' @importFrom dplyr bind_cols
 #' @importFrom stringr str_split_fixed fixed
 #'
 
 
 testStatistic_with_covariates <- function(data, group.name, meta.info, formula.str ,
                                           trend, robust) {
-  combined_data <- do.call(cbind, data)
+  combined_data <- bind_cols(data)
   colnames(combined_data) <- paste0(colnames(combined_data), "." , seq(1,ncol(combined_data)))
     covariates.p <- data.frame()
     meta.info.temp <- meta.info

### 2024-09-18 R/LimROTS.r

@@ -50,6 +50,7 @@
 #' }
 #' @importFrom limma voom lmFit eBayes
 #' @importFrom stats model.matrix formula p.adjust
+#' @importFrom dplyr bind_cols
 #' @import parallel
 #' @import foreach
 #' @import doRNG

### 2024-09-18 paper.qmd
.

@@
 # Run LimROTS analysis with trend and robust settings enabled
+
+trace(LimROTS, edit = T)
+
 limrots.result <- LimROTS(data.exp = UPS1.Case0, 
@@
 
-trace(LimROTS::testStatistic_with_covariates_Fit, edit = T)
-trace(LimROTS::testStatistic_with_covariates_permutating, edit = T)
-trace(LimROTS::testStatistic_with_covariates, edit = T)
+trace(LimROTS, edit = T)
 
@@
 ```
+
+```{r}
+length(which(limrots.result$q_values$qvalues < 0.05))
+
+length(which(limrots.result$FDR < 0.05))
+
+length(which(limrots.result$BH.pvalue < 0.05))
+
+summary(limrots.result$q_values)
+plot(limrots.result$q_values)
+hist(limrots.result$q_values)
+p.value = limrots.result$pvalue
+hist(p.value)
+
+library(qvalue)
+q <- qvalue(p.value, pi0.method = "bootstrap", lambda  = seq(0.01, 0.95, 0.01))
+plot(q)
+hist(q)
+summary(q)
+
+pi0est(p.value, pi0.method = "bootstrap")
+
+limrots.result.df <- data.frame(row.names  = row.names(limrots.result$data),
+                                LimROTS.FC = limrots.result$corrected.logf,
+                                FDR = limrots.result$FDR, p.value = limrots.result$pvalue, bh = limrots.result$BH.pvalue)
+```
+
+```{r}
+
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+
+load("../LimROTS paper/metabolomics/BM21.Case1.rda")
+
+print("BM21.Case1")
+data <- assay(BM21.Case1)
+formula.str <- "~ 0 + mixing.ratio"
+meta.info <- data.frame(colData(BM21.Case1), check.names = FALSE, row.names = row.names(colData(BM21.Case1)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=T, robust=T)
+limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
+
+length(which(limma.result$adj.P.Val < 0.05))
+
+saveRDS(limma.result, "../LimROTS paper/BM21.case0.limma.rds")
+```

### 2024-09-18 R/LM_with_limma.fit.r

@@ -44,7 +44,11 @@
 testStatistic_with_covariates_Fit <- function(data, group.name, meta.info , formula.str,
                                                       trend, robust) {
 
-  combined_data <- bind_cols(data)
+  combined_data <- data.frame(check.rows = F, check.names = F , none = rep("none" , nrow(data[[1]])))
+  for(k.list in names(data)){
+    combined_data <- cbind(combined_data, data.frame(data[[k.list]], check.rows = F, check.names = F))
+  }
+  combined_data <- combined_data[,-1]
   design.matrix <- model.matrix(formula(formula.str), data = meta.info)
   colnames(design.matrix) <- make.names(colnames(design.matrix) )
   fit <- limma::lmFit(combined_data, design.matrix)

### 2024-09-18 R/LM_with_limma.permuting.r

@@ -40,7 +40,11 @@
 
 testStatistic_with_covariates_permutating <- function(data, group.name, meta.info, formula.str ,
                                           trend, robust) {
-  combined_data <- bind_cols(data)
+  combined_data <- data.frame(check.rows = F, check.names = F , none = rep("none" , nrow(data[[1]])))
+  for(k.list in names(data)){
+    combined_data <- cbind(combined_data, data.frame(data[[k.list]], check.rows = F, check.names = F))
+  }
+  combined_data <- combined_data[,-1]
   colnames(combined_data) <- paste0(colnames(combined_data), "." , seq(1,ncol(combined_data)))
   covariates.p <- data.frame()
   meta.info.temp <- meta.info

### 2024-09-18 R/LM_with_limma.r

@@ -39,7 +39,11 @@
 
 testStatistic_with_covariates <- function(data, group.name, meta.info, formula.str ,
                                           trend, robust) {
-  combined_data <- bind_cols(data)
+  combined_data <- data.frame(check.rows = F, check.names = F , none = rep("none" , nrow(data[[1]])))
+  for(k.list in names(data)){
+    combined_data <- cbind(combined_data, data.frame(data[[k.list]], check.rows = F, check.names = F))
+  }
+  combined_data <- combined_data[,-1]
   colnames(combined_data) <- paste0(colnames(combined_data), "." , seq(1,ncol(combined_data)))
     covariates.p <- data.frame()
     meta.info.temp <- meta.info

### 2024-09-18 paper.qmd
.

@@
 # Run LimROTS analysis with trend and robust settings enabled
-
-trace(LimROTS, edit = T)
 
@@
 
-trace(LimROTS, edit = T)
-
 # Run LimROTS analysis with trend and robust settings enabled

### 2024-09-18 paper.qmd
.

@@
 ```
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+
+# Load the dataset
+load("../LimROTS paper/metabolomics/BM21.Case3.rda")
+
+print(BM21.Case3)
+
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- c("mixing.ratio" , "fake.batch")
+B <- 1000  # Number of bootstrap samples
+K <- dim(BM21.Case3)[1]/2  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30   # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "mixing.ratio"
+formula.str <- "~ 0 + mixing.ratio + fake.batch"  # Formula for group comparison
+
+# Run LimROTS analysis with trend and robust settings enabled
+limrots.result <- LimROTS(data.exp = BM21.Case3, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+```
+
+```{r}
+
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+
+load("../LimROTS paper/metabolomics/BM21.Case3.rda")
+
+print("BM21.Case3")
+data <- assay(BM21.Case3)
+formula.str <- "~ 0 + mixing.ratio + fake.batch"
+meta.info <- data.frame(colData(BM21.Case3), check.names = FALSE, row.names = row.names(colData(BM21.Case3)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=T, robust=T)
+limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
+
+length(which(limma.result$adj.P.Val < 0.05))
+
+limma.result <- limma.result[limma.result$adj.P.Val < 0.05,]
+
+length(which(grepl("_pluseffect", row.names(limma.result)) ))
+
+length(which(grepl("synthetic", row.names(limma.result)) ))
+
+```

### 2024-09-18 paper.qmd
.

@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case3)[1]/2  # Set the value for K based on the data size
+K <- dim(BM21.Case3)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+
+summary(limrots.result$q_values)
+plot(limrots.result$q_values)
+hist(limrots.result$q_values)
+
+limrots.result.df <- data.frame(row.names = row.names(limrots.result$data), q.value = limrots.result$q_values$qvalues , fdr = limrots.result$FDR,
+                                bh = limrots.result$BH.pvalue , fc = limrots.result$corrected.logfc)
+
+length(which(limrots.result.df$q.value < 0.05))
+length(which(limrots.result.df$fdr < 0.05))
+
+limrots.result.df <- limrots.result.df[limrots.result.df$q.value < 0.05,]
+
+length(which(grepl("_pluseffect", row.names(limrots.result.df)) ))
+
+length(which(grepl("synthetic", row.names(limrots.result.df)) ))
 ```

### 2024-09-18 paper.qmd
.

@@
 
-# UPS1 Case Study 0
+# UPS1-spiked DataSet
+
+## UPS1 Case Study 0
 
@@
 
-## Run LimROTS
+### Run LimROTS
 
@@
 
-## Run ROTS
+### Run ROTS
 
@@
 
-## Run limma
+### Run limma
 
@@
 
-## Run t-test
+### Run t-test
 
@@
 
-## Comparison
+### Comparison
 
@@
 
-# UPS1 Case Study 1
+## UPS1 Case Study 1
 
-## Run LimROTS
+### Run LimROTS
 
@@
 
-## Run ANCOVA
+### Run ANCOVA
 
@@
 
-# UPS1 Case Study 2
+## UPS1 Case Study 2
 
@@
 
-# BM21 Case study 3
+# BM21 DataSet
 
-## Run LimROTS
+To demonstrate the capability of LimROTS to also handle metabolomics data, the dataset was obtained from the Metabolomics Workbench (ST002454).
+Different matrix ratios of human plasma and vegetable juice were analyzed, and three different datasets were summarized into three *SummarizedExperiment* objects.
+
+## BM21 Case study 3
+
+This dataset contains six different mixing ratios: 1:1024, 1:256, 1:64, and their reverse, each with three technical replicates.
+The samples were grouped into high plasma ratio samples and low ratio samples.
+Peaks with more than 50% missing values were removed, and peaks with a variance less than 1 were excluded.
+Furthermore, only peaks quantified in samples with 100% plasma and 100% vegetable juice were selected.
+Therefore, we expect that all the peaks in this dataset represent true significant differences between the two mixing ratio groups.
+
+### Run LimROTS
 
@@
 # Load the dataset
-load("../LimROTS paper/metabolomics/BM21.Case0.rda")
+data(BM21.Case0)
 
@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case0)[1]/2  # Set the value for K based on the data size
+K <- dim(BM21.Case0)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
@@
 
-```{r}
-length(which(limrots.result$q_values$qvalues < 0.05))
-
-length(which(limrots.result$BH.pvalue < 0.05))
-
-summary(limrots.result$q_values)
-plot(limrots.result$q_values)
-hist(limrots.result$q_values)
-
-limrots.result.df <- data.frame(row.names  = row.names(limrots.result$data),
-                                LimROTS.FC = limrots.result$corrected.logf,
-                                FDR = limrots.result$FDR, p.value = limrots.result$pvalue, bh = limrots.result$BH.pvalue)
-```
-
-## Run ROTS
+### Run ROTS
 
@@
 # Load the dataset
-load("../LimROTS paper/metabolomics/BM21.Case0.rda")
 library(SummarizedExperiment)
@@
 library(LimROTS)
+
+data(BM21.Case0)
 print(BM21.Case0)
@@
 
-```{r}
-length(which(rots.result$FDR < 0.05))
-summary(rots.result$q_values)
-plot(rots.result$q_values)
-hist(rots.result$q_values)
-hist(rots.result$p)
-hist(rots.result$FDR)
-```
-
-## Run Limma
+### Run Limma
 
@@
 
-load("../LimROTS paper/metabolomics/BM21.Case0.rda")
+data(BM21.Case0)
 
@@
 ```
+
+## BM21 Case study 4
 
@@
 
-load("../LimROTS paper/metabolomics/BM21.Case1.rda")
-
-print("BM21.Case1")
+data(BM21.Case1)
+print(BM21.Case1)
 data <- assay(BM21.Case1)
@@
 limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
-
-length(which(limma.result$adj.P.Val < 0.05))
-
-saveRDS(limma.result, "../LimROTS paper/BM21.case0.limma.rds")
 ```

### 2024-09-18 paper.qmd
paperwork

@@
 K <- floor(K) 
-num_cores <- 30   # Number of cores for parallel processing
+num_cores <- 10   # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+
+saveRDS(limrots.result , "paperdata/BM21.Case0.limrots.rds")
 ```
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = T)
+
+saveRDS(rots.result , "paperdata/BM21.Case0.rots.rds")
 ```
@@
 
-length(which(limma.result$adj.P.Val < 0.05))
+saveRDS(limma.result , "paperdata/BM21.Case0.limma.rds")
+```
 
-saveRDS(limma.result, "../LimROTS paper/BM21.case0.limma.rds")
+### Comparison
+
+```{r}
+library(tidyverse, quietly = T, warn.conflicts = T)
+library(ggsci)
+
+limrots.result <- readRDS("paperdata/BM21.Case0.limrots.rds")
+rots.result <- readRDS("paperdata/BM21.Case0.rots.rds")
+limma.result <- readRDS("paperdata/BM21.Case0.limma.rds")
+
+alpha_thresholds <- seq(0, 0.075, by = 0.01)
+significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$q_values$qvalues < alpha))
+significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
+significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+
+plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma )
+
+plot_data.melt <-  plot_data %>%
+  pivot_longer(cols = starts_with("significant"), names_to = "Method", values_to = "FDR")
+
+# Plot using ggplot2
+ggplot(plot_data.melt, aes(x = Alpha, y = FDR, color = Method)) +
+  geom_line( size = 1) +
+  theme_minimal() +
+  labs(title = "Significant Proteins vs FDR Case study 1",
+       x = "FDR",
+       y = "Number of Significant Proteins") +
+   geom_vline(xintercept = 0.01, linetype="dotted", 
+                color = "red", size=1) + 
+    geom_vline(xintercept = 0.05, linetype="dotted", 
+                color = "red", size=1)+
+    scale_color_npg() + 
+    theme(
+    plot.title = element_text(face = "bold", color = "black"),
+    axis.title = element_text(face = "bold", color = "black"),
+    axis.text = element_text(face = "bold", color = "black"),
+    axis.ticks = element_line(color = "black")
+  )
 ```
@@
 ## BM21 Case study 4
+
+### Run LimROTS
 
@@
 # Load the dataset
-load("../LimROTS paper/metabolomics/BM21.Case1.rda")
+data(BM21.Case1)
 
@@
 
-length(which(limrots.result.df$q.value < 0.05))
-length(which(limrots.result.df$fdr < 0.05))
+```
 
-limrots.result.df <- limrots.result.df[limrots.result.df$q.value < 0.05,]
-
-length(which(grepl("_pluseffect", row.names(limrots.result.df)) ))
-
-length(which(grepl("synthetic", row.names(limrots.result.df)) ))
-```
+### Run limma
 

### 2024-09-19 paper.qmd
.

@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case0)[1]/2 # Set the value for K based on the data size
+K <- dim(BM21.Case0)[1]/4 # Set the value for K based on the data size
 K <- floor(K) 
@@
 
+Another feature in LimROTS is the ability to handle more than two groups using moderated F-statistics and ROTS optimization.
+In this example, three different ratio mixtures were extracted from the BM21 dataset and processed similarly to the dataset in case study 3.
+Therefore, we expect to see the maximum number of features that are significantly different with a true effect.
+
 ### Run LimROTS
@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case1)[1]/2  # Set the value for K based on the data size
+K <- dim(BM21.Case1)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+
+saveRDS(limrots.result , "paperdata/BM21.Case1.limrots.rds")
 ```
 
-```{r}
-length(which(limrots.result$q_values$qvalues < 0.05))
-
-length(which(limrots.result$FDR < 0.05))
-
-length(which(limrots.result$BH.pvalue < 0.05))
-
-summary(limrots.result$q_values)
-plot(limrots.result$q_values)
-hist(limrots.result$q_values)
-p.value = limrots.result$pvalue
-hist(p.value)
-
-library(qvalue)
-q <- qvalue(p.value, pi0.method = "bootstrap", lambda  = seq(0.01, 0.95, 0.01))
-plot(q)
-hist(q)
-summary(q)
-
-pi0est(p.value, pi0.method = "bootstrap")
-
-limrots.result.df <- data.frame(row.names  = row.names(limrots.result$data),
-                                LimROTS.FC = limrots.result$corrected.logf,
-                                FDR = limrots.result$FDR, p.value = limrots.result$pvalue, bh = limrots.result$BH.pvalue)
-```
+### Run limma
 
@@
 ```
+
+## BM21 Case study 5
+
+### Run LimROTS
 

### 2024-09-19 paper.qmd
.

@@
 
+### Run t-test
+
+```{r}
+library(SummarizedExperiment, quietly = T, warn.conflicts = F)
+data("BM21.Case0")
+data <- assay(BM21.Case0)
+meta.info <- data.frame(colData(BM21.Case0), check.names = FALSE, row.names = row.names(colData(BM21.Case0)))
+t.test.result <- data.frame()
+
+for(i in 1:nrow(data)){
+  p <- t.test(data[i,c(1:3,13:18)], data[i,4:12])
+  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
+}
+
+t.test.result$fdr <- p.adjust(t.test.result$p.value)
+saveRDS(t.test.result , "paperdata/BM21.Case0.t.test.rds")
+```
+
 ### Comparison
@@
 limma.result <- readRDS("paperdata/BM21.Case0.limma.rds")
+t.test.result <- readRDS("paperdata/BM21.Case0.t.test.rds")
 
@@
 significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+significant_t.test <- sapply(alpha_thresholds, function(alpha) sum(t.test.result$fdr < alpha))
 
 plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
-                        significant_ROTS = significant_ROTS , significant_limma = significant_limma )
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma, significant_t.test = significant_t.test )
 
@@
 saveRDS(limrots.result , "paperdata/BM21.Case1.limrots.rds")
+```
+
+### Run ROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+
+# Load the dataset
+data(BM21.Case1)
+
+print(BM21.Case1)
+
+# Set random seed for reproducibility
+set.seed(1234, kind = "default")
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- "mixing.ratio"
+B <- 1000  # Number of bootstrap samples
+K <- dim(BM21.Case1)[1]/4  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30   # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "mixing.ratio"
+formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison
+
+# Run LimROTS analysis with trend and robust settings enabled
+limrots.result <- LimROTS(data.exp = BM21.Case1, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
+
+saveRDS(limrots.result , "paperdata/BM21.Case1.rots.rds")
 ```

### 2024-09-19 paper.qmd
paper work

@@
 
+### Run samr
+
+```{r}
+library(samr)
+data("UPS1.Case0")
+data <- assay(UPS1.Case0)
+meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
+
+data <- list(x = as.matrix(data), y = rev(meta.info$Conc.), geneid = row.names(data), genenames = row.names(data), logged2 = TRUE)
+
+# Set up SAM input data
+samfit <- samr(data, resp.type="Two class unpaired", nperms = 1000 )
+
+delta.table <- samr.compute.delta.table(samfit, min.foldchange = 0)
+
+delta <- delta.table[delta.table[, "median FDR"] < 0.05, ][1, ][1] 
+siggenes.table <- samr.compute.siggenes.table(samfit, delta, data, delta.table, min.foldchange = 0 , all.genes = TRUE)
+
+both <- rbind(siggenes.table$genes.up, siggenes.table$genes.lo)
+dim(both)
+both <- data.frame(both)
+
+both$q.value... <- as.numeric(both$q.value...) / 100
+
+saveRDS(both,"paperdata/UPS1.Case0.samr.rds")
+```
+
 ### Comparison
@@
 
-limrots.result <- readRDS("../LimROTS paper/UPS1.Case0.LimROTS.rds")
-rots.result <- readRDS("../LimROTS paper/UPS1.Case0.ROTS.rds")
-limma.result <- readRDS("../LimROTS paper/UPS1.Case0.limma.rds")
-t.test.result <- readRDS("../LimROTS paper/UPS1.Case0.t.test.rds")
+data(UPS1.Case0)
+
+limrots.result <- readRDS("paperdata/UPS1.Case0.LimROTS.rds")
+rots.result <- readRDS("paperdata/UPS1.Case0.ROTS.rds")
+limma.result <- readRDS("paperdata/UPS1.Case0.limma.rds")
+t.test.result <- readRDS("paperdata/UPS1.Case0.t.test.rds")
+samr.result <- readRDS("paperdata/UPS1.Case0.samr.rds")
 
@@
 
+predictions_samr <- samr.result$q.value... < 0.05
+predictions_samr <- factor(predictions_samr, levels = c(TRUE, FALSE))
+true_labels_samr <- ifelse(samr.result$Gene.ID %in% TP , TRUE, FALSE)
+true_labels_samr <- factor(true_labels_samr, levels = c(TRUE, FALSE))
+conf_matrix_samr <- confusionMatrix(predictions_samr, true_labels_samr)
+
 ```
@@
 metrics_t.test <- extract_metrics(conf_matrix_t.test, "t.test")
+metrics_samr <- extract_metrics(conf_matrix_samr, "samr")
 
-all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma,metrics_t.test)
+all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma,metrics_t.test , metrics_samr)
 
@@
 
--   **limma** performs significantly worse than both **limROTS** and **ROTS**, especially in PPV and F1 score, making it less reliable for this particular dataset and analysis task.
+-   **limma, and samr** performs significantly worse than both **limROTS** and **ROTS**, especially in PPV and F1 score, making it less reliable for this particular dataset and analysis task.
 The lower sensitivity and specificity suggest that limma struggles with both true positive and true negative identification compared to the other two methods.
@@
 
+### Run samr
+
+```{r}
+library(samr)
+
+data("UPS1.Case1")
+data <- assay(UPS1.Case1)
+meta.info <- data.frame(colData(UPS1.Case1), check.names = FALSE, row.names = row.names(colData(UPS1.Case1)))
+
+data <- list(x = as.matrix(data), y = rev(meta.info$Conc.), geneid = row.names(data), genenames = row.names(data), logged2 = TRUE)
+
+# Set up SAM input data
+samfit <- samr(data, resp.type="Two class unpaired", nperms = 1000 )
+
+delta.table <- samr.compute.delta.table(samfit, min.foldchange = 0)
+
+delta <- delta.table[delta.table[, "median FDR"] < 0.05, ][1, ][1] 
+siggenes.table <- samr.compute.siggenes.table(samfit, delta, data, delta.table, min.foldchange = 0 , all.genes = TRUE)
+
+both <- rbind(siggenes.table$genes.up, siggenes.table$genes.lo)
+dim(both)
+both <- data.frame(both)
+
+both$q.value... <- as.numeric(both$q.value...) / 100
+
+saveRDS(both,"paperdata/UPS1.Case1.samr.rds")
+```
+
 ### Comparison
@@
 
-limrots.result <- readRDS("../LimROTS paper/UPS1.Case1.LimROTS.rds")
-rots.result <- readRDS("../LimROTS paper/UPS1.Case1.ROTS.rds")
-limma.result <- readRDS("../LimROTS paper/UPS1.Case1.limma.rds")
-ancova.result <- readRDS("../LimROTS paper/UPS1.Case1.ancova.rds")
+limrots.result <- readRDS("paperdata/UPS1.Case1.LimROTS.rds")
+rots.result <- readRDS("paperdata/UPS1.Case1.ROTS.rds")
+limma.result <- readRDS("paperdata/UPS1.Case1.limma.rds")
+ancova.result <- readRDS("paperdata/UPS1.Case1.ancova.rds")
+samar.result <- readRDS("paperdata/UPS1.Case1.samr.rds")
 
@@
 significant_ancova <- sapply(alpha_thresholds, function(alpha) sum(ancova.result$FDR < alpha))
+significant_samr <- sapply(alpha_thresholds, function(alpha) sum(samar.result$q.value... < alpha))
 
 plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
-                        significant_ROTS = significant_ROTS , significant_limma = significant_limma  ,significant_ancova = significant_ancova)
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma  ,significant_ancova = significant_ancova , significant_samr = significant_samr)
 
@@
 
+### Run samar
+
+```{r}
+library(samr)
+
+data("BM21.Case0")
+data <- assay(BM21.Case0)
+meta.info <- data.frame(colData(BM21.Case0), check.names = FALSE, row.names = row.names(colData(BM21.Case0)))
+
+data <- list(x = as.matrix(data), y = rev(meta.info$mixing.ratio), geneid = row.names(data), genenames = row.names(data), logged2 = TRUE)
+
+# Set up SAM input data
+samfit <- samr(data, resp.type="Two class unpaired", nperms = 1000 )
+
+delta.table <- samr.compute.delta.table(samfit, min.foldchange = 0)
+
+delta <- delta.table[delta.table[, "median FDR"] < 0.05, ][1, ][1] 
+siggenes.table <- samr.compute.siggenes.table(samfit, delta, data, delta.table, min.foldchange = 0 , all.genes = TRUE)
+
+both <- rbind(siggenes.table$genes.up, siggenes.table$genes.lo)
+dim(both)
+both <- data.frame(both)
+
+both$q.value... <- as.numeric(both$q.value...) / 100
+
+saveRDS(both,"paperdata/BM21.Case0.samr.rds")
+```
+
 ### Comparison
@@
 t.test.result <- readRDS("paperdata/BM21.Case0.t.test.rds")
+samr.result <- readRDS("paperdata/BM21.Case0.samr.rds")
 
@@
 significant_t.test <- sapply(alpha_thresholds, function(alpha) sum(t.test.result$fdr < alpha))
+significant_samr <- sapply(alpha_thresholds, function(alpha) sum(samr.result$q.value... < alpha))
 
-plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
+plot_data <- data.frame(Alpha = alpha_thresholds, significant_samr = significant_samr,
+                        significant_limROTS = significant_limROTS,
                         significant_ROTS = significant_ROTS , significant_limma = significant_limma, significant_t.test = significant_t.test  )
@@
 # Run LimROTS analysis with trend and robust settings enabled
-limrots.result <- LimROTS(data.exp = BM21.Case1, 
+rots.result <- LimROTS(data.exp = BM21.Case1, 
                           B = B, K = K, meta.info = meta.info, 
@@
 
-saveRDS(limrots.result , "paperdata/BM21.Case1.rots.rds")
+saveRDS(rots.result , "paperdata/BM21.Case1.rots.rds")
 ```
@@
 limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
+
+saveRDS(limma.result , "paperdata/BM21.Case1.limma.rds")
+```
+
+### Run ANOVA
+
+```{r}
+formula.str <- "~ 0 + mixing.ratio"
+data(BM21.Case1)
+
+data <- assay(BM21.Case1)
+meta.info <- data.frame(colData(BM21.Case1), check.names = FALSE, row.names = row.names(colData(BM21.Case1)))
+meta.info$sample.id <- NULL
+ancova.result <- data.frame()
+for(i in 1:nrow(data)){
+  df.temp <- data[i,]
+
+  data.temp <- data.frame(response = as.numeric(df.temp) , mixing.ratio = meta.info$mixing.ratio)
+
+  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
+  coef_table <- anova(ancova_model)
+  coef_table <- data.frame(coef_table)
+  significant_groups <- coef_table[rownames(coef_table) == "mixing.ratio" , ]
+  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
+}
+ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
+saveRDS(ancova.result, "paperdata/BM21.Case1.anova.rds")
+```
+
+Run samr
+
+```{r}
+library(samr)
+
+data("BM21.Case1")
+data <- assay(BM21.Case1)
+meta.info <- data.frame(colData(BM21.Case1), check.names = FALSE, row.names = row.names(colData(BM21.Case1)))
+
+data <- list(x = as.matrix(data), y = rev(meta.info$mixing.ratio), geneid = row.names(data), genenames = row.names(data), logged2 = TRUE)
+
+# Set up SAM input data
+samfit <- samr(data, resp.type="Multiclass", nperms = 1000 )
+
+delta.table <- samr.compute.delta.table(samfit, min.foldchange = 0)
+
+delta <- delta.table[delta.table[, "median FDR"] < 0.05, ][1, ][1] 
+siggenes.table <- samr.compute.siggenes.table(samfit, delta, data, delta.table, min.foldchange = 0 , all.genes = TRUE)
+
+both <- rbind(siggenes.table$genes.up, siggenes.table$genes.lo)
+dim(both)
+both <- data.frame(both)
+
+both$q.value... <- as.numeric(both$q.value...) / 100
+
+saveRDS(both,"paperdata/BM21.Case1.samr.rds")
+```
+
+### Comparison
+
+```{r}
+library(tidyverse, quietly = T, warn.conflicts = T)
+library(ggsci)
+
+limrots.result <- readRDS("paperdata/BM21.Case1.limrots.rds")
+rots.result <- readRDS("paperdata/BM21.Case1.rots.rds")
+limma.result <- readRDS("paperdata/BM21.Case1.limma.rds")
+anova.result <- readRDS("paperdata/BM21.Case1.anova.rds")
+samr.result <- readRDS("paperdata/BM21.Case1.samr.rds")
+
+alpha_thresholds <- seq(0, 0.075, by = 0.01)
+significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$q_values$qvalues < alpha))
+significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
+significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
+significant_anova <- sapply(alpha_thresholds, function(alpha) sum(anova.result$FDR < alpha))
+significant_samr <- sapply(alpha_thresholds, function(alpha) sum(samr.result$q.value... < alpha))
+
+plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
+                        significant_ROTS = significant_ROTS , significant_limma = significant_limma, significant_anova = significant_anova, significant_samr = significant_samr)
+
+plot_data.melt <-  plot_data %>%
+  pivot_longer(cols = starts_with("significant"), names_to = "Method", values_to = "FDR")
+
+plot_data.melt$size <- ifelse(plot_data.melt$Method == "significant_anova", 2.5,1.2)
+
+# Plot using ggplot2
+ggplot(plot_data.melt, aes(x = Alpha, y = FDR, color = Method , size =size)) +
+  geom_line() +
+  theme_minimal() + scale_size_identity()+
+  labs(title = "Significant Proteins vs FDR BM21 Case study 4",
+       x = "FDR",
+       y = "Number of Significant Proteins") +
+   geom_vline(xintercept = 0.01, linetype="dotted", 
+                color = "red", size=1) + 
+    geom_vline(xintercept = 0.05, linetype="dotted", 
+                color = "red", size=1)+
+    scale_color_npg() 
+    theme(
+    plot.title = element_text(face = "bold", color = "black"),
+    axis.title = element_text(face = "bold", color = "black"),
+    axis.text = element_text(face = "bold", color = "black"),
+    axis.ticks = element_line(color = "black")
+  )
 ```
@@
 # Load the dataset
-load("../LimROTS paper/metabolomics/BM21.Case3.rda")
-
+data(BM21.Case3)
 print(BM21.Case3)
@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case3)[1]/4  # Set the value for K based on the data size
+K <- dim(BM21.Case3)[1]/2  # Set the value for K based on the data size
 K <- floor(K) 
@@
 
+length(which(limrots.result.df$q.value < 0.01))
+
+limrots.result.df <- limrots.result.df[limrots.result.df$q.value < 0.01,]
+
+length(which(grepl("_pluseffect", row.names(limrots.result.df)) ))
+
+length(which(grepl("synthetic", row.names(limrots.result.df)) ))
 ```
@@
 
-length(which(limma.result$adj.P.Val < 0.05))
+length(which(limma.result$adj.P.Val < 0.01))
 
-limma.result <- limma.result[limma.result$adj.P.Val < 0.05,]
+limma.result <- limma.result[limma.result$adj.P.Val < 0.01,]
 

### 2024-09-20 R/BM21.Case0.r
.

@@ -1,0 +1,28 @@
+#' BM21 Dataset case 0
+#'
+#' A `SummarizedExperiment` object with a dataset was obtained from the Metabolomics Workbench (ST002454). Different matrix ratios of human plasma and vegetable juice were analyzed, This dataset contains six different mixing ratios: 1:1024, 1:256, 1:64, and their reverse, each with three technical replicates. The samples were grouped into high plasma ratio samples and low ratio samples.
+#'
+#' @name BM21.Case0
+#' @docType data
+#' @format An instance of the `SummarizedExperiment` class with the following assays:
+#' \describe{
+#'   \item{log2}{This assay includes log2 metabolites peaks}
+#' }
+#' The object also contains colData and rowData:
+#' \describe{
+#'   \item{colData}{A DataFrame with metadata for samples.}
+#'   \item{rowData}{A DataFrame with metadata for metabolites peaks}
+#' }
+#'
+#' The `colData` contains the following columns:
+#' \describe{
+#'   \item{SampleID}{Unique identifier for each sample.}
+#'   \item{mixing.ratio}{Experimental condition or group for each sample , representing different different mixing ratios of human plasma and vegetable juice .}
+#' }
+#'
+#' @source Bloody Mary (BM21)- Serial mixtures of vegetable juice/water and human plasma, Metabolomics Workbench (ST002454).
+#' @references
+#' This data is available at the NIH Common Fund's National Metabolomics Data Repository (NMDR) website, the Metabolomics Workbench, https://www.metabolomicsworkbench.org, where it has been assigned Project ID PR001582. The data can be accessed directly via it's Project DOI: 10.21228/M86Q7N This work is supported by NIH grant, U2C- DK119886.
+#'
+
+NULL

### 2024-09-20 tomakeapackae.r

@@ -13,6 +13,8 @@ Rcpp::compileAttributes()
 library(devtools)
 #test()
 
+roxygen2::roxygenize()
+
 devtools::check(vignettes = FALSE)
 
 devtools::check(vignettes = TRUE)

### 2024-09-20 paper.qmd
Update paper.qmd

@@
 ```
-
-## BM21 Case study 5
-
-### Run LimROTS
-
-```{r}
-# Load necessary packages
-library(LimROTS)
-library(parallel)
-
-# Load the dataset
-data(BM21.Case3)
-print(BM21.Case3)
-
-# Set random seed for reproducibility
-set.seed(1234, kind = "default")
-
-# Set metadata and formula for LimROTS analysis
-meta.info <- c("mixing.ratio" , "fake.batch")
-B <- 1000  # Number of bootstrap samples
-K <- dim(BM21.Case3)[1]/2  # Set the value for K based on the data size
-K <- floor(K) 
-num_cores <- 30   # Number of cores for parallel processing
-cluster <- makeCluster(num_cores)  # Create a parallel cluster
-group.name <- "mixing.ratio"
-formula.str <- "~ 0 + mixing.ratio + fake.batch"  # Formula for group comparison
-
-# Run LimROTS analysis with trend and robust settings enabled
-limrots.result <- LimROTS(data.exp = BM21.Case3, 
-                          B = B, K = K, meta.info = meta.info, 
-                          cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE)
-
-summary(limrots.result$q_values)
-plot(limrots.result$q_values)
-hist(limrots.result$q_values)
-
-limrots.result.df <- data.frame(row.names = row.names(limrots.result$data), q.value = limrots.result$q_values$qvalues , fdr = limrots.result$FDR,
-                                bh = limrots.result$BH.pvalue , fc = limrots.result$corrected.logfc)
-
-length(which(limrots.result.df$q.value < 0.01))
-
-limrots.result.df <- limrots.result.df[limrots.result.df$q.value < 0.01,]
-
-length(which(grepl("_pluseffect", row.names(limrots.result.df)) ))
-
-length(which(grepl("synthetic", row.names(limrots.result.df)) ))
-```
-
-### Run limma
-
-```{r}
-
-library(limma)
-library(SummarizedExperiment, quietly = T, warn.conflicts = T)
-
-load("../LimROTS paper/metabolomics/BM21.Case3.rda")
-
-print("BM21.Case3")
-data <- assay(BM21.Case3)
-formula.str <- "~ 0 + mixing.ratio + fake.batch"
-meta.info <- data.frame(colData(BM21.Case3), check.names = FALSE, row.names = row.names(colData(BM21.Case3)))
-meta.info$sample.id <- NULL
-design.matrix <- model.matrix(formula(formula.str), data = meta.info)
-colnames(design.matrix) <- make.names(colnames(design.matrix))
-fit <- lmFit(data, design.matrix)
-cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
-fit2 <- contrasts.fit(fit, cont_matrix)
-fit.ebayes <- eBayes(fit2, trend=T, robust=T)
-limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
-
-length(which(limma.result$adj.P.Val < 0.01))
-
-limma.result <- limma.result[limma.result$adj.P.Val < 0.01,]
-
-length(which(grepl("_pluseffect", row.names(limma.result)) ))
-
-length(which(grepl("synthetic", row.names(limma.result)) ))
-
-```

### 2024-09-22 paper.qmd
.

@@
 ---
-
-##
 
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE)
+
 saveRDS(limrots.result, "../LimROTS paper/UPS1.Case0.LimROTS.rds")
@@
                           cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE)
 saveRDS(limrots.result, "../LimROTS paper/UPS1.Case1.LimROTS.rds")

### 2024-09-22 R/calculateFalseDiscoveryRate.R

@@ -1,5 +1,4 @@
-
-#' Calculate False Discovery Rate (FDR) Using Permuted Values
+#' Calculate False Discovery Rate (FDR) Using Permuted Values (Optimized)
 #'
 #' This function calculates the false discovery rate (FDR) by comparing observed values to permuted values.
 #' The function sorts observed values, compares them against permuted data, and computes FDR using the median of permutation results.
@@ -8,19 +7,11 @@
 #' @param permutedValues Numeric matrix. The permuted test statistics or values, with rows corresponding to the same values as in `observedValues` and columns representing different permutations.
 #' @param showProgress Logical. If `TRUE`, a progress bar will be shown during the computation.
 #'
-#' @details
-#' This function computes the FDR by comparing the magnitude of observed values against permuted values in a sorted order.
-#' The FDR is calculated as the median ratio of the number of permuted values greater than or equal to the observed values.
-#' The results are adjusted to ensure that FDR values are non-increasing, as required for multiple testing procedures.
-#'
 #' @return A numeric vector of the same length as `observedValues`, containing the estimated FDR for each observed value.
 #'
-#' @importFrom stats median
-#'
 #' @export
-#'
 
-calculateFalseDiscoveryRate <- function(observedValues, permutedValues, showProgress) {
+calculateFalseDiscoveryRate <- function(observedValues, permutedValues, showProgress = FALSE) {
   observedAbs <- abs(observedValues)
   permutedAbs <- abs(permutedValues)
 
@@ -36,8 +27,10 @@ calculateFalseDiscoveryRate <- function(observedValues, permutedValues, showProg
 
   for (i in seq_len(numPermutations)) {
     a.rand <- sort(permutedAbs[, i], decreasing = TRUE, na.last = TRUE)
-    n.bigger <- countLargerThan(a, a.rand)  # Use the new function
-    FDRmatrix[ord, i] <- n.bigger / seq_along(a)
+
+    n.bigger <- countLargerThan(a, a.rand)
+
+    FDRmatrix[, i] <- n.bigger / seq_along(a)
 
     if (showProgress) {
       setTxtProgressBar(progressBar, i)
@@ -50,16 +43,15 @@ calculateFalseDiscoveryRate <- function(observedValues, permutedValues, showProg
 
[…]

### 2024-09-22 paper.qmd
.

@@
 
-saveRDS(limrots.result, "../LimROTS paper/UPS1.Case0.LimROTS.rds")
+saveRDS(limrots.result, "paperdata/UPS1.Case0.LimROTS.rds")
 ```
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
-saveRDS(rots.result, "../LimROTS paper/UPS1.Case0.ROTS.rds")
+saveRDS(rots.result, "paperdata/UPS1.Case0.ROTS.rds")
 ```
@@
 limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
-saveRDS(limma.result, "../LimROTS paper/UPS1.Case0.limma.rds")
+saveRDS(limma.result, "paperdata/UPS1.Case0.limma.rds")
 ```
@@
 K <- floor(K) 
-num_cores <- 50  # Number of cores for parallel processing
+num_cores <- 10  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
                           cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE)
-saveRDS(limrots.result, "../LimROTS paper/UPS1.Case1.LimROTS.rds")
+                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+saveRDS(limrots.result, "paperdata/UPS1.Case1.LimROTS.rds")
 ```
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
-saveRDS(rots.result, "../LimROTS paper/UPS1.Case1.ROTS.rds")
+saveRDS(rots.result, "paperdata/UPS1.Case1.ROTS.rds")
 ```
@@
 limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
-saveRDS(limma.result, "../LimROTS paper/UPS1.Case1.limma.rds")
+saveRDS(limma.result, "paperdata/UPS1.Case1.limma.rds")
 ```
@@
 ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
-saveRDS(ancova.result, "../LimROTS paper/UPS1.Case1.ancova.rds")
+saveRDS(ancova.result, "paperdatar/UPS1.Case1.ancova.rds")
 ```
@@
 K <- floor(K) 
-num_cores <- 30  # Number of cores for parallel processing
+num_cores <- 10  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
                           cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE)
-saveRDS(limrots.result, "../LimROTS paper/UPS1.Case2.LimROTS.rds")
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE)
+saveRDS(limrots.result, "paperdata/UPS1.Case2.LimROTS.rds")
 ```
@@
 limma.result <- topTable(fit.ebayes, coef="Conc.1-Conc.2" , number = "Inf")
-saveRDS(limma.result, "../LimROTS paper/UPS1.Case2.limma.rds")
+saveRDS(limma.result, "paperdata/UPS1.Case2.limma.rds")
 ```
@@
 ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
-saveRDS(ancova.result, "../LimROTS paper/UPS1.Case2.ancova.rds")
+saveRDS(ancova.result, "paperdata/UPS1.Case2.ancova.rds")
 
@@
                           cluster = cluster, group.name = group.name, 
-                          formula.str = formula.str, trend = TRUE, robust = TRUE)
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE)
 
@@
 
-Run samr
+### Run samr
 

### 2024-09-23 R/Bootstrap_functions.r
.

@@ -1,5 +1,3 @@
-
-
 #' Generate Bootstrap Samples with Optional Pairing
 #'
 #' This function generates bootstrap samples from the input metadata. It samples with replacement
@@ -28,23 +26,25 @@
 #' bootstrapS(B = 10, meta.info = meta.info, group.name = "group", paired = TRUE)
 
 
+
+
 bootstrapS <- function (B, meta.info, group.name , paired) {
-  groups <- meta.info[, group.name]
-  bootsamples <- matrix(nrow = B, ncol = length(groups))
-  for (i in seq_len(B)) {
-    for (g in unique(groups)) {
-      g.names <- row.names(meta.info)[which(groups == g)]
-      bootsamples[i, which(groups == g)] <- sample(g.names, length(g.names), replace = TRUE)
-    }
-  }
-  if (paired) {
+    groups <- meta.info[, group.name]
+    bootsamples <- matrix(nrow = B, ncol = length(groups))
     for (i in seq_len(B)) {
-      g.names1 <-  bootsamples[i, which(groups == unique(groups)[1])]
-      g.names2 <- match(g.names1 , row.names(meta.info)) + length(g.names1)
-      bootsamples[i, which(groups == unique(groups)[2])] <- row.names(meta.info)[g.names2]
+        for (g in unique(groups)) {
+            g.names <- row.names(meta.info)[which(groups == g)]
+            bootsamples[i, which(groups == g)] <- sample(g.names, length(g.names), replace = TRUE)
+        }
     }
-  }
-  return(bootsamples)
+    if (paired) {
+        for (i in seq_len(B)) {
+            g.names1 <-  bootsamples[i, which(groups == unique(groups)[1])]
+            g.names2 <- match(g.names1 , row.names(meta.info)) + length(g.names1)
+            bootsamples[i, which(groups == unique(groups)[2])] <- row.names(meta.info)[g.names2]
+        }
+    }
+    return(bootsamples)
 }
 
 
@@ -69,13 +69,14 @@ bootstrapS <- function (B, meta.info, group.name , paired) {
 #' meta.info <- data.frame(group = rep(c("A", "B"), each = 5), row.names = paste0("Sample", 1:10))
 #' permutatedS(meta.info = meta.info, B = 10)
 
+
 permutatedS <- function (meta.info, B)
 {
-  persamples <- matrix(nrow = B, ncol = nrow(meta.info))
-  for (i in seq_len(B)) {
[…]

### 2024-09-23 R/LimROTS.r

@@ -61,6 +61,7 @@
 #' @import SummarizedExperiment
 #' @export
 
+
 LimROTS <- function (data.exp,
                      B = 1000,
                      K = NULL,
@@ -80,349 +81,370 @@ LimROTS <- function (data.exp,
                      robust = TRUE,
                      trend = TRUE)
 {
-  SanityChecK.list <- SanityChecK(
-    data.exp,
-    B = B,
-    K = K,
-    a1 = a1,
-    a2 = a2,
-    meta.info = meta.info,
-    group.name = group.name ,
-    formula.str = formula.str,
-    survival = survival,
-    paired = paired,
-    n.ROTS = n.ROTS,
-    verbose = verbose,
-    log = log
-  )
+    SanityChecK.list <- SanityChecK(
+        data.exp,
+        B = B,
+        K = K,
+        a1 = a1,
+        a2 = a2,
+        meta.info = meta.info,
+        group.name = group.name ,
+        formula.str = formula.str,
+        survival = survival,
+        paired = paired,
+        n.ROTS = n.ROTS,
+        verbose = verbose,
+        log = log
+    )
 
 
-  meta.info <-  SanityChecK.list$meta.info
-  data <-  SanityChecK.list$data
-  groups <-  SanityChecK.list$groups
-  event <-  SanityChecK.list$event
-  K <-  SanityChecK.list$K
+    meta.info <-  SanityChecK.list$meta.info
+    data <-  SanityChecK.list$data
+    groups <-  SanityChecK.list$groups
+    event <-  SanityChecK.list$event
+    K <-  SanityChecK.list$K
 
-  #### FC
+    #### FC
 
-  if (length(unique(groups)) == 2 && survival == FALSE) {
-    group1_data <- data[, groups == 1]
-    group2_data <- data[, groups == 2]
+    if (length(unique(groups)) == 2 && survival == FALSE) {
+        group1_data <- data[, groups == 1]
+        group2_data <- data[, groups == 2]
 
-    if (log) {
-      logfc <- rowMeans(group1_data, na.rm = TRUE) - rowMeans(group2_data, na.rm = TRUE)
+        if (log) {
+            logfc <- rowMeans(group1_data, na.rm = TRUE) - rowMeans(group2_data, na.rm = TRUE)
+        } else {
[…]

### 2024-09-23 R/Optimizing.R

@@ -32,95 +32,96 @@
 #'
 
 
+
 Optimizing <- function(B, ssq, N, D, S, pD, pS, verbose, progress) {
-  if (verbose)
-    message("Optimizing parameters")
-  reprotable <- matrix(nrow = length(ssq) + 1, ncol = length(N))
-  colnames(reprotable) <- N
-  row.names(reprotable) <- c(ssq, "slr")
-  reprotable.P <- matrix(nrow = length(ssq) + 1, ncol = length(N))
-  colnames(reprotable.P) <- N
-  row.names(reprotable.P) <- c(ssq, "slr")
-  reprotable.sd <- matrix(nrow = length(ssq) + 1, ncol = length(N))
-  colnames(reprotable.sd) <- N
-  row.names(reprotable.sd) <- c(ssq, "slr")
-  if (progress)
-    pb <- txtProgressBar(min = 0,
-                         max = length(ssq),
-                         style = 3)
-  for (i in seq_len(length(ssq))) {
+    if (verbose)
+        message("Optimizing parameters")
+    reprotable <- matrix(nrow = length(ssq) + 1, ncol = length(N))
+    colnames(reprotable) <- N
+    row.names(reprotable) <- c(ssq, "slr")
+    reprotable.P <- matrix(nrow = length(ssq) + 1, ncol = length(N))
+    colnames(reprotable.P) <- N
+    row.names(reprotable.P) <- c(ssq, "slr")
+    reprotable.sd <- matrix(nrow = length(ssq) + 1, ncol = length(N))
+    colnames(reprotable.sd) <- N
+    row.names(reprotable.sd) <- c(ssq, "slr")
+    if (progress)
+        pb <- txtProgressBar(min = 0,
+                             max = length(ssq),
+                             style = 3)
+    for (i in seq_len(length(ssq))) {
+        overlaps <- matrix(0, nrow = B, ncol = length(N))
+        overlaps.P <- matrix(0, nrow = B, ncol = length(N))
+        cResults <- calOverlaps(
+            D,
+            S,
+            pD,
+            pS,
+            nrow(D),
+            as.integer(N),
+            length(N),
+            ssq[i],
+            as.integer(B),
+            overlaps,
+            overlaps.P
+        )
+        reprotable[i, ] <- colMeans(cResults[["overlaps"]])
+        reprotable.P[i, ] <- colMeans(cResults[["overlaps_P"]])
+        reprotable.sd[i, ] <- sqrt(rowSums((t(cResults[["overlaps"]]) -
[…]

### 2024-09-23 R/SanityChecK.R

@@ -30,6 +30,7 @@
 #'
 
 
+
 SanityChecK <- function(data.exp,
                         B = 1000,
                         K = NULL,
@@ -43,133 +44,143 @@ SanityChecK <- function(data.exp,
                         n.ROTS = FALSE,
                         verbose = TRUE,
                         log = TRUE) {
-
-
-  if(survival == TRUE){
-    if(n.ROTS == FALSE){
-      stop("survival is TRUE, Survival analysis is only available through the old ROTS implementation. To enable this, please set n.ROTS to TRUE.")
+    if (survival == TRUE) {
+        if (n.ROTS == FALSE) {
+            stop(
+                "survival is TRUE, Survival analysis is only available through the old ROTS implementation. To enable this, please set n.ROTS to TRUE."
+            )
+        }
     }
-  }
 
-  if(paired == TRUE){
-    if(n.ROTS == FALSE){
-      stop("paired if TRUE, Paired analysis is only available through the old ROTS implementation. To enable this, please set n.ROTS to TRUE.")
+    if (paired == TRUE) {
+        if (n.ROTS == FALSE) {
+            stop(
+                "paired if TRUE, Paired analysis is only available through the old ROTS implementation. To enable this, please set n.ROTS to TRUE."
+            )
+        }
     }
-  }
 
-  if(log == FALSE){
-    if(n.ROTS == FALSE){
-      stop("log is FALSE, Unlogged values can only be handled through the old ROTS implementation. To enable this, please set n.ROTS to TRUE.")
+    if (log == FALSE) {
+        if (n.ROTS == FALSE) {
+            stop(
+                "log is FALSE, Unlogged values can only be handled through the old ROTS implementation. To enable this, please set n.ROTS to TRUE."
+            )
 
+        }
     }
-  }
 
-  ### SummarizedExperiment
-  if (inherits(data.exp, "SummarizedExperiment")) {
-    message("Data is SummarizedExperiment object")
+    ### SummarizedExperiment
+    if (inherits(data.exp, "SummarizedExperiment")) {
+        message("Data is SummarizedExperiment object")
 
-    if (is.null(meta.info)) {
-      stop("meta.info should be a vector of colData names to be used")
+        if (is.null(meta.info)) {
+            stop("meta.info should be a vector of colData names to be used")
+        } else{
+            meta.info.colnames <- meta.info
+            meta.info <- data.frame(
+                colData(data.exp)[, meta.info],
+                check.names = FALSE,
[…]

### 2024-09-23 R/calculateFalseDiscoveryRate.R

@@ -23,37 +23,40 @@
 #'
 #'
 
+
 calculateFalseDiscoveryRate <- function(observedValues,
                                         permutedValues,
                                         showProgress = FALSE) {
-  observedAbs <- abs(observedValues)
-  permutedAbs <- abs(permutedValues)
-  ord <- order(observedAbs, decreasing = TRUE, na.last = TRUE)
-  a <- observedAbs[ord]
-  numPermutations <- ncol(permutedValues)
-  FDRmatrix <- matrix(NA, nrow = length(a), ncol = numPermutations)
-  if (showProgress) {
-    progressBar <- txtProgressBar(min = 0,
-                                  max = numPermutations,
-                                  style = 3)
-  }
-  for (i in seq_len(numPermutations)) {
-    a.rand <- sort(permutedAbs[, i], decreasing = TRUE, na.last = TRUE)
-    n.bigger <- countLargerThan(a, a.rand)
-    FDRmatrix[, i] <- n.bigger / seq_along(a)
+    observedAbs <- abs(observedValues)
+    permutedAbs <- abs(permutedValues)
+    ord <- order(observedAbs, decreasing = TRUE, na.last = TRUE)
+    a <- observedAbs[ord]
+    numPermutations <- ncol(permutedValues)
+    FDRmatrix <- matrix(NA, nrow = length(a), ncol = numPermutations)
     if (showProgress) {
-      setTxtProgressBar(progressBar, i)
+        progressBar <- txtProgressBar(min = 0,
+                                      max = numPermutations,
+                                      style = 3)
     }
-  }
-  if (showProgress) {
-    close(progressBar)
-  }
-  falseDiscoveryRate <- apply(FDRmatrix, 1, median)
-  falseDiscoveryRate[falseDiscoveryRate > 1] <- 1
-  for (i in length(falseDiscoveryRate):1) {
-    falseDiscoveryRate[i] <- min(falseDiscoveryRate[i:length(falseDiscoveryRate)])
-  }
-  return(falseDiscoveryRate)
+    for (i in seq_len(numPermutations)) {
+        a.rand <- sort(permutedAbs[, i],
+                       decreasing = TRUE,
+                       na.last = TRUE)
+        n.bigger <- countLargerThan(a, a.rand)
+        FDRmatrix[, i] <- n.bigger / seq_along(a)
+        if (showProgress) {
[…]

### 2024-09-23 R/testStatOptimized.R

@@ -19,55 +19,56 @@
 
 
 
+
 testStatOptimized <- function(isPaired, sampleGroups) {
-  if (length(sampleGroups) == 2) {
-    groupA <- sampleGroups[[1]]
-    groupB <- sampleGroups[[2]]
-    meanA <- rowMeans(groupA, na.rm = TRUE)
-    meanB <- rowMeans(groupB, na.rm = TRUE)
-    sumSqA <- rowSums((groupA - meanA) ^ 2, na.rm = TRUE)
-    sumSqB <- rowSums((groupB - meanB) ^ 2, na.rm = TRUE)
-    if (!isPaired) {
-      nonNAcountA <- rowSums(!is.na(groupA))
-      nonNAcountB <- rowSums(!is.na(groupB))
-      meanDiff <- meanB - meanA
-      pooledSD <- sqrt(((sumSqA + sumSqB) / (nonNAcountA + nonNAcountB - 2)) *
-                         (1 / nonNAcountA + 1 / nonNAcountB))
-      insufficientSamples <- which(nonNAcountA < 2 |
-                                     nonNAcountB < 2)
-      meanDiff[insufficientSamples] <- 0
-      pooledSD[insufficientSamples] <- 1
-      return(list(d = meanDiff, s = pooledSD))
-    } else {
-      covarianceAB <- rowSums((groupA - meanA) * (groupB - meanB), na.rm = TRUE)
-      pairedCount <- rowSums(!is.na(groupA * groupB))
-      meanDiff <- meanB - meanA
-      pairedSD <- sqrt(((sumSqA + sumSqB) / (2 * pairedCount - 2)) * (2 / pairedCount) -
-                         2 / (pairedCount * (pairedCount - 1)) * covarianceAB)
-      insufficientSamples <- which(pairedCount < 2)
-      meanDiff[insufficientSamples] <- 0
-      pairedSD[insufficientSamples] <- 1
+    if (length(sampleGroups) == 2) {
+        groupA <- sampleGroups[[1]]
+        groupB <- sampleGroups[[2]]
+        meanA <- rowMeans(groupA, na.rm = TRUE)
+        meanB <- rowMeans(groupB, na.rm = TRUE)
+        sumSqA <- rowSums((groupA - meanA) ^ 2, na.rm = TRUE)
+        sumSqB <- rowSums((groupB - meanB) ^ 2, na.rm = TRUE)
+        if (!isPaired) {
[…]

### 2024-09-24 inst/extdata/paperdata/paper.qmd
.

--- 2024-09-22 paper.qmd
@@
 ```
+
+## Columbia Case Study
+
+### Run LimROTS
+
+```{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+# Load the dataset
+#data("Columbia.RealCase")
+print(Columbia.RealCase)
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- c("groups" , "Age" , "gender" , "LRRK2.status" , "GBA.status")
+B <- 1000  # Number of bootstrap samples
+K <- dim(Columbia.RealCase)[1]/2  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "groups"
+formula.str <- "~ 0 + groups + Age + gender + LRRK2.status + GBA.status"  # Formula for group comparison
+# Run LimROTS analysis with trend and robust settings enabled
+
+trace(LimROTS, edit = T)
+
+limrots.result <- LimROTS(data.exp = Columbia.RealCase, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE, seed.cl = 1234)
+
+saveRDS(limrots.result, "inst/extdata/paperdata/Columbia.RealCase.LimROTS.rds")
+```
+
+### Q Plot
+
+```{r}
+plot(limrots.result$q_values)
+hist(limrots.result$q_values)
+summary(limrots.result$q_values)
+```
+
+### Volcano Plot with ggplot2
+
+```{r}
+library(ggplot2)
+# Create a data frame from the LimROTS results
+limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
+                                LimROTS.FC = limrots.result$logfc,
+                                q.value = limrots.result$q_values$pvalues)
+
+# Create a volcano plot
+ggplot(limrots.result.df, aes(x = limrots.result$corrected.logfc, y = -log10(limrots.result$pvalue) )) +
+  geom_point(alpha = 0.8) +
+  theme_bw() +
+  labs(title = "Volcano Plot", x = "Log Fold Change", y = "-Log10 q.value", color = "True Positive") +
+  scale_color_manual(values = c("grey", "red"))+
+  geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "blue")+
+  geom_hline(yintercept=-log10(0.01), linetype="dashed", color = "black")
+```
+
+```{r}
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+data("Columbia.RealCase")
+data <- assay(Columbia.RealCase)
+formula.str <- "~ 0 + groups + Age + gender + LRRK2.status + GBA.status"
+meta.info <- data.frame(colData(Columbia.RealCase), check.names = FALSE, row.names = row.names(colData(Columbia.RealCase)))
+meta.info$sample.id <- NULL
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("groups1-groups2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+limma.result <- topTable(fit.ebayes, coef="groups1-groups2" , number = "Inf")
+
+# Create a volcano plot
+ggplot(limma.result, aes(x = logFC, y = -log10(P.Value))) +
+  geom_point(alpha = 0.8) +
+  theme_bw() +
+  labs(title = "Volcano Plot", x = "Log Fold Change", y = "-Log10 q.value", color = "True Positive") +
+  scale_color_manual(values = c("grey", "red"))+
+  geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "blue")+
+  geom_hline(yintercept=-log10(0.01), linetype="dashed", color = "black")
+```

### 2024-09-24 inst/extdata/paperdata/paper.qmd
.

@@
 # Load the dataset
-#data("Columbia.RealCase")
+data("Columbia.RealCase")
 print(Columbia.RealCase)
+
+set.seed(1234, kind = "default")
+
+data.assay <- assay(Columbia.RealCase)
+data.assay <- preprocessCore::normalize.quantiles(as.matrix(data.assay))
 
@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(Columbia.RealCase)[1]/2  # Set the value for K based on the data size
+K <- dim(Columbia.RealCase)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
-num_cores <- 30  # Number of cores for parallel processing
+num_cores <- 10  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
@@
 hist(limrots.result$q_values)
+hist(limrots.result$q_values$qvalues)
+
 summary(limrots.result$q_values)
@@
 # Create a volcano plot
-ggplot(limrots.result.df, aes(x = limrots.result$corrected.logfc, y = -log10(limrots.result$pvalue) )) +
+ggplot(limrots.result.df, aes(x = limrots.result$corrected.logfc, y = -log10(limrots.result$q_values$qvalues) )) +
   geom_point(alpha = 0.8) +
@@
 
+volcanoplot(fit.ebayes)
+plotMD(fit.ebayes)
+
 # Create a volcano plot
@@
 ```
+
+```{r}
+library(ROTS)
+
+rots <- ROTS(data, groups = as.numeric(meta.info$groups), B = 1000, progress = TRUE,
+             K = 600, a1= 5, a2 = 1)
+```
+
+```{r}
+rots.df <- data.frame(p = rots$pvalue, f = rots$logfc)
+
+ggplot(rots.df, aes(x = f, y = -log10(p))) +
+  geom_point(alpha = 0.8) +
+  theme_bw() +
+  labs(title = "Volcano Plot", x = "Log Fold Change", y = "-Log10 q.value", color = "True Positive") +
+  scale_color_manual(values = c("grey", "red"))+
+  geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "blue")+
+  geom_hline(yintercept=-log10(0.01), linetype="dashed", color = "black")
+```

### 2024-09-25 inst/extdata/paperdata/paper.qmd
proteogenomics

@@
 # Load the dataset
-data("Columbia.RealCase")
-print(Columbia.RealCase)
+load("D:\\MJFF\\MJFF V2 10.30.2023\\proteogenomic study\\soma.rda")
+print(soma)
 
@@
 
-data.assay <- assay(Columbia.RealCase)
-data.assay <- preprocessCore::normalize.quantiles(as.matrix(data.assay))
-
 # Set metadata and formula for LimROTS analysis
-meta.info <- c("groups" , "Age" , "gender" , "LRRK2.status" , "GBA.status")
+meta.info <- c("group_comp" , "gender" , "AGE") 
 B <- 1000  # Number of bootstrap samples
-K <- dim(Columbia.RealCase)[1]/4  # Set the value for K based on the data size
+K <- dim(soma)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
@@
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
-group.name <- "groups"
-formula.str <- "~ 0 + groups + Age + gender + LRRK2.status + GBA.status"  # Formula for group comparison
+group.name <- "group_comp"
+formula.str <- "~ 0 + group_comp + gender + AGE"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled
 
-trace(LimROTS, edit = T)
-
-limrots.result <- LimROTS(data.exp = Columbia.RealCase, 
+limrots.result <- LimROTS(data.exp = soma, 
                           B = B, K = K, meta.info = meta.info, 
@@
 
-saveRDS(limrots.result, "inst/extdata/paperdata/Columbia.RealCase.LimROTS.rds")
+saveRDS(limrots.result, "inst/extdata/paperdata/soma.LimROTS.rds")
 ```
@@
 limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
-                                LimROTS.FC = limrots.result$logfc,
-                                q.value = limrots.result$q_values$pvalues)
+                                LimROTS.FC = limrots.result$corrected.logfc,
+                                p.value = limrots.result$q_values$pvalues)
 
 # Create a volcano plot
-ggplot(limrots.result.df, aes(x = limrots.result$corrected.logfc, y = -log10(limrots.result$q_values$qvalues) )) +
+ggplot(limrots.result.df, aes(x = LimROTS.FC, y = -log10(p.value) )) +
   geom_point(alpha = 0.8) +
@@
   geom_hline(yintercept=-log10(0.01), linetype="dashed", color = "black") 
+```
+
+```{r}
+library(SomaScan.db)
+soma@NAMES
+
+select(SomaScan.db, keys = "9073-35", columns = c("SYMBOL", "UNIPROT"))
+
 ```

### 2024-09-25 inst/extdata/paperdata/paper.qmd
.

@@
 
-## Columbia Case Study
+## Proteogenomic PD Case Study
 
@@
 # Load the dataset
-load("D:\\MJFF\\MJFF V2 10.30.2023\\proteogenomic study\\soma.rda")
+data("proteogenomicPD")
 print(soma)
@@
 # Set metadata and formula for LimROTS analysis
-meta.info <- c("group_comp" , "gender" , "AGE") 
+meta.info <- names(soma@colData@listData)
 B <- 1000  # Number of bootstrap samples
-K <- dim(soma)[1]/4  # Set the value for K based on the data size
+K <- dim(soma)[1]/2  # Set the value for K based on the data size
 K <- floor(K) 
-num_cores <- 10  # Number of cores for parallel processing
+num_cores <- 30  # Number of cores for parallel processing
 cluster <- makeCluster(num_cores)  # Create a parallel cluster
 group.name <- "group_comp"
-formula.str <- "~ 0 + group_comp + gender + AGE"  # Formula for group comparison
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled
+
+soma@colData@listData[["CNO"]] <- paste0("CNO" , soma@colData@listData[["CNO"]])
+
+soma@colData@listData[["CNO"]] <- as.factor(soma@colData@listData[["CNO"]])
 
@@
 
-saveRDS(limrots.result, "inst/extdata/paperdata/soma.LimROTS.rds")
+# saveRDS(limrots.result, "inst/extdata/paperdata/proteogenomicPD.LimROTS.rds")
 ```
@@
 hist(limrots.result$q_values$qvalues)
+hist(limrots.result$FDR)
 
@@
 ```{r}
+limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
+                                LimROTS.FC = limrots.result$corrected.logfc,
+                                q = limrots.result$q_values$qvalues)
+limrots.result.df <- limrots.result.df[limrots.result.df$q < 0.05,]
+```
+
+```{r}
 library(SomaScan.db)
@@
 library(SummarizedExperiment, quietly = T, warn.conflicts = T)
-data("Columbia.RealCase")
-data <- assay(Columbia.RealCase)
-formula.str <- "~ 0 + groups + Age + gender + LRRK2.status + GBA.status"
-meta.info <- data.frame(colData(Columbia.RealCase), check.names = FALSE, row.names = row.names(colData(Columbia.RealCase)))
+data("soma")
+data <- assay(soma)
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"
+meta.info <- data.frame(colData(soma), check.names = FALSE, row.names = row.names(colData(soma)))
 meta.info$sample.id <- NULL
+meta.info$CNO <- as.factor(meta.info$CNO)
 design.matrix <- model.matrix(formula(formula.str), data = meta.info)
@@
 fit <- lmFit(data, design.matrix)
-cont_matrix <- makeContrasts("groups1-groups2",  levels=design.matrix)
+cont_matrix <- makeContrasts("group_comp1-group_comp2",  levels=design.matrix)
 fit2 <- contrasts.fit(fit, cont_matrix)
 fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
-limma.result <- topTable(fit.ebayes, coef="groups1-groups2" , number = "Inf")
+limma.result <- topTable(fit.ebayes, coef="group_comp1-group_comp2" , number = "Inf")
 
@@
 plotMD(fit.ebayes)
+
+length(which(limma.result$adj.P.Val < 0.05))
 
@@
 
-rots <- ROTS(data, groups = as.numeric(meta.info$groups), B = 1000, progress = TRUE,
+rots <- ROTS(data, groups = as.numeric(meta.info$group_comp), B = 1000, progress = TRUE,
              K = 600, a1= 5, a2 = 1)

### 2024-09-25 inst/extdata/paperdata/paper.qmd
.

@@
 set.seed(1234, kind = "default")
+keep_samples <- colData(soma)$CNO != 12
+soma <- soma[, keep_samples]
 
 # Set metadata and formula for LimROTS analysis
-meta.info <- names(soma@colData@listData)
+meta.info <- names(soma@colData@listData)[1:6]
 B <- 1000  # Number of bootstrap samples
@@
 group.name <- "group_comp"
-formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"  # Formula for group comparison
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled

### 2024-09-26 inst/extdata/paperdata/paper.qmd
.

@@
 
-limrots.result <- readRDS("paperdata/UPS1.Case0.LimROTS.rds")
-rots.result <- readRDS("paperdata/UPS1.Case0.ROTS.rds")
-limma.result <- readRDS("paperdata/UPS1.Case0.limma.rds")
-t.test.result <- readRDS("paperdata/UPS1.Case0.t.test.rds")
-samr.result <- readRDS("paperdata/UPS1.Case0.samr.rds")
+#limrots.result <- readRDS("paperdata/UPS1.Case0.LimROTS.rds")
+rots.result <- readRDS("../../extdata/paperdata/UPS1.Case0.ROTS.rds")
+limma.result <- readRDS("../../extdata/paperdata/UPS1.Case0.limma.rds")
+t.test.result <- readRDS("../../extdata/paperdata/UPS1.Case0.t.test.rds")
+samr.result <- readRDS("../../extdata/paperdata/UPS1.Case0.samr.rds")
 
@@
 library(parallel)
+library(SummarizedExperiment)
 # Load the dataset
@@
 # Set metadata and formula for LimROTS analysis
-meta.info <- names(soma@colData@listData)[1:6]
-B <- 1000  # Number of bootstrap samples
+meta.info <- names(soma@colData@listData)
+B <- 2000  # Number of bootstrap samples
 K <- dim(soma)[1]/2  # Set the value for K based on the data size
@@
 group.name <- "group_comp"
-formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2"  # Formula for group comparison
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled
@@
 library(SummarizedExperiment, quietly = T, warn.conflicts = T)
-data("soma")
+
 data <- assay(soma)
@@
 
-length(which(limma.result$adj.P.Val < 0.05))
+length(which(limma.result$adj.P.Val < 0.01))
 

### 2024-09-26 R/LM_with_limma.permuting.r

@@ -60,7 +60,8 @@ testStatistic_with_covariates_permutating <- function(data, group.name, meta.inf
         df.temp$sample.id <- i
         covariates.p <- rbind(covariates.p , df.temp)
     }
-    covariates.p <- covariates.p[sample(nrow(covariates.p)), ]
+    #covariates.p <- covariates.p[sample(nrow(covariates.p)), ]
+    covariates.p[,group.name] <- sample(covariates.p[,group.name])
     covariates.p$sample.id <- NULL
     row.names(covariates.p) <- NULL
     design.matrix <- model.matrix(formula(formula.str), data = covariates.p)

### 2024-09-26 inst/extdata/paperdata/paper.qmd
.

@@
 meta.info <- names(soma@colData@listData)
-B <- 2000  # Number of bootstrap samples
-K <- dim(soma)[1]/2  # Set the value for K based on the data size
+B <- 1000  # Number of bootstrap samples
+K <- dim(soma)[1]/4  # Set the value for K based on the data size
 K <- floor(K) 
@@
 
-limrots.result <- LimROTS(data.exp = soma, 
+limrots.result2 <- LimROTS(data.exp = soma, 
                           B = B, K = K, meta.info = meta.info, 
@@
                                 LimROTS.FC = limrots.result$corrected.logfc,
-                                q = limrots.result$q_values$qvalues)
+                                q = limrots.result$q_values$qvalues,
+                                p.value = limrots.result$pvalue)
+
 limrots.result.df <- limrots.result.df[limrots.result.df$q < 0.05,]
+```
+
+```{r}
+set.seed(1234, kind = "default")
+
+soma0.5 <- soma[, sample(1:262, 262) , drop = FALSE]
+
+set.seed(1234, kind = "default")
+keep_samples <- colData(soma0.5)$CNO != 12
+soma0.5 <- soma0.5[, keep_samples]
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- names(soma0.5@colData@listData)
+B <- 1000  # Number of bootstrap samples
+K <- dim(soma0.5)[1]/4  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "group_comp"
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"  # Formula for group comparison
+# Run LimROTS analysis with trend and robust settings enabled
+
+soma0.5@colData@listData[["CNO"]] <- paste0("CNO" , soma0.5@colData@listData[["CNO"]])
+
+soma0.5@colData@listData[["CNO"]] <- as.factor(soma0.5@colData@listData[["CNO"]])
+
+limrots.result2 <- LimROTS(data.exp = soma0.5, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE, seed.cl = 1234)
+
 ```

### 2024-09-27 inst/extdata/paperdata/paper.qmd

@@
 ```{r}
+{r}
+# Load necessary packages
+library(LimROTS)
+library(parallel)
+library(SummarizedExperiment)
+# Load the dataset
+data("proteogenomicPD")
+print(soma)
+
 set.seed(1234, kind = "default")
-
-soma0.5 <- soma[, sample(1:262, 262) , drop = FALSE]
-
+keep_samples <- colData(soma)$CNO != 12
+soma <- soma[, keep_samples]
+
+# Set metadata and formula for LimROTS analysis
+meta.info <- names(soma@colData@listData)
+B <- 1000  # Number of bootstrap samples
+K <- dim(soma)[1]/4  # Set the value for K based on the data size
+K <- floor(K) 
+num_cores <- 30  # Number of cores for parallel processing
+cluster <- makeCluster(num_cores)  # Create a parallel cluster
+group.name <- "group_comp"
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO"  # Formula for group comparison
+# Run LimROTS analysis with trend and robust settings enabled
+
+soma@colData@listData[["CNO"]] <- paste0("CNO" , soma@colData@listData[["CNO"]])
+
+soma@colData@listData[["CNO"]] <- as.factor(soma@colData@listData[["CNO"]])
+
+limrots.result <- LimROTS(data.exp = soma, 
+                          B = B, K = K, meta.info = meta.info, 
+                          cluster = cluster, group.name = group.name, 
+                          formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE, seed.cl = 1234)
+
+# saveRDS(limrots.result, "inst/extdata/paperdata/proteogenomicPD.LimROTS.rds")
+```
+
+```{r}
+plot(limrots.result2$q_values)
+hist(limrots.result2$q_values)
+hist(limrots.result2$q_values$qvalues)
+hist(limrots.result2$FDR)
+
+summary(limrots.result$q_values)
+```
+
+```{r}
 set.seed(1234, kind = "default")
-keep_samples <- colData(soma0.5)$CNO != 12
-soma0.5 <- soma0.5[, keep_samples]
+
+soma0.5 <- soma[, sample(1:524, 393) , drop = FALSE]
+
+keep_samples <- colData(soma0.5)$CNO %in% names(which(table(colData(soma0.5)$CNO) <= 2))
+soma0.5 <- soma0.5[, !keep_samples]
 
@@
 
-limrots.result2 <- LimROTS(data.exp = soma0.5, 
+soma0.5 <- LimROTS(data.exp = soma0.5, 
                           B = B, K = K, meta.info = meta.info, 
@@
 
+```
+
+```{r}
+plot(soma0.5$q_values)
+hist(soma0.5$q_values)
+hist(soma0.5$q_values$qvalues)
+hist(soma0.5$FDR)
+
+summary(soma0.5$q_values)
 ```

### 2024-09-27 inst/extdata/paperdata/paper.qmd
.

@@
 ```{r}
-{r}
+
 # Load necessary packages
@@
 
-# saveRDS(limrots.result, "inst/extdata/paperdata/proteogenomicPD.LimROTS.rds")
-```
-
-```{r}
-plot(limrots.result2$q_values)
-hist(limrots.result2$q_values)
-hist(limrots.result2$q_values$qvalues)
-hist(limrots.result2$FDR)
+saveRDS(limrots.result, "D:/LimROTS/inst/extdata/paperdata/Soma.noPCs.LimROTS.rds")
+
+```
+
+```{r}
+plot(limrots.result$q_values)
+hist(limrots.result$q_values)
+hist(limrots.result$q_values$qvalues)
+hist(limrots.result$FDR)
 
@@
 group.name <- "group_comp"
-formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"  # Formula for group comparison
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled
-
-soma0.5@colData@listData[["CNO"]] <- paste0("CNO" , soma0.5@colData@listData[["CNO"]])
-
-soma0.5@colData@listData[["CNO"]] <- as.factor(soma0.5@colData@listData[["CNO"]])
 
@@
                           formula.str = formula.str, trend = TRUE, robust = TRUE, progress = TRUE, seed.cl = 1234)
+
+saveRDS(soma0.5, "D:/LimROTS/inst/extdata/paperdata/Soma0.5.noPCs.LimROTS.rds")
 

### 2024-09-27 inst/extdata/paperdata/paper.qmd
.

@@
 
-soma0.5 <- soma[, sample(1:524, 393) , drop = FALSE]
+soma0.5 <- soma[, sample(1:524, 262) , drop = FALSE]
 
@@
 B <- 1000  # Number of bootstrap samples
-K <- dim(soma0.5)[1]/4  # Set the value for K based on the data size
+K <- dim(soma0.5)[1]/2  # Set the value for K based on the data size
 K <- floor(K) 
@@
 
-saveRDS(soma0.5, "D:/LimROTS/inst/extdata/paperdata/Soma0.5.noPCs.LimROTS.rds")
-
-```
-
-```{r}
+saveRDS(soma0.5, "../paperdata/Soma0.5.noPCs.LimROTS.rds")
+
+```
+
+```{r}
+soma0.5 <- readRDS("Soma0.5.noPCs.LimROTS.rds")
+
 plot(soma0.5$q_values)
@@
 ```{r}
+
+soma.limrots <- readRDS("Soma.noPCs.LimROTS.rds")
+soma0.5.limrots <- readRDS("Soma0.5.noPCs.LimROTS.rds")
+
+soma.limrots <- row.names(soma.limrots$data)[which(soma.limrots$q_values$qvalues < 0.05)]
+soma0.5.limrots <- row.names(soma0.5.limrots$data)[which(soma0.5.limrots$q_values$qvalues < 0.05)]
+
+length(intersect(soma.sig, soma0.5.sig))
+```
+
+```{r}
 library(SomaScan.db)
@@
 
+soma0.5 <- soma[, sample(1:524, 262) , drop = FALSE]
+
+keep_samples <- colData(soma0.5)$CNO %in% names(which(table(colData(soma0.5)$CNO) <= 2))
+soma0.5 <- soma0.5[, !keep_samples]
+
+data <- assay(soma0.5)
+
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO"
+meta.info <- data.frame(colData(soma0.5), check.names = FALSE, row.names = row.names(colData(soma0.5)))
+meta.info$sample.id <- NULL
+meta.info$CNO <- as.factor(meta.info$CNO)
+design.matrix <- model.matrix(formula(formula.str), data = meta.info)
+colnames(design.matrix) <- make.names(colnames(design.matrix))
+fit <- lmFit(data, design.matrix)
+cont_matrix <- makeContrasts("group_comp1-group_comp2",  levels=design.matrix)
+fit2 <- contrasts.fit(fit, cont_matrix)
+fit.ebayes <- eBayes(fit2, trend=TRUE, robust=TRUE)
+limma.result.0.5 <- topTable(fit.ebayes, coef="group_comp1-group_comp2" , number = "Inf")
+
+volcanoplot(fit.ebayes)
+plotMD(fit.ebayes)
+
+length(which(limma.result.0.5$adj.P.Val < 0.05))
+
+```
+
+```{r}
+library(limma)
+library(SummarizedExperiment, quietly = T, warn.conflicts = T)
+
+keep_samples <- colData(soma)$CNO != 12
+soma <- soma[, keep_samples]
+
 data <- assay(soma)
-formula.str <- "~ 0 + group_comp + gender + AGE + CNO + pc1 + pc2 + pc3 + pc4"
+
+formula.str <- "~ 0 + group_comp + gender + AGE + CNO"
 meta.info <- data.frame(colData(soma), check.names = FALSE, row.names = row.names(colData(soma)))
@@
 
-length(which(limma.result$adj.P.Val < 0.01))
-
-# Create a volcano plot
-ggplot(limma.result, aes(x = logFC, y = -log10(P.Value))) +
-  geom_point(alpha = 0.8) +
-  theme_bw() +
-  labs(title = "Volcano Plot", x = "Log Fold Change", y = "-Log10 q.value", color = "True Positive") +
-  scale_color_manual(values = c("grey", "red"))+
-  geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "blue")+
-  geom_hline(yintercept=-log10(0.01), linetype="dashed", color = "black")
+length(which(limma.result$adj.P.Val < 0.05))
+
+length(intersect(which(limma.result.0.5$adj.P.Val < 0.05) , which(limma.result$adj.P.Val < 0.05)))
 ```

### 2024-09-28 inst/extdata/paperdata/paper.qmd
.

@@
 # Set metadata and formula for LimROTS analysis
-meta.info <- names(soma@colData@listData)
+meta.info <- names(soma@colData@listData)[1:4]
 B <- 1000  # Number of bootstrap samples
@@
 
-saveRDS(limrots.result, "D:/LimROTS/inst/extdata/paperdata/Soma.noPCs.LimROTS.rds")
+saveRDS(limrots.result, "../paperdata/Soma.noPCs.LimROTS.rds")
 
@@
 ```{r}
+library(LimROTS)
+library(parallel)
+
 set.seed(1234, kind = "default")
 
-soma0.5 <- soma[, sample(1:524, 262) , drop = FALSE]
+data("proteogenomicPD")
+
+soma@colData@listData[["CNO"]] <- paste0("CNO" , soma@colData@listData[["CNO"]])
+
+soma@colData@listData[["CNO"]] <- as.factor(soma@colData@listData[["CNO"]])
+
+group1 <- sample(which(colData(soma)$group_comp == 1), length(which(colData(soma)$group_comp == 1))/1.8 )
+
+group2 <- sample(which(colData(soma)$group_comp == 2), length(which(colData(soma)$group_comp == 2))/1.8 )
+
+soma0.5 <- soma[, c(group1, group2) , drop = FALSE]
 
@@
 
+colData(soma0.5)$CNO <- droplevels(colData(soma0.5)$CNO)
+
+soma0.5@assays@data@listData[["norm"]] <- preprocessCore::normalize.quantiles(as.matrix(soma0.5@assays@data@listData[["norm"]]))
+
 # Set metadata and formula for LimROTS analysis
-meta.info <- names(soma0.5@colData@listData)
+meta.info <- names(soma0.5@colData@listData)[1:3]
 B <- 1000  # Number of bootstrap samples
-K <- dim(soma0.5)[1]/2  # Set the value for K based on the data size
+K <- dim(soma0.5)[1]/4 # Set the value for K based on the data size
 K <- floor(K) 
@@
 group.name <- "group_comp"
-formula.str <- "~ 0 + group_comp + gender + AGE + CNO"  # Formula for group comparison
+formula.str <- "~ 0 + group_comp + gender + AGE"  # Formula for group comparison
 # Run LimROTS analysis with trend and robust settings enabled
 
-soma0.5 <- LimROTS(data.exp = soma0.5, 
+trace(LimROTS, edit = T)
+
+soma0.5.result <- LimROTS(data.exp = soma0.5, 
                           B = B, K = K, meta.info = meta.info, 
@@
 
-saveRDS(soma0.5, "../paperdata/Soma0.5.noPCs.LimROTS.rds")
-
-```
-
-```{r}
-soma0.5 <- readRDS("Soma0.5.noPCs.LimROTS.rds")
-
-plot(soma0.5$q_values)
-hist(soma0.5$q_values)
-hist(soma0.5$q_values$qvalues)
-hist(soma0.5$FDR)
-
-summary(soma0.5$q_values)
+saveRDS(soma0.5.result, "../paperdata/Soma0.5.noPCs.LimROTS.rds")
+
+```
+
+```{r}
+soma0.5.result <- readRDS("Soma0.5.noPCs.LimROTS.rds")
+
+plot(soma0.5.result$q_values)
+hist(soma0.5.result$q_values)
+hist(soma0.5.result$q_values$qvalues)
+hist(soma0.5.result$FDR)
+
+summary(soma0.5.result$q_values)
 ```
@@
 
-length(intersect(soma.sig, soma0.5.sig))
+length(intersect(soma.limrots, soma0.5.limrots))
 ```
@@
 
-soma0.5 <- soma[, sample(1:524, 262) , drop = FALSE]
-
-keep_samples <- colData(soma0.5)$CNO %in% names(which(table(colData(soma0.5)$CNO) <= 2))
-soma0.5 <- soma0.5[, !keep_samples]
+set.seed(1234, kind = "default")
 
@@
 
-keep_samples <- colData(soma)$CNO != 12
-soma <- soma[, keep_samples]
+data("proteogenomicPD")
+
+soma@colData@listData[["CNO"]] <- paste0("CNO" , soma@colData@listData[["CNO"]])
+
+soma@colData@listData[["CNO"]] <- as.factor(soma@colData@listData[["CNO"]])
+
+keep_samples <- colData(soma)$CNO %in% names(which(table(colData(soma0.5)$soma) <= 2))
+soma <- soma[, !keep_samples]
+
+colData(soma0.5)$CNO <- droplevels(colData(soma0.5)$CNO)
 

### 2024-09-28 inst/extdata/paperdata/tomakeapackae.r
.

@@ -1,35 +1,0 @@
-
-gitcreds::gitcreds_set()
-
-usethis::create_package(".")
-
-############################
-
-roxygen2::roxygenize()
-
-#Rcpp::compileAttributes()
-
-
-usethis::use_testthat()
-
-library(devtools)
-#test()
-
-
-
-roxygen2::roxygenize()
-
-devtools::check(vignettes = FALSE)
-
-devtools::test()
-
-
-devtools::check(vignettes = TRUE)
-
-
-devtools::build(vignettes = FALSE)
-
-devtools::build_vignettes(quiet = FALSE)
-
-library(BiocCheck)
-BiocCheck(checkDir = "../LimROTS paper/")

### 2024-09-29 R/Bootstrap_functions.r
.

@@ -24,11 +24,7 @@
 #'
 #' # Paired bootstrap sampling
 #' bootstrapS(B = 10, meta.info = meta.info, group.name = "group", paired = TRUE)
-
-
-
-
-bootstrapS <- function (B, meta.info, group.name , paired) {
+bootstrapS <- function(B, meta.info, group.name, paired) {
     groups <- meta.info[, group.name]
     bootsamples <- matrix(nrow = B, ncol = length(groups))
     for (i in seq_len(B)) {
@@ -39,8 +35,8 @@ bootstrapS <- function (B, meta.info, group.name , paired) {
     }
     if (paired) {
         for (i in seq_len(B)) {
-            g.names1 <-  bootsamples[i, which(groups == unique(groups)[1])]
-            g.names2 <- match(g.names1 , row.names(meta.info)) + length(g.names1)
+            g.names1 <- bootsamples[i, which(groups == unique(groups)[1])]
+            g.names2 <- match(g.names1, row.names(meta.info)) + length(g.names1)
             bootsamples[i, which(groups == unique(groups)[2])] <- row.names(meta.info)[g.names2]
         }
     }
@@ -68,9 +64,7 @@ bootstrapS <- function (B, meta.info, group.name , paired) {
 #' set.seed(123)
 #' meta.info <- data.frame(group = rep(c("A", "B"), each = 5), row.names = paste0("Sample", 1:10))
 #' permutatedS(meta.info = meta.info, B = 10)
-
-
-permutatedS <- function (meta.info, B)
+permutatedS <- function(meta.info, B)
 {
     persamples <- matrix(nrow = B, ncol = nrow(meta.info))
     for (i in seq_len(B)) {
@@ -103,13 +97,11 @@ permutatedS <- function (meta.info, B)
 #' # Example usage:
 #' set.seed(123)
 #' meta.info <- data.frame(group = rep(c(1, 2), each = 5),
-#'                         batch = rep(c("A", "B"), 5),
-#'                         row.names = paste0("Sample", 1:10))
+#'     batch = rep(c("A", "B"), 5),
+#'     row.names = paste0("Sample", 1:10))
 #' meta.info$batch <- as.factor(meta.info$batch)
 #' bootstrapSamples.limRots(B = 10, meta.info = meta.info, group.name = "group")
-
-
-bootstrapSamples.limRots <- function (B, meta.info , group.name)
+bootstrapSamples.limRots <- function(B, meta.info, group.name)
 {
[…]

### 2024-09-29 R/LM_with_limma.fit.r

@@ -48,48 +48,48 @@
 
 testStatistic_with_covariates_Fit <- function(x,
                                               group.name,
-                                              meta.info ,
+                                              meta.info,
                                               formula.str,
                                               trend,
                                               robust) {
     data <- x
     combined_data <- data.frame(
         check.rows = FALSE,
-        check.names = FALSE ,
-        none = rep("none" , nrow(data[[1]]))
+        check.names = FALSE,
+        none = rep("none", nrow(data[[1]]))
     )
     for (k.list in names(data)) {
         combined_data <- cbind(combined_data,
-                               data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE))
+            data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE))
     }
     combined_data <- combined_data[, -1]
     design.matrix <- model.matrix(formula(formula.str), data = meta.info)
     colnames(design.matrix) <- make.names(colnames(design.matrix))
     fit <- lmFit(combined_data, design.matrix)
     if (length(data) == 2) {
-        pairwise_contrasts <- paste0(group.name , unique(meta.info[, group.name]))
+        pairwise_contrasts <- paste0(group.name, unique(meta.info[, group.name]))
         pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x)
             paste(x[1], "-", x[2]))
         cont_matrix <- makeContrasts(contrasts = pairwise_contrasts, levels =
-                                         design.matrix)
+            design.matrix)
         fit2 <- contrasts.fit(fit, cont_matrix)
         fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
         d_values <- topTable(
             fit.ebayes,
-            coef = pairwise_contrasts ,
+            coef = pairwise_contrasts,
             number = "Inf",
             sort.by = 'none'
         )
         corrected.logfc <- d_values$logFC
         d_values <- abs(d_values$logFC)
-        s_values <- as.numeric(sqrt(fit.ebayes$s2.post)  * fit.ebayes$stdev.unscaled[, 1])
+        s_values <- as.numeric(sqrt(fit.ebayes$s2.post) * fit.ebayes$stdev.unscaled[, 1])
         return(list(
             d = d_values,
-            s = s_values ,
+            s = s_values,
             corrected.logfc = corrected.logfc
         ))
[…]

### 2024-09-29 R/LM_with_limma.r

@@ -42,29 +42,29 @@
 testStatistic_with_covariates <- function(x,
                                           group.name,
                                           meta.info,
-                                          formula.str ,
+                                          formula.str,
                                           trend,
                                           robust) {
     data <- x
     combined_data <- data.frame(
         check.rows = FALSE,
-        check.names = FALSE ,
-        none = rep("none" , nrow(data[[1]]))
+        check.names = FALSE,
+        none = rep("none", nrow(data[[1]]))
     )
     for (k.list in names(data)) {
         combined_data <- cbind(combined_data,
-                               data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE))
+            data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE))
     }
     combined_data <- combined_data[, -1]
-    colnames(combined_data) <- paste0(colnames(combined_data), "." , seq(1, ncol(combined_data)))
+    colnames(combined_data) <- paste0(colnames(combined_data), ".", seq(1, ncol(combined_data)))
     covariates.p <- data.frame()
     meta.info.temp <- meta.info
     meta.info.temp$sample.id <- row.names(meta.info.temp)
     for (i in colnames(combined_data)) {
-        real_SampleNames <-  str_split_fixed(i , fixed(".") , 2)[, 1]
+        real_SampleNames <- str_split_fixed(i, fixed("."), 2)[, 1]
         df.temp <- meta.info.temp[row.names(meta.info.temp) %in% real_SampleNames, ]
         df.temp$sample.id <- i
-        covariates.p <- rbind(covariates.p , df.temp)
+        covariates.p <- rbind(covariates.p, df.temp)
     }
     covariates.p$sample.id <- NULL
     row.names(covariates.p) <- NULL
@@ -72,24 +72,24 @@ testStatistic_with_covariates <- function(x,
     colnames(design.matrix) <- make.names(colnames(design.matrix))
     fit <- lmFit(combined_data, design.matrix)
     if (length(data) == 2) {
-        pairwise_contrasts <- paste0(group.name , unique(covariates.p[, group.name]))
+        pairwise_contrasts <- paste0(group.name, unique(covariates.p[, group.name]))
         pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x)
             paste(x[1], "-", x[2]))
[…]

### 2024-09-29 R/SanityChecK.R

@@ -37,7 +37,7 @@ SanityChecK <- function(x,
                         a1 = NULL,
                         a2 = NULL,
                         meta.info = NULL,
-                        group.name = NULL ,
+                        group.name = NULL,
                         formula.str = NULL,
                         survival = FALSE,
                         paired = FALSE,
@@ -76,7 +76,7 @@ SanityChecK <- function(x,
 
         if (is.null(meta.info)) {
             stop("meta.info should be a vector of colData names to be used")
-        } else{
+        } else {
             meta.info.colnames <- meta.info
             meta.info <- data.frame(
                 colData(data.exp)[, meta.info],
@@ -85,7 +85,7 @@ SanityChecK <- function(x,
             )
             if (length(meta.info) != length(meta.info.colnames)) {
                 stop("meta.info should be a vector of colData names to be used")
-            } else{
+            } else {
                 colnames(meta.info) <- meta.info.colnames
             }
 
@@ -95,9 +95,9 @@ SanityChecK <- function(x,
                 "group.name should be a string specifying the column in `meta.info` that represents the groups or conditions for comparison."
             )
         }
-        message(sprintf("Assay: %s will be used" , assayNames(data.exp)[1]))
-        data <- assay(data.exp , assayNames(data.exp)[1])
-    } else{
+        message(sprintf("Assay: %s will be used", assayNames(data.exp)[1]))
+        data <- assay(data.exp, assayNames(data.exp)[1])
+    } else {
         data <- data.exp
         groups <- meta.info[, group.name]
     }
@@ -105,7 +105,7 @@ SanityChecK <- function(x,
     if (any(!row.names(meta.info) %in% colnames(data))) {
         stop("rownames for meta.info should match the data colnames (samples names)")
     }
-    if (any(grepl("." , colnames(data) , fixed = TRUE))) {
+    if (any(grepl(".", colnames(data), fixed = TRUE))) {
         stop("Sample names should contains no '.', please remove it if any")
     }
     if (!is.null(meta.info) & n.ROTS == FALSE) {
@@ -116,7 +116,7 @@ SanityChecK <- function(x,
             if (is.null(formula.str)) {
                 stop("formula.str should by provided for the model")
             }
-        } else{
+        } else {
             message(
                 "A meta.info table is provided with covariates >>> LimROTS with covariates will be used"
             )
@@ -124,7 +124,7 @@ SanityChecK <- function(x,
[…]

### 2024-09-29 R/calOverlaps.R

@@ -48,10 +48,10 @@ calOverlaps <- function(D,
     for (b in idx_b) {
         res1 <- abs(D[((b - 1) * nrow + 1):(b * nrow)] / (S[((b - 1) * nrow + 1):(b * nrow)] + ssq))
         res2 <- abs(D[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] /
-                        (S[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] + ssq))
+            (S[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] + ssq))
         pres1 <- abs(pD[((b - 1) * nrow + 1):(b * nrow)] / (pS[((b - 1) * nrow + 1):(b * nrow)] + ssq))
         pres2 <- abs(pD[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] /
-                         (pS[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] + ssq))
+            (pS[((b + idx_offset - 1) * nrow + 1):((b + idx_offset) * nrow)] + ssq))
 
         sorted_res <- sort2_1R(res1, res2)
         res1 <- sorted_res$a

### 2024-09-29 tests/testthat/testBootLim.R

@@ -3,32 +3,32 @@ library(limma)
 library(stringr)
 
 test_that("testStatistic_with_covariates_Fit returns expected output structure", {
-  # Sample data with 10 samples per group
-  data <- list(
-    group1 = t(matrix(rnorm(400), nrow = 10)),  # 10 samples, 4 features
-    group2 = t(matrix(rnorm(400), nrow = 10))   # 10 samples, 4 features
-  )
+    # Sample data with 10 samples per group
+    data <- list(
+        group1 = t(matrix(rnorm(400), nrow = 10)), # 10 samples, 4 features
+        group2 = t(matrix(rnorm(400), nrow = 10)) # 10 samples, 4 features
+    )
 
-  data[[1]] <- data.frame(data[[1]])
-  colnames(data[[1]]) <- paste0("sample", 1:10)
-  data[[2]] <- data.frame(data[[2]])
-  colnames(data[[2]]) <- paste0("sample", 11:20)
+    data[[1]] <- data.frame(data[[1]])
+    colnames(data[[1]]) <- paste0("sample", 1:10)
+    data[[2]] <- data.frame(data[[2]])
+    colnames(data[[2]]) <- paste0("sample", 11:20)
 
-  # Sample meta info with corresponding group labels
-  meta.info <- data.frame(
-    row.names  = c(paste0("sample", 1:10), paste0("sample", 11:20)),
-    group = as.factor( c(rep(1, 10), rep(2, 10)) )
-  )
+    # Sample meta info with corresponding group labels
+    meta.info <- data.frame(
+        row.names = c(paste0("sample", 1:10), paste0("sample", 11:20)),
+        group = as.factor(c(rep(1, 10), rep(2, 10)))
+    )
 
-  # Formula string for modeling
-  formula.str <- "~ 0 + group"
+    # Formula string for modeling
+    formula.str <- "~ 0 + group"
 
-  # Call the function
-  result <- testStatistic_with_covariates(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
+    # Call the function
+    result <- testStatistic_with_covariates(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
 
-  # Check output structure
-  expect_type(result, "list")
-  expect_equal(length(result), 2)  # Expecting 3 elements in the output list
-  expect_true("d" %in% names(result))  # Check for 'd' in result
-  expect_true("s" %in% names(result))  # Check for 's' in result
+    # Check output structure
[…]

### 2024-09-29 tests/testthat/testBootstrapLimROTS.R

@@ -1,23 +1,23 @@
 library(testthat)
 
 test_that("bootstrapSamples.limRots returns expected output structure", {
-  # Sample meta info with two groups and one factor
-  meta.info <- data.frame(
-    row.names  = paste0("sample", 1:20),
-    group = c(rep(1, 10), rep(2, 10)),
-    factor1 = factor(rep(c("A", "B"), each = 10))
-  )
-  row.names(meta.info) <- meta.info$sample_id
+    # Sample meta info with two groups and one factor
+    meta.info <- data.frame(
+        row.names = paste0("sample", 1:20),
+        group = c(rep(1, 10), rep(2, 10)),
+        factor1 = factor(rep(c("A", "B"), each = 10))
+    )
+    row.names(meta.info) <- meta.info$sample_id
 
-  # Parameters
-  B <- 10  # Number of bootstrap samples
-  group.name <- "group"
+    # Parameters
+    B <- 10 # Number of bootstrap samples
+    group.name <- "group"
 
-  # Call the function
-  result <- bootstrapSamples.limRots(data = NULL, B = B, meta.info = meta.info, group.name = group.name)
+    # Call the function
+    result <- bootstrapSamples.limRots(data = NULL, B = B, meta.info = meta.info, group.name = group.name)
 
-  # Check output structure
-  expect_type(result, "character")
-  expect_equal(dim(result), c(B, nrow(meta.info)))  # Expecting B rows and nrow(meta.info) columns
+    # Check output structure
+    expect_type(result, "character")
+    expect_equal(dim(result), c(B, nrow(meta.info))) # Expecting B rows and nrow(meta.info) columns
 
 })

### 2024-09-29 tests/testthat/testFitLim.R

@@ -2,28 +2,28 @@ library(testthat)
 library(limma)
 
 test_that("testStatistic_with_covariates_Fit returns expected output structure", {
-  # Sample data with 10 samples per group
-  data <- list(
-    group1 = t(matrix(rnorm(400), nrow = 10)),  # 10 samples, 4 features
-    group2 = t(matrix(rnorm(400), nrow = 10))   # 10 samples, 4 features
-  )
+    # Sample data with 10 samples per group
+    data <- list(
+        group1 = t(matrix(rnorm(400), nrow = 10)), # 10 samples, 4 features
+        group2 = t(matrix(rnorm(400), nrow = 10)) # 10 samples, 4 features
+    )
 
-  # Sample meta info with corresponding group labels
-  meta.info <- data.frame(
-    sample_id = c(paste0("sample", 1:10), paste0("sample", 11:20)),
-    group = as.factor( c(rep(1, 10), rep(2, 10)) )
-  )
+    # Sample meta info with corresponding group labels
+    meta.info <- data.frame(
+        sample_id = c(paste0("sample", 1:10), paste0("sample", 11:20)),
+        group = as.factor(c(rep(1, 10), rep(2, 10)))
+    )
 
-  # Formula string for modeling
-  formula.str <- "~ 0 + group"
+    # Formula string for modeling
+    formula.str <- "~ 0 + group"
 
-  # Call the function
-  result <- testStatistic_with_covariates_Fit(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
+    # Call the function
+    result <- testStatistic_with_covariates_Fit(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
 
-  # Check output structure
-  expect_type(result, "list")
-  expect_equal(length(result), 3)  # Expecting 3 elements in the output list
-  expect_true("d" %in% names(result))  # Check for 'd' in result
-  expect_true("s" %in% names(result))  # Check for 's' in result
-  expect_true("corrected.logfc" %in% names(result))  # Check for 'corrected.logfc' in result
+    # Check output structure
+    expect_type(result, "list")
+    expect_equal(length(result), 3) # Expecting 3 elements in the output list
+    expect_true("d" %in% names(result)) # Check for 'd' in result
+    expect_true("s" %in% names(result)) # Check for 's' in result
+    expect_true("corrected.logfc" %in% names(result)) # Check for 'corrected.logfc' in result
 })

### 2024-09-29 tests/testthat/testPermutatedS.R

@@ -1,25 +1,25 @@
 library(testthat)
 
 test_that("permutatedS returns expected output structure", {
-  # Sample meta info with some samples
-  meta.info <- data.frame(
-    sample_id = paste0("sample", 1:20)
-  )
-  row.names(meta.info) <- meta.info$sample_id
+    # Sample meta info with some samples
+    meta.info <- data.frame(
+        sample_id = paste0("sample", 1:20)
+    )
+    row.names(meta.info) <- meta.info$sample_id
 
-  # Parameters
-  B <- 5  # Number of permuted samples
+    # Parameters
+    B <- 5 # Number of permuted samples
 
-  # Call the function
-  result <- permutatedS(meta.info, B)
+    # Call the function
+    result <- permutatedS(meta.info, B)
 
-  # Check output structure
-  expect_type(result, "character")
-  expect_equal(dim(result), c(B, nrow(meta.info)))  # Expecting B rows and nrow(meta.info) columns
+    # Check output structure
+    expect_type(result, "character")
+    expect_equal(dim(result), c(B, nrow(meta.info))) # Expecting B rows and nrow(meta.info) columns
 
-  # Check that each row is a permutation of the original row names
-  for (i in seq_len(B)) {
-    expect_true(all(result[i, ] %in% row.names(meta.info)))  # All sampled names should be in meta.info
-    expect_equal(length(unique(result[i, ])), nrow(meta.info))  # Each row should have unique samples
-  }
-} )
+    # Check that each row is a permutation of the original row names
+    for (i in seq_len(B)) {
+        expect_true(all(result[i, ] %in% row.names(meta.info))) # All sampled names should be in meta.info
+        expect_equal(length(unique(result[i, ])), nrow(meta.info)) # Each row should have unique samples
+    }
+})

### 2024-09-29 tests/testthat/testPermutatingLim.R

@@ -3,32 +3,32 @@ library(limma)
 library(stringr)
 
 test_that("testStatistic_with_covariates_Fit returns expected output structure", {
-  # Sample data with 10 samples per group
-  data <- list(
-    group1 = t(matrix(rnorm(400), nrow = 10)),  # 10 samples, 4 features
-    group2 = t(matrix(rnorm(400), nrow = 10))   # 10 samples, 4 features
-  )
+    # Sample data with 10 samples per group
+    data <- list(
+        group1 = t(matrix(rnorm(400), nrow = 10)), # 10 samples, 4 features
+        group2 = t(matrix(rnorm(400), nrow = 10)) # 10 samples, 4 features
+    )
 
-  data[[1]] <- data.frame(data[[1]])
-  colnames(data[[1]]) <- paste0("sample", 1:10)
-  data[[2]] <- data.frame(data[[2]])
-  colnames(data[[2]]) <- paste0("sample", 11:20)
+    data[[1]] <- data.frame(data[[1]])
+    colnames(data[[1]]) <- paste0("sample", 1:10)
+    data[[2]] <- data.frame(data[[2]])
+    colnames(data[[2]]) <- paste0("sample", 11:20)
 
-  # Sample meta info with corresponding group labels
-  meta.info <- data.frame(
-    row.names  = c(paste0("sample", 1:10), paste0("sample", 11:20)),
-    group = as.factor( c(rep(1, 10), rep(2, 10)) )
-  )
+    # Sample meta info with corresponding group labels
+    meta.info <- data.frame(
+        row.names = c(paste0("sample", 1:10), paste0("sample", 11:20)),
+        group = as.factor(c(rep(1, 10), rep(2, 10)))
+    )
 
-  # Formula string for modeling
-  formula.str <- "~ 0 + group"
+    # Formula string for modeling
+    formula.str <- "~ 0 + group"
 
-  # Call the function
-  result <- testStatistic_with_covariates_permutating(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
+    # Call the function
+    result <- testStatistic_with_covariates_permutating(data, "group", meta.info, formula.str, trend = FALSE, robust = FALSE)
 
-  # Check output structure
-  expect_type(result, "list")
-  expect_equal(length(result), 2)  # Expecting 3 elements in the output list
-  expect_true("d" %in% names(result))  # Check for 'd' in result
-  expect_true("s" %in% names(result))  # Check for 's' in result
+    # Check output structure
[…]

### 2024-10-01 R/LimROTS.r
vignettes

@@ -58,11 +58,8 @@
 #' @import utils
 #' @import SummarizedExperiment
 #'
-#' @details Differential expression analysis is a prevalent method utilised in the examination of diverse biological data.
-#'     The reproducibility-optimized test statistic (ROTS) modifies a t-statistic based on the data's intrinsic characteristics and ranks features according to their statistical significance for differential expression between two or more groups, as shown by the f-statistic.
-#'     Focussing on proteomics and metabolomics, the current ROTS implementation cannot account for technical or biological covariates such as MS batches or gender differences among the samples.
-#'     Consequently, we developed LimROTS, which employs a reproducibility-optimized test statistic utilising the limma methodology to simulate complex experimental designs.
-#'
+#' @details The **LimROTS** approach initially uses the \link{limma} package to simulate the intensity data of proteins and metabolites. A linear model is subsequently fitted using the design matrix. Empirical Bayes variance shrinking is then implemented. To obtain the moderated t-statistics, the adjusted standard error \eqn{SEpost = √(s2.post) \times unscaled SD} for each feature is computed, along with the regression coefficient for each feature (indicating the impact of variations in the experimental settings). The \link{ROTS} approach establishes optimality based on the largest overlap of top-ranked features within group-preserving bootstrap datasets, Finally based on the optimized parameters \eqn{\alpha1} and \eqn{\alpha2} this equation used to calculates the final statistics: \deqn{t_{\alpha}(g) = \frac{\beta}{\alpha1 + \alpha2 \times SEpost}}
+#'          where \eqn{t_{\alpha}(g)} is the final statistics for each feature, \eqn{\beta} is the coefficient, and SEpost is the the adjusted standard error. LimROTS generates p-values from permutation samples using the implementation available in \link{qvalue} package, along with internal implementation of FDR adapted from ROTS package. Additionally, the qvalue package is used to calculate q-values, were the proportion of true null p-values is set to the bootstrap method \link{pi0est}. We recommend using permutation-derived p-values and qvalues.
 #'
 #' @references
 #'   Ritchie, M.E., Phipson, B., Wu, D., Hu, Y., Law, C.W., Shi, W., and Smyth, G.K. (2015). limma powers differential
@@ -71,6 +68,9 @@
[…]

### 2024-10-01 R/LimROTS.r
vignettes

@@ -1,4 +1,4 @@
-#' LimROTS: An Extension of the ROTS Method with Limma Integration
+#' LimROTS: A Hybrid Method Integrating Empirical Bayes and Reproducibility-Optimized Statistics for Robust Analysis of Protein and Metabolite Data
 #'
 #' The `LimROTS` function employs a reproducibility-optimized test statistic utilising the limma and ROTS methodology to simulate complex experimental designs.
 #'

### 2024-10-02 R/LimROTS.r
.

@@ -20,7 +20,7 @@
 #' @param formula.str A formula string used when covariates are present in meta.info for modeling. It should include "~ 0 + ..." to exclude the intercept from the model.
 #' @param robust indicating whether robust fitting should be used. Default is TRUE, see \link{eBayes}.
 #' @param trend indicating whether to include trend fitting in the differential expression analysis. Default is TRUE. see \link{eBayes}.
-#'
+#' @param permutating.group Logical, If \code{TRUE}, the permutation for calculating the null distribution is performed by permuting the target group only specified in \code{group.name}. If FALSE, the entire \code{meta.info} will be permuted (recommended to be set to TRUE).
 #'
 #' @return A list of class `"list"` with the following elements:
 #' \item{data}{The original data matrix.}
@@ -94,7 +94,8 @@ LimROTS <- function(x,
                     n.ROTS = FALSE,
                     seed.cl = 1234,
                     robust = TRUE,
-                    trend = TRUE)
+                    trend = TRUE,
+                    permutating.group = TRUE)
 {
     SanityChecK.list <- SanityChecK(
         x,
@@ -264,7 +265,7 @@ LimROTS <- function(x,
                 formula.str = formula.str,
                 trend =
                     trend,
-                robust = robust
+                robust = robust , permutating.group = permutating.group
             )
         } else {
             pSamples.R <- split(pSamples[i, ], groups)

### 2024-10-04 R/LM_with_limma.fit.r
corrections to BiocCheck

@@ -1,106 +1,0 @@
-#' Perform Linear Modeling with Covariates using Limma
-#'
-#' This function performs linear modeling using the Limma package, incorporating covariates
-#' in the model fitting process. It is designed to handle both two-group comparisons and
-#' multi-group settings with covariates.
-#'
-#' @param x A list containing two or more data matrices where rows represent features
-#' (e.g., genes, proteins) and columns represent samples. The list should contain at least
-#' two matrices for pairwise group comparison.
-#' @param group.name A character string indicating the name of the group variable in
-#' `meta.info` to be used in the analysis.
-#' @param meta.info A data frame containing the metadata for the samples. This includes
-#' sample grouping and any covariates to be included in the model.
-#' @param formula.str A string specifying the formula to be used in model fitting. It should
-#' follow the standard R formula syntax (e.g., `~ covariate1 + covariate2`).
-#' @param trend A logical value indicating whether to allow for an intensity-dependent trend in
-#' the prior variance.
-#' @param robust A logical value indicating whether to use a robust fitting procedure to protect
-#' against outliers.
-#'
-#' @details
-#' This function combines the data matrices from different groups and fits a linear model
-#' using covariates provided in the `meta.info`. For two-group comparisons, the function
-#' computes contrasts between the two groups and applies empirical Bayes moderation. For
-#' multi-group analysis with a single covariate, pairwise contrasts are computed, and the
-#' moderated F-statistic is calculated for each feature.
-#'
-#' @return A list containing the following elements:
-#' \item{d}{A vector of the test statistics (log-fold changes or F-statistics) for each feature.}
-#' \item{s}{A vector of the standard deviations for each feature, adjusted by the empirical Bayes
-#' procedure.}
-#' \item{corrected.logfc}{The log-fold changes for each feature after fitting the model.}
-#'
-#' @seealso \code{\link[limma]{lmFit}}, \code{\link[limma]{eBayes}}, \code{\link[limma]{topTable}},
-#' \code{\link[limma]{makeContrasts}}
-#'
-#'
-#'
-#' @importFrom stats model.matrix formula
-#' @importFrom dplyr bind_cols
-#' @importFrom limma makeContrasts lmFit contrasts.fit eBayes topTable
-#' @import utils
-#'
-#'
-#'
-
-
-
[…]

### 2024-10-04 R/LM_with_limma.permuting.r

@@ -1,108 +1,0 @@
-#' Perform Permutation-Based Linear Modeling with Covariates using Limma
-#'
-#' This function performs linear modeling using the Limma package with permutation of the covariates
-#' to evaluate the test statistics under random assignments. It handles two-group comparisons and
-#' multi-group settings.
-#'
-#' @param x A list containing two or more data matrices where rows represent features (e.g., genes, proteins)
-#' and columns represent samples. The list should contain at least two matrices for pairwise group comparison.
-#' @param group.name A character string indicating the name of the group variable in `meta.info` to be used
-#' in the analysis.
-#' @param meta.info A data frame containing the metadata for the samples. This includes sample grouping and any
-#' covariates to be included in the model.
-#' @param formula.str A string specifying the formula to be used in model fitting. It should follow the standard
-#' R formula syntax (e.g., `~ covariate1 + covariate2`).
-#' @param trend A logical value indicating whether to allow for an intensity-dependent trend in the prior variance.
-#' @param robust A logical value indicating whether to use a robust fitting procedure to protect against outliers.
-#' @param permutating.group Logical, If \code{TRUE}, the permutation for calculating the null distribution is performed by permuting the target group only specified in \code{group.name}. If FALSE, the entire \code{meta.info} will be permuted (recommended to be set to TRUE).
-#'
-#' @details
-#' This function combines the data matrices from different groups and permutes the covariates from `meta.info`
-#' before fitting a linear model using Limma. Permutation helps assess how the covariates behave under random
-#' conditions, providing a null distribution of the test statistics. For two-group comparisons, the function
-#' computes contrasts between the two groups and applies empirical Bayes moderation. For multi-group analysis
-#' with a single covariate, pairwise contrasts are computed, and the moderated F-statistic is calculated for
-#' each feature.
-#'
-#' @return A list containing the following elements:
-#' \item{d}{A vector of the test statistics (log-fold changes or F-statistics) for each feature.}
-#' \item{s}{A vector of the standard deviations for each feature, adjusted by the empirical Bayes procedure.}
-#'
[…]

### 2024-10-04 R/LM_with_limma_fit.R

@@ -1,16 +1,14 @@
 #' Perform Linear Modeling with Covariates using Limma
 #'
-#' This function performs linear modeling using the Limma package while
-#' accounting for covariates specified
-#' in the `meta.info`. It supports two-group comparisons and multi-group
-#' analysis, incorporating covariates
-#' through a design matrix.
+#' This function performs linear modeling using the Limma package, incorporating
+#'  covariates in the model fitting process. It is designed to handle both
+#'  two-group comparisons and multi-group settings with covariates.
 #'
 #' @param x A list containing two or more data matrices where rows represent
 #' features (e.g., genes, proteins) and columns represent samples. The list
 #' should contain at least two matrices for pairwise group comparison.
 #' @param group.name A character string indicating the name of the group
-#' variable in `meta.info` to be used in the analysis.
+#' variable in`meta.info` to be used in the analysis.
 #' @param meta.info A data frame containing the metadata for the samples.
 #' This includes sample grouping and any covariates to be included in the model.
 #' @param formula.str A string specifying the formula to be used in model
@@ -18,42 +16,45 @@
 #' (e.g., `~ covariate1 + covariate2`).
 #' @param trend A logical value indicating whether to allow for an
 #' intensity-dependent trend in the prior variance.
-#' @param robust A logical value indicating whether to use a robust
-#' fitting procedure to protect against outliers.
+#' @param robust A logical value indicating whether to use a robust fitting
+#' procedure to protect against outliers.
 #'
 #' @details
-#' This function first combines the data matrices from different groups and
-#' prepares a design matrix based on the covariates specified in `meta.info`
-#' using the provided formula. It fits a linear model using Limma,
-#' computes contrasts between groups, and applies empirical Bayes moderation.
-#' For two-group comparisons, the function returns log-fold changes and
-#' associated statistics. In multi-group settings with a single covariate,
-#' it calculates pairwise contrasts and moderated F-statistics.
+#' This function combines the data matrices from different groups and fits a
+#' linear model using covariates provided in the `meta.info`. For two-group
[…]

### 2024-10-04 R/LM_with_limma_permuting.R

@@ -1,0 +1,138 @@
+#' Perform Permutation-Based Linear Modeling with Covariates using Limma
+#'
+#' This function performs linear modeling using the Limma package with
+#' permutation of the covariates to evaluate the test statistics under random
+#' assignments. It handles two-group comparisons and multi-group settings.
+#'
+#' @param x A list containing two or more data matrices where rows represent
+#' features (e.g., genes, proteins) and columns represent samples. The list
+#' should contain at least two matrices for pairwise group comparison.
+#' @param group.name A character string indicating the name of the group
+#' variable in `meta.info` to be used in the analysis.
+#' @param meta.info A data frame containing the metadata for the samples.
+#' This includes sample grouping and any covariates to be included in the model.
+#' @param formula.str A string specifying the formula to be used in model
+#' fitting. It should follow the standard R formula syntax
+#' (e.g., `~ covariate1 + covariate2`).
+#' @param trend A logical value indicating whether to allow for an
+#' intensity-dependent trend in the prior variance.
+#' @param robust A logical value indicating whether to use a robust
+#' fitting procedure to protect against outliers.
+#' @param permutating.group Logical, If \code{TRUE}, the permutation for
+#' calculating the null distribution is performed by permuting the target group
+#' only specified in \code{group.name}. If FALSE, the entire \code{meta.info}
+#' will be permuted (recommended to be set to TRUE).
+#'
+#' @details
+#' This function combines the data matrices from different groups and permutes
+#' the covariates from `meta.info`before fitting a linear model using Limma.
+#' Permutation helps assess how the covariates behave under random conditions,
+#' providing a null distribution of the test statistics. For two-group
+#' comparisons, the function computes contrasts between the two groups and
+#' applies empirical Bayes moderation. For multi-group analysis with a single
+#' covariate, pairwise contrasts are computed, and the moderated F-statistic is
+#' calculated for each feature.
+#'
+#' @return A list containing the following elements:
+#' \item{d}{A vector of the test statistics (log-fold changes or F-statistics)
+#' for each feature.}
+#' \item{s}{A vector of the standard deviations for each feature, adjusted by
+#' the empirical Bayes procedure.}
+#'
+#' @seealso \code{\link[limma]{lmFit}}, \code{\link[limma]{eBayes}},
[…]

### 2024-10-04 R/LimROTS.r

@@ -1,26 +1,59 @@
-#' LimROTS: A Hybrid Method Integrating Empirical Bayes and Reproducibility-Optimized Statistics for Robust Analysis of Proteomics and Metabolomics Data
+#' LimROTS: A Hybrid Method Integrating Empirical Bayes and
+#' Reproducibility-Optimized Statistics for Robust Analysis of Proteomics and
+#' Metabolomics Data
 #'
-#' The `LimROTS` function employs a reproducibility-optimized test statistic utilising the limma and ROTS methodology to simulate complex experimental designs.
+#' The `LimROTS` function employs a reproducibility-optimized test statistic
+#' utilising the limma and ROTS methodology to simulate complex experimental
+#' designs.
 #'
-#' @param x A \code{SummarizedExperiment} object or a matrix where rows represent features (e.g., genes, proteins) and columns represent samples. The values should be log-transformed.
-#' @param B An integer specifying the number of bootstrap iterations. Default is 1000.
-#' @param K An optional integer representing the top list size for ranking. If not specified, it is set to one-fourth of the number of features.
-#' @param a1 Optional numeric value used in the optimization process. If defined by the user, no optimization occurs.
-#' @param a2 Optional numeric value used in the optimization process. If defined by the user, no optimization occurs.
-#' @param log Logical, indicating whether the data is already log-transformed. Default is \code{TRUE}.
-#' @param progress Logical, indicating whether to display a progress bar during bootstrap sampling. Default is \code{FALSE}.
-#' @param verbose Logical, indicating whether to display messages during the function's execution. Default is \code{TRUE}.
-#' @param meta.info A data frame containing sample-level metadata, where each row corresponds to a sample. It should include the grouping variable specified in \code{group.name}. If \code{x} is a \code{SummarizedExperiment} object, \code{meta.info} must be a vector of the metadata needed for the model to run and can be retrieved using \code{colData()}.
-#' @param group.name A string specifying the column in \code{meta.info} that represents the groups or conditions for comparison.
-#' @param seed.cl An integer specifying the seed for randomization; if not provided, the default is 1234.
-#' @param cluster A parallel cluster object for distributed computation, e.g., created by \code{makeCluster()}. Default is 2.
[…]

### 2024-10-04 R/SanityChecK.R

@@ -1,22 +1,40 @@
 #' Sanity Check for Input Data and Parameters
 #'
-#' This function performs a series of checks and initial setups for input data, metadata, and parameters, ensuring everything is correctly formatted for downstream analysis.
+#' This function performs a series of checks and initial setups for input data,
+#' metadata, and parameters, ensuring everything is correctly formatted for
+#' downstream analysis.
 #'
-#' @param x A matrix-like object or a `SummarizedExperiment` containing the data to be analyzed.
-#' @param B Integer. Number of bootstrap samples or resampling iterations. Default is 1000.
-#' @param K Integer. Top list size. If NULL, it will be set to a quarter of the number of rows in the data matrix. Default is NULL.
+#' @param x A matrix-like object or a `SummarizedExperiment` containing the
+#' data to be analyzed.
+#' @param B Integer. Number of bootstrap samples or resampling iterations.
+#' Default is 1000.
+#' @param K Integer. Top list size. If NULL, it will be set to a quarter of
+#' the number of rows in the data matrix. Default is NULL.
 #' @param a1,a2 Optional numeric parameters related to optimization.
-#' @param meta.info Data frame. Metadata associated with the samples (columns of `data.exp`). If `data.exp` is a `SummarizedExperiment`, `meta.info` can be a vector of `colData` column names to use.
-#' @param group.name Character. Column name in `meta.info` that defines the groups or conditions for comparison.
-#' @param formula.str Optional character string representing the formula for the model.
-#' @param survival Logical. If TRUE, survival analysis is used, requiring `time` and `event` columns in `meta.info`. Default is FALSE.
-#' @param paired Logical. If TRUE, indicates paired test setup. Default is FALSE.
-#' @param n.ROTS Logical. If TRUE, uses the ROTS method instead of LimROTS. Default is FALSE.
-#' @param verbose Logical, indicating whether to display messages during the function's execution. Default is TRUE.
-#' @param log Logical, indicating whether the data is already log-transformed. Default is TRUE.
+#' @param meta.info Data frame. Metadata associated with the samples
+#' (columns of `data.exp`). If `data.exp` is a `SummarizedExperiment`,
+#' `meta.info` can be a vector of `colData` column names to use.
+#' @param group.name Character. Column name in `meta.info` that defines the
+#' groups or conditions for comparison.
+#' @param formula.str Optional character string representing the formula for
+#' the model.
+#' @param survival Logical. If TRUE, survival analysis is used, requiring
[…]

### 2024-10-04 R/calOverlaps_slr.R

@@ -1,22 +1,30 @@
 #' Calculate Overlaps for Single-Label Replicates (SLR)
 #'
-#' This function computes the overlap between two sets of observed and permuted values for single-label replicates (SLR).
-#' It calculates the proportion of overlap between pairs of vectors (res1/res2 and pres1/pres2) after sorting them.
+#' This function computes the overlap between two sets of observed and permuted
+#' 'values for single-label replicates (SLR).
+#' It calculates the proportion of overlap between pairs of vectors (res1/res2
+#' and pres1/pres2) after sorting them.
 #'
 #' @param D Numeric vector. Observed data values (e.g., differences).
 #' @param pD Numeric vector. Permuted data values.
 #' @param nrow Integer. Number of rows in each block of data.
-#' @param N Integer vector. Number of top values to consider for overlap calculation.
+#' @param N Integer vector. Number of top values to consider for overlap
+#' calculation.
 #' @param N_len Integer. Length of the `N` vector.
 #' @param B Integer. Number of bootstrap samples or resampling iterations.
-#' @param overlaps Numeric matrix. Matrix to store overlap results for observed data.
-#' @param overlaps_P Numeric matrix. Matrix to store overlap results for permuted data.
+#' @param overlaps Numeric matrix. Matrix to store overlap results for
+#' observed data.
+#' @param overlaps_P Numeric matrix. Matrix to store overlap results for
+#' permuted data.
 #'
 #' @details
-#' The function calculates the overlap for two sets of comparisons: one for observed data (`res1`/`res2`) and one for permuted data (`pres1`/`pres2`).
-#' For each bootstrap sample, the function orders the two vectors being compared, then computes the proportion of overlap for the top `N` values.
+#' The function calculates the overlap for two sets of comparisons: one for
+#'  observed data (`res1`/`res2`) and one for permuted data (`pres1`/`pres2`).
+#' For each bootstrap sample, the function orders the two vectors being
+#' compared, then computes the proportion of overlap for the top `N` values.
 #'
-#' @return A list containing two matrices: \code{overlaps} for observed data and \code{overlaps_P} for permuted data.
+#' @return A list containing two matrices: \code{overlaps} for observed data and
[…]

### 2024-10-04 R/testStatOptimized.R

@@ -1,13 +1,23 @@
 #' Optimized Test Statistic Calculation
 #'
-#' This function calculates the test statistic (mean differences and standard deviations) for comparing two or more groups of samples, with support for paired or unpaired samples.
+#' This function calculates the test statistic (mean differences and standard
+#' deviations) for comparing two or more groups of samples, with support for
+#' paired or unpaired samples.
 #'
-#' @param isPaired Logical. If TRUE, calculates test statistics for paired samples. If FALSE, for unpaired samples.
-#' @param x List of matrices or data frames. Each element represents a group of samples (columns) with the same set of features (rows).
+#' @param isPaired Logical. If TRUE, calculates test statistics for paired
+#' samples. If FALSE, for unpaired samples.
+#' @param x List of matrices or data frames. Each element represents a group of
+#' samples (columns) with the same set of features (rows).
 #'
 #' @details
-#' The function supports comparison between two groups or multiple groups of samples. For two groups, it computes the mean differences and pooled standard deviations for unpaired samples, or paired standard deviations for paired samples.
-#' When comparing more than two groups, it calculates the mean differences and standard deviations across all groups. For unpaired samples, a scaling factor based on sample size is used. Paired comparisons are only supported for two groups.
+#' The function supports comparison between two groups or multiple groups of
+#' samples. For two groups, it computes the mean differences and pooled
+#' standard deviations for unpaired samples, or paired standard deviations for
+#' paired samples.
+#' When comparing more than two groups, it calculates the mean differences and
+#' standard deviations across all groups. For unpaired samples, a scaling factor
+#'  based on sample size is used. Paired comparisons are only supported for two
+#'  groups.
 #'
 #' @return A list containing:
 #' \itemize{
@@ -15,11 +25,6 @@
 #'   \item \code{s}: Standard deviations (pooled or paired).
 #' }
 #'
-#'
-
-
-
-
 testStatOptimized <- function(isPaired, x) {
     sampleGroups <- x
     if (length(sampleGroups) == 2) {
@@ -33,16 +38,24 @@ testStatOptimized <- function(isPaired, x) {
             nonNAcountA <- rowSums(!is.na(groupA))
[…]

### 2024-10-04 R/testStatSurvivalOptimized.R

@@ -1,65 +1,74 @@
 #' @title Test Statistics for Survival Data
-#' @description This function calculates mean differences and standard deviations for survival data
-#' across different sample groups, considering at-risk samples at each unique event time.
+#' @description This function calculates mean differences and standard
+#' deviations for survival data across different sample groups, considering
+#' at-risk samples at each unique event time.
 #'
-#' @param x A list of matrices or data frames, where each element represents a group of samples
-#' (columns) with the same set of features (rows).
-#' @param survivalTime A numeric vector containing the survival times for each sample.
-#' @param survivalEvent A binary vector indicating the occurrence of an event (1 for event, 0 for censored)
-#' for each sample.
+#' @param x A list of matrices or data frames, where each element represents
+#' a group of samples (columns) with the same set of features (rows).
+#' @param survivalTime A numeric vector containing the survival times for
+#' each sample.
+#' @param survivalEvent A binary vector indicating the occurrence of an event
+#' (1 for event, 0 for censored) for each sample.
 #'
 #' @return A list containing:
-#' \item{d}{A numeric vector of mean differences for each feature across the groups.}
+#' \item{d}{A numeric vector of mean differences for each feature across
+#' the groups.}
 #' \item{s}{A numeric vector of standard deviations for the mean differences.}
 #'
-#' @details The function identifies unique times from the survival data and computes the mean difference
-#' for each time point by comparing samples that experienced the event against those at risk.
-#' It calculates the standard deviation based on the variation among at-risk samples.
+#' @details The function identifies unique times from the survival data and
+#' computes the mean difference for each time point by comparing samples that
+#' experienced the event against those at risk. It calculates the standard
+#' deviation based on the variation among at-risk samples.
 #'
 #'
 
 
 
 
-testStatSurvivalOptimized <- function(x,
-                                      survivalTime,
-                                      survivalEvent) {
-    sampleGroups <- x
-    allSamples <- do.call("cbind", sampleGroups)
-    uniqueTimes <- unique(survivalTime[survivalEvent == 1])
+testStatSurvivalOptimized <-
+    function(x, survivalTime, survivalEvent) {
+        sampleGroups <- x
[…]

### 2024-10-05 R/LimROTS.r
correcting the rworkflow error

@@ -106,7 +106,7 @@
 #' \eqn{SEpost = √(s2.post)
 #' \times unscaled SD} for each feature is computed, along with the regression
 #' coefficient for each feature (indicating the impact of variations in the
-#' experimental settings). The \link[ROTS]{ROTS} approach establishes optimality
+#' experimental settings). The \link{ROTS} approach establishes optimality
 #' based on the largest overlap of top-ranked features within group-preserving
 #' bootstrap datasets, Finally based on the optimized parameters
 #' \eqn{\alpha1} and

### 2024-10-08 R/Boot_parallel.R
bioccheck

@@ -1,6 +1,8 @@
 #' Parallel processing handling function
 #'
-#'
+#' @param data A \code{SummarizedExperiment} object or a matrix where rows
+#' represent features (e.g., genes, proteins) and columns represent samples.
+#' The values should be log-transformed.
 #' @param a1 Optional numeric value used in the optimization process.
 #' If defined by the user, no optimization occurs.
 #' @param a2 Optional numeric value used in the optimization process.
@@ -33,6 +35,13 @@
 #' @return A list containing: \code{D, S, pD, pS} for bootstrapped data and
 #'  for permuted data.
 #'
+#' @importFrom parallel makeCluster clusterSetRNGStream clusterExport
+#' stopCluster
+#' @importFrom doParallel registerDoParallel
+#' @importFrom foreach foreach
+#' @import doRNG
+#'
+#'
 #'
 
 Boot_parallel <- function(cluster, seed.cl , samples, data,
@@ -48,7 +57,7 @@ Boot_parallel <- function(cluster, seed.cl , samples, data,
     }
     clusterSetRNGStream(cluster, iseed = seed.cl)
     clusterExport( cluster,
-            varlist = c("samples", "pSamples", "data",
+            varlist = c("samples", "data",
                             "formula.str", "group.name", "groups", "meta.info",
                                                 "a1", "a2", "trend", "robust" ,
                                                 "permutating.group"),

### 2024-10-08 R/Check_meta_info.R

@@ -7,6 +7,7 @@
 #' @param log Logical, indicating whether the data is already log-transformed.
 #' Default is TRUE.
 #'
+#' @return Logical
 #'
 #'
 
@@ -28,4 +29,6 @@ Check_meta_info <- function(meta.info, data, log) {
                 If the data is log2 values then only set log to TRUE."
         )
     }
+
+    return(TRUE)
 }

### 2024-10-08 R/LimROTS.r

@@ -77,17 +77,10 @@
 #'     formula.str = formula.str, B = 10
 #' )
 #'
-#' @importFrom limma voom lmFit eBayes
 #' @importFrom stats model.matrix formula p.adjust
 #' @importFrom dplyr bind_cols
-#' @importFrom parallel makeCluster clusterSetRNGStream clusterExport
-#' stopCluster
-#' @importFrom doParallel registerDoParallel
-#' @importFrom foreach foreach
-#' @import doRNG
 #' @importFrom qvalue empPvals qvalue
 #' @import utils
-#' @import SummarizedExperiment
 #'
 #' @details The **LimROTS** approach initially uses the
 #' \link{limma} package to simulate the intensity data of proteins and
@@ -257,7 +250,7 @@ LimROTS <- function(x,
         R <- optimized.parameters$R
         Z <- optimized.parameters$Z
         ztable <- optimized.parameters$ztable
-        fit <- testStatistic_with_covariates_Fit(
+        fit <- Limma_fit(
             x = lapply(split(seq_len(length(
                 groups
             )), groups), function(x)

### 2024-10-09 R/LimROTS.r
vignettes

@@ -9,7 +9,7 @@
 #' @param x A \code{SummarizedExperiment} object or a matrix where rows
 #' represent features (e.g., genes, proteins) and columns represent samples.
 #' The values should be log-transformed.
-#' @param B An integer specifying the number of bootstrap iterations.
+#' @param B An integer representing the amount of bootstrap iterations.
 #' Default is 1000.
 #' @param K An optional integer representing the top list size for ranking.
 #' If not specified, it is set to one-fourth of the number of features.
@@ -95,9 +95,9 @@
 #' bootstrap datasets, Finally based on the optimized parameters
 #' \eqn{\alpha1} and
 #' \eqn{\alpha2} this equation used to calculates the final statistics:
-#' \deqn{t_{\alpha}(g) = \frac{\beta}{\alpha1 + \alpha2 \times SEpost}}where
-#'          \eqn{t_{\alpha}(g)} is the final statistics for each feature,
-#'          \eqn{\beta} is the coefficient, and SEpost is the the adjusted
+#' \deqn{t_{\alpha_{(p)}} = \frac{\beta_{(p)}}{\alpha1 + \alpha2 \times SEpost_{(p)}}}where
+#'          \eqn{t_{\alpha_{(p)}}} is the final statistics for each feature,
+#'          \eqn{\beta_{(p)}} is the coefficient, and \eqn{SEpost_{(p)}} is the the adjusted
 #'          standard error. LimROTS generates p-values from permutation samples
 #'          using the implementation available in
 #'          \link{qvalue} package, along with internal implementation of FDR

### 2024-10-09 R/LimROTS.r
.

@@ -1,13 +1,9 @@
-#' LimROTS: A Hybrid Method Integrating Empirical Bayes and
+#' `LimROTS`: A Hybrid Method Integrating Empirical Bayes and
 #' Reproducibility-Optimized Statistics for Robust Analysis of Proteomics and
 #' Metabolomics Data
 #'
-#' The `LimROTS` function employs a reproducibility-optimized test statistic
-#' utilising the limma and ROTS methodology to simulate complex experimental
-#' designs.
-#'
 #' @param x A \code{SummarizedExperiment} object or a matrix where rows
-#' represent features (e.g., genes, proteins) and columns represent samples.
+#' represent features (e.g., proteins, metabolites) and columns represent samples.
 #' The values should be log-transformed.
 #' @param B An integer representing the amount of bootstrap iterations.
 #' Default is 1000.
@@ -19,8 +15,8 @@
 #' If defined by the user, no optimization occurs.
 #' @param log Logical, indicating whether the data is already log-transformed.
 #' Default is \code{TRUE}.
-#' @param progress Logical, indicating whether to display a progress bar during
-#' bootstrap sampling. Default is \code{FALSE}.
+#' @param progress Logical, indicating whether to display a progress bar the
+#' function's execution. Default is \code{FALSE}.
 #' @param verbose Logical, indicating whether to display messages during the
 #' function's execution. Default is \code{TRUE}.
 #' @param meta.info A data frame containing sample-level metadata, where each
@@ -34,29 +30,33 @@
 #' if not provided, the default is 1234.
 #' @param cluster A parallel cluster object for distributed computation,
 #' e.g., created by \code{makeCluster()}. Default is 2.
-#' @param formula.str A formula string used when covariates are present in meta.
-#' info for modeling. It should include "~ 0 + ..." to exclude the
-#' intercept from the model.
+#' @param formula.str A formula string for modeling.
+#' It should include "~ 0 + ..." to exclude the intercept from the model.
+#' All the model parameters must be present in \code{meta.info}.
 #' @param robust indicating whether robust fitting should be used.
 #' Default is TRUE, see \link{eBayes}.
 #' @param trend indicating whether to include trend fitting in the
 #' differential expression analysis. Default is TRUE. see \link{eBayes}.
 #' @param permutating.group Logical, If \code{TRUE}, the permutation for
-#' calculating the null distribution is performed by permuting the target
[…]

### 2024-10-09 R/LimROTS.r
.

@@ -3,7 +3,8 @@
 #' Metabolomics Data
 #'
 #' @param x A \code{SummarizedExperiment} object or a matrix where rows
-#' represent features (e.g., proteins, metabolites) and columns represent samples.
+#' represent features (e.g., proteins, metabolites) and columns
+#' represent samples.
 #' The values should be log-transformed.
 #' @param B An integer representing the amount of bootstrap iterations.
 #' Default is 1000.
@@ -83,7 +84,8 @@
 #' @import utils
 #'
 #' @details The **LimROTS** approach initially uses
-#' \link{limma} package functionality to simulate the intensity data of proteins and
+#' \link{limma} package functionality to simulate the intensity data of
+#' proteins and
 #' metabolites. A linear model is subsequently fitted using the design matrix.
 #' Empirical Bayes variance shrinking is then implemented. To obtain the
 #' moderated t-statistics, the adjusted standard error
@@ -96,9 +98,12 @@
 #' bootstrap datasets, Finally based on the optimized parameters
 #' \eqn{\alpha1} and
 #' \eqn{\alpha2} this equation used to calculates the final statistics:
-#' \deqn{t_{\alpha_{(p)}} = \frac{\beta_{(p)}}{\alpha1 + \alpha2 \times SEpost_{(p)}}}where
+#'
+#' \deqn{t_{\alpha_{(p)}} = \frac{\beta_{(p)}}
+#'          {\alpha1 + \alpha2 \times SEpost_{(p)}}}where
 #'          \eqn{t_{\alpha_{(p)}}} is the final statistics for each feature,
-#'          \eqn{\beta_{(p)}} is the coefficient, and \eqn{SEpost_{(p)}} is the the adjusted
+#'          \eqn{\beta_{(p)}} is the coefficient, and \eqn{SEpost_{(p)}}
+#'          is the the adjusted
 #'          standard error. LimROTS generates p-values from permutation samples
 #'          using the implementation available in
 #'          \link{qvalue} package, along with internal implementation of FDR

### 2024-12-05 R/Optimizing.R
SanityChecKandB

@@ -4,7 +4,7 @@
 #' and permuted data for multiple values of a smoothing constant (`ssq`) and a
 #' single-label replicate (SLR) comparison.
 #'
-#' @param B Integer. Number of bootstrap samples or resampling iterations.
+#' @param niter Integer. Number of bootstrap samples or resampling iterations.
 #' @param ssq Numeric vector. Smoothing constants to be evaluated.
 #' @param N Integer vector. Number of top values to consider for overlap
 #' calculation.
@@ -37,8 +37,8 @@
 #'
 #'
 #'
-Optimizing <- function(B, ssq, N, D, S, pD, pS, verbose, progress) {
-    if (verbose) {message("Optimizing parameters")}
+Optimizing <- function(niter, ssq, N, D, S, pD, pS, verbose, progress) {
+    if (verbose) {message("Optimizing a1 and a2")}
     reprotable <- matrix(nrow = length(ssq) + 1, ncol = length(N))
     colnames(reprotable) <- N
     row.names(reprotable) <- c(ssq, "slr")
@@ -49,10 +49,10 @@ Optimizing <- function(B, ssq, N, D, S, pD, pS, verbose, progress) {
     colnames(reprotable.sd) <- N
     row.names(reprotable.sd) <- c(ssq, "slr")
     for (i in seq_len(length(ssq))) {
-        overlaps <- matrix(0, nrow = B, ncol = length(N))
-        overlaps.P <- matrix(0, nrow = B, ncol = length(N))
+        overlaps <- matrix(0, nrow = niter, ncol = length(N))
+        overlaps.P <- matrix(0, nrow = niter, ncol = length(N))
         cResults <- calOverlaps(D, S, pD, pS, nrow(D), as.integer(N), length(N),
-                                ssq[i], as.integer(B), overlaps, overlaps.P)
+                                ssq[i], as.integer(niter), overlaps, overlaps.P)
         reprotable[i, ] <- colMeans(cResults[["overlaps"]])
         reprotable.P[i, ] <- colMeans(cResults[["overlaps_P"]])
         reprotable.sd[i, ] <- sqrt(rowSums((t(cResults[["overlaps"]]) -
@@ -60,10 +60,10 @@ Optimizing <- function(B, ssq, N, D, S, pD, pS, verbose, progress) {
                                             (nrow(cResults[["overlaps"]]) - 1))
     }
     i <- length(ssq) + 1
[…]

### 2024-12-05 R/SanityChecK.R

@@ -6,7 +6,7 @@
 #'
 #' @param x A matrix-like object or a `SummarizedExperiment` containing the
 #' data to be analyzed.
-#' @param B Integer. Number of bootstrap samples or resampling iterations.
+#' @param niter Integer. Number of bootstrap samples or resampling iterations.
 #' Default is 1000.
 #' @param K Integer. Top list size. If NULL, it will be set to a quarter of
 #' the number of rows in the data matrix. Default is NULL.
@@ -39,9 +39,10 @@
 #' }
 #'
 #'
-SanityChecK <- function(x, B = 1000, K = NULL, a1 = NULL, a2 = NULL,
-                        meta.info = NULL, group.name = NULL,
-                        formula.str = NULL, verbose = TRUE, log = TRUE) {
+SanityChecK <- function(x, niter = 1000, K = NULL,
+                        meta.info , group.name,
+                        formula.str, verbose = TRUE, log = TRUE) {
+  
     data.exp <- x
     Check_SExp <- Check_SummarizedExperiment(data.exp = x ,
                                                 meta.info = meta.info,

### 2024-12-05 R/calOverlaps.R

@@ -15,7 +15,7 @@
 #' calculation.
 #' @param N_len Integer. Length of the `N` vector.
 #' @param ssq Numeric. A small constant added to standard errors for stability.
-#' @param B Integer. Number of bootstrap samples or resampling iterations.
+#' @param niter Integer. Number of bootstrap samples or resampling iterations.
 #' @param overlaps Numeric matrix. Matrix to store overlap results for observed
 #' data.
 #' @param overlaps_P Numeric matrix. Matrix to store overlap results for
@@ -32,7 +32,7 @@
 #'
 #'
 
-calOverlaps <- function(D, S, pD, pS, nrow, N, N_len, ssq, B, overlaps,
+calOverlaps <- function(D, S, pD, pS, nrow, N, N_len, ssq, niter, overlaps,
                                 overlaps_P) {
     sort2_1R <- function(a, b) {
         order_a <- order(a, b, decreasing = TRUE)
@@ -40,8 +40,8 @@ calOverlaps <- function(D, S, pD, pS, nrow, N, N_len, ssq, B, overlaps,
         b <- b[order_a]
         list(a = a, b = b)
     }
-    idx_b <- seq_len(B)
-    idx_offset <- B
+    idx_b <- seq_len(niter)
+    idx_offset <- niter
     D <- abs(D)
     for (b in idx_b) {
         res1 <-

### 2024-12-05 R/calOverlaps_slr.R

@@ -11,7 +11,7 @@
 #' @param N Integer vector. Number of top values to consider for overlap
 #' calculation.
 #' @param N_len Integer. Length of the `N` vector.
-#' @param B Integer. Number of bootstrap samples or resampling iterations.
+#' @param niter Integer. Number of bootstrap samples or resampling iterations.
 #' @param overlaps Numeric matrix. Matrix to store overlap results for
 #' observed data.
 #' @param overlaps_P Numeric matrix. Matrix to store overlap results for
@@ -30,15 +30,15 @@
 
 
 
-calOverlaps_slr <- function(D, pD, nrow, N, N_len, B, overlaps, overlaps_P) {
+calOverlaps_slr <- function(D, pD, nrow, N, N_len, niter, overlaps, overlaps_P) {
     sort2_1R <- function(a, b) {
         order_a <- order(a, b, decreasing = TRUE)
         a <- a[order_a]
         b <- b[order_a]
         list(a = a, b = b)
     }
-    idx_b <- seq_len(B)
-    idx_offset <- B
+    idx_b <- seq_len(niter)
+    idx_offset <- niter
     D <- abs(D)
     for (b in idx_b) {
         res1 <- abs(D[((b - 1) * nrow + 1):(b * nrow)])

### 2024-12-05 tests/testthat/testLimROTS.R
SanityChecKandB

@@ -10,12 +10,12 @@ test_that("LiROTS the main function returns expected output structure", {
   formula.str <- "~ 0 + group"
   result <- LimROTS(data,
       meta.info = meta.info, group.name = "group",
-      formula.str = formula.str, B = 10, seed.cl = 1234 )
+      formula.str = formula.str, niter = 10, seed.cl = 1234 )
   
   expect_type(result, "list")
   expect_length(result, 16)
   
-  expected_names <- c("data", "B", "d", "logfc", "pvalue", "FDR",
+  expected_names <- c("data", "niter", "d", "logfc", "pvalue", "FDR",
                       "a1", "a2", "k", "R", "Z", "ztable", "groups",
                       "corrected.logfc" , "q_values", "BH.pvalue")
   

### 2024-12-05 R/Boot_parallel.R
SanityChecKandB

@@ -40,59 +40,59 @@
 #' stopCluster
 #' @importFrom doParallel registerDoParallel
 #' @importFrom foreach foreach
-#' @importFrom doRNG %dorng%
+#' @import doRNG
 #'
 #'
 #'
 
 Boot_parallel <- function(cluster, seed.cl , samples, data,
-                                        formula.str, group.name, groups,
-                                        meta.info, a1, a2, trend, robust,
-                                        permutating.group) {
-    if (is.null(cluster)) {
-        cluster <- makeCluster(2)
-        registerDoParallel(cluster)
-        message("No cluster found; only two cores will be used!")
-    } else {
-        registerDoParallel(cluster)
+                          formula.str, group.name, groups,
+                          meta.info, a1, a2, trend, robust,
+                          permutating.group) {
+  if (is.null(cluster)) {
+    cluster <- makeCluster(2)
+    registerDoParallel(cluster)
+    message("No cluster found; only two cores will be used!")
+  } else {
+    registerDoParallel(cluster)
+  }
+  clusterSetRNGStream(cluster, iseed = seed.cl)
+  clusterExport( cluster,
+                 varlist = c("samples", "data",
+                             "formula.str", "group.name", "groups", "meta.info",
+                             "a1", "a2", "trend", "robust" ,
+                             "permutating.group"),
+                 envir = environment())
+  i <- NULL
+  results_list <- foreach(
+    i = seq_len(nrow(samples)),
+    .combine = "c", .packages = c("utils", "stringr", "stats", "limma"),
+    .export = c("Limma_bootstrap",
+                "Limma_permutating")
+  ) %dorng% {
+    samples.R <- split(samples[i, ], groups)
+    d_result <- s_result <- pd_result <- ps_result <- NULL
+    if (is.null(a1) | is.null(a2)) {
+      fit <- Limma_bootstrap(
+        x = lapply(samples.R, function(x) data[, x]),
+        group.name = group.name, meta.info = meta.info,
+        formula.str = formula.str, trend = trend, robust = robust
+      )
     }
-    clusterSetRNGStream(cluster, iseed = seed.cl)
-    clusterExport( cluster,
-            varlist = c("samples", "data",
-                            "formula.str", "group.name", "groups", "meta.info",
-                                                "a1", "a2", "trend", "robust" ,
-                                                "permutating.group"),
-                        envir = environment())
-    i <- NULL
[…]

### 2024-12-05 R/LimROTS.r

@@ -82,6 +82,7 @@
 #' @importFrom dplyr bind_cols
 #' @importFrom qvalue empPvals qvalue
 #' @importFrom utils txtProgressBar setTxtProgressBar
+#' @importFrom BiocParallel SnowParam MulticoreParam bplapply bpoptions bpRNGseed
 #'
 #' @details The **LimROTS** approach initially uses
 #' \link{limma} package functionality to simulate the intensity data of
@@ -166,7 +167,6 @@ LimROTS <- function(x,
     event <- SanityChecK.list$event
     K <- SanityChecK.list$K
 
-    #### FC
     if (length(unique(groups)) == 2) {
         group1_data <- data[, groups == 1]
         group2_data <- data[, groups == 2]
@@ -201,6 +201,7 @@ LimROTS <- function(x,
     if (progress) {
         setTxtProgressBar(pb, 50)
     }
+    
     results_list <- Boot_parallel(cluster = cluster, seed.cl = seed.cl,
                                         samples = samples, data = data,
                                         formula.str = formula.str,

### 2024-12-06 R/Boot_parallel.R
BiocParallel

@@ -36,63 +36,83 @@
 #' @return A list containing: \code{D, S, pD, pS} for bootstrapped data and
 #'  for permuted data.
 #'
-#' @importFrom parallel makeCluster clusterSetRNGStream clusterExport
-#' stopCluster
-#' @importFrom doParallel registerDoParallel
-#' @importFrom foreach foreach
-#' @import doRNG
-#'
-#'
+#' @importFrom BiocParallel SnowParam MulticoreParam bplapply
+#' @importFrom stats model.matrix formula p.adjust
+#' @importFrom basilisk.utils isWindows
 #'
 
-Boot_parallel <- function(cluster, seed.cl , samples, data,
+
+Boot_parallel <- function(cluster = NULL, seed.cl, samples, data,
                           formula.str, group.name, groups,
                           meta.info, a1, a2, trend, robust,
                           permutating.group) {
+  
   if (is.null(cluster)) {
-    cluster <- makeCluster(2)
-    registerDoParallel(cluster)
-    message("No cluster found; only two cores will be used!")
-  } else {
-    registerDoParallel(cluster)
-  }
-  clusterSetRNGStream(cluster, iseed = seed.cl)
-  clusterExport( cluster,
-                 varlist = c("samples", "data",
-                             "formula.str", "group.name", "groups", "meta.info",
-                             "a1", "a2", "trend", "robust" ,
-                             "permutating.group"),
-                 envir = environment())
-  i <- NULL
-  results_list <- foreach(
-    i = seq_len(nrow(samples)),
-    .combine = "c", .packages = c("utils", "stringr", "stats", "limma"),
-    .export = c("Limma_bootstrap",
-                "Limma_permutating")
-  ) %dorng% {
-    samples.R <- split(samples[i, ], groups)
-    d_result <- s_result <- pd_result <- ps_result <- NULL
-    if (is.null(a1) | is.null(a2)) {
-      fit <- Limma_bootstrap(
-        x = lapply(samples.R, function(x) data[, x]),
-        group.name = group.name, meta.info = meta.info,
-        formula.str = formula.str, trend = trend, robust = robust
-      )
+    if (isWindows()) {
+      cluster <- SnowParam(workers = 2)
+      message("Using SnowParam (Windows) with two workers.")
+    } else {
+      cluster <- MulticoreParam(workers = 2)
+      message("Using MulticoreParam (Unix-like OS) with two workers.")
     }
[…]

### 2024-12-06 R/Limma_bootstrap.r

@@ -44,6 +44,7 @@
 #' @importFrom dplyr bind_cols
 #' @importFrom stringr str_split_fixed fixed
 #' @importFrom limma makeContrasts lmFit contrasts.fit eBayes topTable
+#' @importFrom utils combn
 #'
 
 Limma_bootstrap <-

### 2024-12-06 R/Limma_permutating.R

@@ -47,6 +47,9 @@
 #' @importFrom dplyr bind_cols
 #' @importFrom limma makeContrasts lmFit contrasts.fit eBayes topTable
 #' @importFrom stringr str_split_fixed fixed
+#' @importFrom utils combn
+#' 
+#' 
 #'
 Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
                                     robust, permutating.group) {

### 2024-12-06 R/Optimizing.R

@@ -14,7 +14,6 @@
 #' @param pS Numeric matrix. Standard errors or related values for
 #' permuted data.
 #' @param verbose Logical. If `TRUE`, progress messages will be displayed.
-#' @param progress Logical. If `TRUE`, a progress bar will be shown.
 #'
 #' @details
 #' The function calculates overlaps for a range of smoothing constants and
@@ -37,7 +36,7 @@
 #'
 #'
 #'
-Optimizing <- function(niter, ssq, N, D, S, pD, pS, verbose, progress) {
+Optimizing <- function(niter, ssq, N, D, S, pD, pS, verbose) {
     if (verbose) {message("Optimizing a1 and a2")}
     reprotable <- matrix(nrow = length(ssq) + 1, ncol = length(N))
     colnames(reprotable) <- N

### 2024-12-06 R/SanityChecK.R

@@ -10,7 +10,6 @@
 #' Default is 1000.
 #' @param K Integer. Top list size. If NULL, it will be set to a quarter of
 #' the number of rows in the data matrix. Default is NULL.
-#' @param a1,a2 Optional numeric parameters related to optimization.
 #' @param meta.info Data frame. Metadata associated with the samples
 #' (columns of `data.exp`). If `data.exp` is a `SummarizedExperiment`,
 #' `meta.info` can be a vector of `colData` column names to use.

### 2024-12-06 R/calculateFalseDiscoveryRate.R

@@ -10,8 +10,6 @@
 #' @param permutedValues Numeric matrix. The permuted test statistics or values,
 #'  with rows corresponding to the same values as in `observedValues` and
 #'  columns representing different permutations.
-#' @param showProgress Logical. If `TRUE`, a progress bar will be shown during
-#' the computation.
 #'
 #' @return A numeric vector of the same length as `observedValues`, containing
 #' the estimated FDR for each observed value.
@@ -21,34 +19,23 @@
 #' observedValues <- c(2.5, 1.8, 3.1, 0.7, 2.9)
 #' set.seed(123)
 #' permutedValues <- matrix(rnorm(5 * 5, mean = 2, sd = 1), nrow = 5)
-#' fdr <- calculateFalseDiscoveryRate(observedValues, permutedValues,
-#'     showProgress = FALSE
+#' fdr <- calculateFalseDiscoveryRate(observedValues, permutedValues
 #' )
 #' print(fdr)
 #'
 #' @export
 
-calculateFalseDiscoveryRate <- function(observedValues, permutedValues,
-                                        showProgress = FALSE) {
+calculateFalseDiscoveryRate <- function(observedValues, permutedValues) {
     observedAbs <- abs(observedValues)
     permutedAbs <- abs(permutedValues)
     ord.obs <- order(observedAbs, decreasing = TRUE, na.last = TRUE)
     abs.obs <- observedAbs[ord.obs]
     numPermutations <- ncol(permutedValues)
     FDRmatrix <- matrix(NA, nrow = length(abs.obs), ncol = numPermutations)
-    if (showProgress) {
-        progressBar <- txtProgressBar(min = 0, max = numPermutations, style = 3)
-    }
     for (i in seq_len(numPermutations)) {
         a.rand <- sort(permutedAbs[, i], decreasing = TRUE, na.last = TRUE)
         bigger <- countLargerThan(abs.obs, a.rand)
         FDRmatrix[ord.obs, i] <- bigger / seq_along(abs.obs)
-        if (showProgress) {
-            setTxtProgressBar(progressBar, i)
-        }
-    }
-    if (showProgress) {
-        close(progressBar)
     }
     falseDiscoveryRate <- apply(FDRmatrix, 1, median)
     falseDiscoveryRate[falseDiscoveryRate > 1] <- 1

### 2024-12-06 R/Bootstrap_functions.r
R450

@@ -30,7 +30,6 @@
 #'     niter = 10, meta.info = meta.info, group.name = "group"
 #' )
 #'
-
 bootstrapS <- function(niter, meta.info, group.name) {
     groups <- meta.info[, group.name]
     bootsamples <- matrix(nrow = niter, ncol = length(groups))
@@ -71,8 +70,8 @@ bootstrapS <- function(niter, meta.info, group.name) {
 #' entry is a resampled row name from the metadata, stratified by group and
 #' additional factors.
 #'
-#' 
-#' 
+#'
+#'
 bootstrapSamples_limRots <- function(niter, meta.info, group.name) {
     labels <- as.numeric(meta.info[, group.name])
     samples <- matrix(nrow = niter, ncol = length(labels))

### 2024-12-06 R/LimROTS.r

@@ -24,10 +24,10 @@
 #' represents the groups or conditions for comparison.
 #' @param seed.cl An integer specifying the seed for randomization;
 #' if not provided, the default is 1234.
-#' @param cluster  A \code{BiocParallelParam} object specifying the  
-#' parallelization backend (e.g., \code{MulticoreParam}, \code{SnowParam}).  
-#' The default depends on the operating system: if the user is on Windows,  
-#' \code{SnowParam(workers = 2)} is used; otherwise, 
+#' @param cluster  A \code{BiocParallelParam} object specifying the
+#' parallelization backend (e.g., \code{MulticoreParam}, \code{SnowParam}).
+#' The default depends on the operating system: if the user is on Windows,
+#' \code{SnowParam(workers = 2)} is used; otherwise,
 #' \code{MulticoreParam(workers = 2)}.
 #' @param formula.str A formula string for modeling.
 #' It should include "~ 0 + ..." to exclude the intercept from the model.
@@ -43,7 +43,8 @@
 #' \code{meta.info} will be permuted (recommended to be set to TRUE).
 #'
 #'
-#' @return An object of class `"SummarizedExperiment"` with the following elements:
+#' @return An object of class `"SummarizedExperiment"` with the 
+#' following elements:
 #' \item{data}{The original data matrix.}
 #' \item{niter}{The number of bootstrap samples used.}
 #' \item{d}{The optimized statistics for each feature.}
@@ -80,7 +81,7 @@
 #' @importFrom dplyr bind_cols
 #' @importFrom qvalue empPvals qvalue
 #' @importFrom utils txtProgressBar setTxtProgressBar
-#' @importFrom BiocParallel SnowParam MulticoreParam bplapply bpoptions 
+#' @importFrom BiocParallel SnowParam MulticoreParam bplapply bpoptions
 #' bpRNGseed
 #' @importFrom S4Vectors DataFrame metadata
 #'
@@ -90,7 +91,7 @@
 #' metabolites. A linear model is subsequently fitted using the design matrix.
 #' Empirical Bayes variance shrinking is then implemented. To obtain the
 #' moderated t-statistics, the adjusted standard error
-#' \eqn{SE_{post} = \sqrt{s^2_{\text{post}} } \times unscaled SD} 
+#' \eqn{SE_{post} = \sqrt{s^2_{\text{post}} } \times unscaled SD}
 #' for each feature is computed, along with the regression
 #' coefficient for each feature (indicating the impact of variations in the
[…]

### 2024-12-06 R/Limma_bootstrap.r

@@ -48,72 +48,76 @@
 #'
 
 Limma_bootstrap <-
-    function(x, group.name, meta.info, formula.str, trend,robust) {
+    function(x, group.name, meta.info, formula.str, trend, robust) {
         data <- x
         combined_data <- data.frame(
             check.rows = FALSE,
             check.names = FALSE,
             none = rep("none", nrow(data[[1]]))
-    )
-    for (k.list in names(data)) {
-        combined_data <- cbind(
-            combined_data,
-            data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE)
         )
-    }
-    combined_data <- combined_data[, -1]
-    colnames(combined_data) <-
-        paste0(colnames(combined_data), ".", seq(1, ncol(combined_data)))
-    covariates.p <- data.frame()
-    meta.info.temp <- meta.info
-    meta.info.temp$sample.id <- row.names(meta.info.temp)
-    for (i in colnames(combined_data)) {
-        real_SampleNames <- str_split_fixed(i, fixed("."), 2)[, 1]
-        df.temp <-
-            meta.info.temp[row.names(meta.info.temp) %in% real_SampleNames, ]
-        df.temp$sample.id <- i
-        covariates.p <- rbind(covariates.p, df.temp)
-    }
-    covariates.p$sample.id <- NULL
-    row.names(covariates.p) <- NULL
-    design.matrix <- model.matrix(formula(formula.str), data = covariates.p)
-    colnames(design.matrix) <- make.names(colnames(design.matrix))
-    fit <- lmFit(combined_data, design.matrix)
-    if (length(data) == 2) {
-        pairwise_contrasts <-
-            paste0(group.name, unique(covariates.p[, group.name]))
-        pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x)
-            paste(x[1], "-", x[2]))
-        cont_matrix <- makeContrasts(
-            contrasts = pairwise_contrasts, levels =
-                design.matrix
-        )
-        fit2 <- contrasts.fit(fit, cont_matrix)
-        fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
-        d_values <- topTable(
-            fit.ebayes,
-            coef = pairwise_contrasts,
-            number = "Inf",
-            sort.by = 'none'
-        )
-        d_values <- abs(d_values$logFC)
-        s_values <-
-            as.numeric(sqrt(fit.ebayes$s2.post) *
[…]

### 2024-12-06 R/Limma_fit.R

@@ -46,75 +46,78 @@
 #' @importFrom dplyr bind_cols
 #' @importFrom limma makeContrasts lmFit contrasts.fit eBayes topTable
 #' @importFrom utils combn
-#' 
-#' 
+#'
+#'
 #'
 
-Limma_fit  <-
-    function(x, group.name, meta.info, formula.str, trend, robust){
-    data <- x
-    combined_data <- data.frame(
-        check.rows = FALSE,
-        check.names = FALSE,
-        none = rep("none", nrow(data[[1]]))
-    )
-    for (k.list in names(data)) {
-        combined_data <- cbind(
-            combined_data,
-            data.frame(data[[k.list]], check.rows = FALSE, check.names = FALSE)
+Limma_fit <-
+    function(x, group.name, meta.info, formula.str, trend, robust) {
+        data <- x
+        combined_data <- data.frame(
+            check.rows = FALSE,
+            check.names = FALSE,
+            none = rep("none", nrow(data[[1]]))
         )
-    }
-    combined_data <- combined_data[, -1]
-    design.matrix <- model.matrix(formula(formula.str), data = meta.info)
-    colnames(design.matrix) <- make.names(colnames(design.matrix))
-    fit <- lmFit(combined_data, design.matrix)
-    if (length(data) == 2) {
-        pairwise_contrasts <- paste0(group.name, unique(meta.info[
-            ,
-            group.name
-        ]))
-        pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x)
-            paste(x[1], "-", x[2]))
-        cont_matrix <- makeContrasts(
-            contrasts = pairwise_contrasts, levels =
-                design.matrix
-        )
-        fit2 <- contrasts.fit(fit, cont_matrix)
-        fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
-        d_values <- topTable(
-            fit.ebayes,
-            coef = pairwise_contrasts,
-            number = "Inf",
-            sort.by = 'none'
-        )
-        corrected.logfc <- d_values$logFC
-        d_values <- abs(d_values$logFC)
-        s_values <-
-            as.numeric(sqrt(fit.ebayes$s2.post) *
-                fit.ebayes$stdev.unscaled[, 1])
-        return(list(
-            d = d_values,
-            s = s_values,
-            corrected.logfc = corrected.logfc
-        ))
-    } else if (length(data) > 2 & ncol(meta.info) == 1) {
[…]

### 2024-12-06 R/Limma_permutating.R

@@ -48,19 +48,25 @@
 #' @importFrom limma makeContrasts lmFit contrasts.fit eBayes topTable
 #' @importFrom stringr str_split_fixed fixed
 #' @importFrom utils combn
-#' 
-#' 
+#'
+#'
 #'
 Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
-                                    robust, permutating.group) {
+    robust, permutating.group) {
     data <- x
-    combined_data <- data.frame( check.rows = FALSE, check.names = FALSE,
-                                        none = rep("none", nrow(data[[1]])) )
+    combined_data <- data.frame(
+        check.rows = FALSE, check.names = FALSE,
+        none = rep("none", nrow(data[[1]]))
+    )
     for (k.list in names(data)) {
-        combined_data <- cbind(combined_data,
-                                    data.frame(data[[k.list]],
-                                                    check.rows = FALSE,
-                                                    check.names = FALSE))}
+        combined_data <- cbind(
+            combined_data,
+            data.frame(data[[k.list]],
+                check.rows = FALSE,
+                check.names = FALSE
+            )
+        )
+    }
     combined_data <- combined_data[, -1]
     colnames(combined_data) <-
         paste0(colnames(combined_data), ".", seq(1, ncol(combined_data)))
@@ -70,7 +76,7 @@ Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
     for (i in colnames(combined_data)) {
         real_SampleNames <- str_split_fixed(i, fixed("."), 2)[, 1]
         df.temp <- meta.info.temp[row.names(meta.info.temp) %in%
-                                                    real_SampleNames, ]
+            real_SampleNames, ]
         df.temp$sample.id <- i
         covariates.p <- rbind(covariates.p, df.temp)
     }
@@ -89,25 +95,33 @@ Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
     if (length(data) == 2) {
         pairwise_contrasts <-
             paste0(group.name, unique(covariates.p[, group.name]))
-        pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x)
-                                                    paste(x[1], "-", x[2]))
-        cont_matrix <- makeContrasts(contrasts = pairwise_contrasts,
-                                                    levels = design.matrix)
+        pairwise_contrasts <- combn(pairwise_contrasts, 2, function(x) {
[…]

### 2024-12-06 R/calOverlaps_slr.R

@@ -30,7 +30,8 @@
 
 
 
-calOverlaps_slr <- function(D, pD, nrow, N, N_len, niter, overlaps, overlaps_P) {
+calOverlaps_slr <- function(D, pD, nrow, N, N_len, niter, 
+                                            overlaps, overlaps_P) {
     sort2_1R <- function(a, b) {
         order_a <- order(a, b, decreasing = TRUE)
         a <- a[order_a]

### 2024-12-06 R/calculateFalseDiscoveryRate.R

@@ -19,8 +19,7 @@
 #' observedValues <- c(2.5, 1.8, 3.1, 0.7, 2.9)
 #' set.seed(123)
 #' permutedValues <- matrix(rnorm(5 * 5, mean = 2, sd = 1), nrow = 5)
-#' fdr <- calculateFalseDiscoveryRate(observedValues, permutedValues
-#' )
+#' fdr <- calculateFalseDiscoveryRate(observedValues, permutedValues)
 #' print(fdr)
 #'
 #' @export
@@ -42,8 +41,9 @@ calculateFalseDiscoveryRate <- function(observedValues, permutedValues) {
     falseDiscoveryRate[ord.obs] <-
         rev(vapply(
             length(falseDiscoveryRate):1,
-            function(x)
-                min(falseDiscoveryRate[ord.obs][x:length(falseDiscoveryRate)]),
+            function(x) {
+                min(falseDiscoveryRate[ord.obs][x:length(falseDiscoveryRate)])
+            },
             numeric(1)
         ))
     return(falseDiscoveryRate)

### 2024-12-08 R/Boot_parallel.R
boot_parallel

@@ -14,8 +14,6 @@
 #' model to run and can be retrieved using \code{colData()}.
 #' @param group.name A string specifying the column in \code{meta.info} that
 #' represents the groups or conditions for comparison.
-#' @param seed.cl An integer specifying the seed for randomization;
-#' if not provided, the default is 1234.
 #' @param cluster A parallel cluster object for distributed computation,
 #' e.g., created by \code{makeCluster()}. Default is 2.
 #' @param formula.str A formula string used when covariates are present in meta.
@@ -25,10 +23,6 @@
 #' Default is TRUE, see \link{eBayes}.
 #' @param trend indicating whether to include trend fitting in the
 #' differential expression analysis. Default is TRUE. see \link{eBayes}.
-#' @param permutating.group Logical, If \code{TRUE}, the permutation for
-#' calculating the null distribution is performed by permuting the target
-#' group only specified in \code{group.name}. If FALSE, the entire
-#' \code{meta.info} will be permuted (recommended to be set to FALSE).
 #' @param samples bootstrapped samples matrix
 #' @param groups groups information from `meta.info`
 #'
@@ -43,7 +37,6 @@
 
 
 Boot_parallel <- function(cluster = NULL,
-    seed.cl,
     samples,
     data,
     formula.str,
@@ -54,13 +47,13 @@ Boot_parallel <- function(cluster = NULL,
     a2,
     trend,
     robust,
-    permutating.group) {
+    pSamples) {
     if (is.null(cluster)) {
         if (isWindows()) {
-            cluster <- SnowParam(workers = 2, RNGseed = seed.cl)
+            cluster <- SnowParam(workers = 2)
             message("Using SnowParam (Windows) with two workers.")
         } else {
-            cluster <- MulticoreParam(workers = 2, RNGseed = seed.cl)
+            cluster <- MulticoreParam(workers = 2)
             message("Using MulticoreParam (Unix-like OS) with two workers.")
         }
     } else {
@@ -68,9 +61,6 @@ Boot_parallel <- function(cluster = NULL,
     }
     if (inherits(cluster, "SnowParam")) {
         cluster$exportglobals <- FALSE
-    }
-    if(is.null(cluster$RNGseed)){
-        stop("RNGseed must be set to the Parallel backend")
     }
     export_vars <- list(
         samples = samples,
@@ -83,7 +73,7 @@ Boot_parallel <- function(cluster = NULL,
         a2 = a2,
         trend = trend,
         robust = robust,
[…]

### 2024-12-08 R/LimROTS.r

@@ -190,18 +190,45 @@ LimROTS <- function(x,
         samples <- bootstrapSamples_limRots(
             niter = 2 * niter,
             meta.info = meta.info,
-            group.name = group.name
+            group.name = group.name,
+            seed.cl = seed.cl
         )
-        pSamples <- NULL
+        if(permutating.group == TRUE){
+            pSamples <- list()
+            set.seed(seed = seed.cl) 
+            for (i in seq_len(nrow(samples)) ) {
+                shuffle_df <- meta.info
+                shuffle_df[, group.name] <- sample(shuffle_df[, group.name])
+                colnames(shuffle_df) <- colnames(meta.info)
+                pSamples[[i]] <-  shuffle_df
+            }
+        }else{
+            pSamples <- list()
+            set.seed(seed = seed.cl) 
+            for (i in seq_len(nrow(samples)) ) {
+                shuffle_df <- data.frame(meta.info[sample(nrow(meta.info)), ])
+                colnames(shuffle_df) <- colnames(meta.info)
+                pSamples[[i]] <-  shuffle_df
+            }   
+        }
     } else {
-        samples <- bootstrapS(2 * niter, meta.info, group.name)
+        samples <- bootstrapS(2 * niter, meta.info, group.name, 
+                              seed.cl = seed.cl)
+        pSamples <- list()
+        set.seed(seed = seed.cl) 
+        for (i in seq_len(nrow(samples)) ) {
+            shuffle_df <- data.frame(meta.info[sample(nrow(meta.info)), ])
+            colnames(shuffle_df) <- colnames(meta.info)
+            pSamples[[i]] <-  shuffle_df
+        }
+        
     }
     D <- matrix(nrow = nrow(as.matrix(data)), ncol = nrow(samples))
     S <- matrix(nrow = nrow(as.matrix(data)), ncol = nrow(samples))
     pD <- matrix(nrow = nrow(as.matrix(data)), ncol = nrow(samples))
     pS <- matrix(nrow = nrow(as.matrix(data)), ncol = nrow(samples))
     results_list <- Boot_parallel(
-        cluster = cluster, seed.cl = seed.cl,
+        cluster = cluster,
         samples = samples, data = data,
         formula.str = formula.str,
         group.name = group.name,
@@ -209,7 +236,7 @@ LimROTS <- function(x,
         meta.info = meta.info,
         a1 = a1, a2 = a2,
         trend = trend, robust = robust,
[…]

### 2024-12-08 R/Limma_permutating.R

@@ -4,7 +4,7 @@
 #' permutation of the covariates to evaluate the test statistics under random
 #' assignments. It handles two-group comparisons and multi-group settings.
 #'
-#' @param x A list containing two or more data matrices where rows represent
+#' @param x A data matrices where rows represent
 #' features (e.g., genes, proteins) and columns represent samples. The list
 #' should contain at least two matrices for pairwise group comparison.
 #' @param group.name A character string indicating the name of the group
@@ -18,10 +18,6 @@
 #' intensity-dependent trend in the prior variance.
 #' @param robust A logical value indicating whether to use a robust
 #' fitting procedure to protect against outliers.
-#' @param permutating.group Logical, If \code{TRUE}, the permutation for
-#' calculating the null distribution is performed by permuting the target group
-#' only specified in \code{group.name}. If FALSE, the entire \code{meta.info}
-#' will be permuted (recommended to be set to TRUE).
 #'
 #' @details
 #' This function combines the data matrices from different groups and permutes
@@ -53,38 +49,8 @@
 #'
 Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
     robust, permutating.group) {
-    data <- x
-    combined_data <- data.frame(
-        check.rows = FALSE, check.names = FALSE,
-        none = rep("none", nrow(data[[1]]))
-    )
-    for (k.list in names(data)) {
-        combined_data <- cbind(
-            combined_data,
-            data.frame(data[[k.list]],
-                check.rows = FALSE,
-                check.names = FALSE
-            )
-        )
-    }
-    combined_data <- combined_data[, -1]
-    colnames(combined_data) <-
-        paste0(colnames(combined_data), ".", seq(1, ncol(combined_data)))
-    covariates.p <- data.frame()
-    meta.info.temp <- meta.info
-    meta.info.temp$sample.id <- row.names(meta.info.temp)
-    for (i in colnames(combined_data)) {
-        real_SampleNames <- str_split_fixed(i, fixed("."), 2)[, 1]
-        df.temp <- meta.info.temp[row.names(meta.info.temp) %in%
-            real_SampleNames, ]
-        df.temp$sample.id <- i
-        covariates.p <- rbind(covariates.p, df.temp)
-    }
-    if (permutating.group == TRUE) {
[…]

### 2024-12-08 tests/testthat/testBootstrap.R

@@ -13,7 +13,7 @@ test_that("bootstrapS returns expected output structure", {
     group.name <- "group"
 
     # Call the function
-    result <- bootstrapS(niter, meta.info, group.name)
+    result <- bootstrapS(niter, meta.info, group.name, seed.cl = 123)
 
     # Check output structure
     expect_type(result, "character")

### 2024-12-08 tests/testthat/testPermutatingLim.R

@@ -4,23 +4,11 @@ library(stringr)
 
 test_that("Limma_permutating returns expected output structure", {
     # Sample data with 10 samples per group
-    data <- list(
-        group1 = t(matrix(rnorm(400), nrow = 10)), # 10 samples, 4 features
-        group2 = t(matrix(rnorm(400), nrow = 10)) # 10 samples, 4 features
-    )
-
-    data[[1]] <- data.frame(data[[1]])
-    colnames(data[[1]]) <- paste0("sample", 1:10)
-    data[[2]] <- data.frame(data[[2]])
-    colnames(data[[2]]) <- paste0("sample", 11:20)
-
-    # Sample meta info with corresponding group labels
+    data <- data.frame(matrix(rnorm(500), nrow = 100, ncol = 10))
     meta.info <- data.frame(
-        row.names = c(paste0("sample", 1:10), paste0("sample", 11:20)),
-        group = as.factor(c(rep(1, 10), rep(2, 10)))
+        group = factor(rep(1:2, each = 5)),
+        row.names = colnames(data)
     )
-
-    # Formula string for modeling
     formula.str <- "~ 0 + group"
 
     # Call the function

### 2024-12-08 R/LimROTS.r
Bioc

@@ -38,7 +38,7 @@
 #' calculating the null distribution is performed by permuting the target group
 #' only specified in \code{group.name} Preserving all the other sample
 #' information. If `FALSE`, the entire sample information retrieved from
-#' \code{meta.info} will be permuted (recommended to be set to TRUE).
+#' \code{meta.info} will be permuted (recommended to be set to FALSE).
 #'
 #'
 #' @return An object of class `"SummarizedExperiment"` with the 
@@ -167,7 +167,6 @@ LimROTS <- function(x,
     groups <- SanityChecK.list$groups
     event <- SanityChecK.list$event
     K <- SanityChecK.list$K
-
     if (length(unique(groups)) == 2) {
         group1_data <- data[, groups == 1]
         group2_data <- data[, groups == 2]
@@ -200,6 +199,7 @@ LimROTS <- function(x,
         }else{
             pSamples <- list()
             for (i in seq_len(nrow(samples)) ) {
+                shuffle_df <- meta.info
                 shuffle_df <- data.frame(meta.info[sample(nrow(meta.info)), ])
                 colnames(shuffle_df) <- colnames(meta.info)
                 pSamples[[i]] <-  shuffle_df
@@ -209,6 +209,7 @@ LimROTS <- function(x,
         samples <- bootstrapS(2 * niter, meta.info, group.name)
         pSamples <- list()
         for (i in seq_len(nrow(samples)) ) {
+            shuffle_df <- meta.info
             shuffle_df <- data.frame(meta.info[sample(nrow(meta.info)), ])
             colnames(shuffle_df) <- colnames(meta.info)
             pSamples[[i]] <-  shuffle_df
@@ -227,7 +228,6 @@ LimROTS <- function(x,
         groups = groups,
         meta.info = meta.info,
         a1 = a1, a2 = a2,
-        trend = trend, robust = robust,
         pSamples = pSamples
     )
 
@@ -276,7 +276,7 @@ LimROTS <- function(x,
         if (verbose) {
             message("Computing p-values and FDR")
         }
-        p <- empPvals(
+        p <- qvalue::empPvals(
             stat = d,
             stat0 = pD,
             pool = TRUE
@@ -287,7 +287,7 @@ LimROTS <- function(x,
         gc()
         q_values <- tryCatch(
             {
-                qvalue(
+              qvalue::qvalue(
                     p,
                     pi0.method = "bootstrap",
                     lambda = seq(0.01, 0.95, 0.01)

### 2024-12-08 R/Limma_bootstrap.r

@@ -16,10 +16,6 @@
 #' @param formula.str A string specifying the formula to be used in model
 #' fitting. It should follow the standard R formula syntax
 #' (e.g., `~ covariate1 + covariate2`).
-#' @param trend A logical value indicating whether to allow for an
-#' intensity-dependent trend in the prior variance.
-#' @param robust A logical value indicating whether to use a robust
-#' fitting procedure to protect against outliers.
 #'
 #' @details
 #' This function first combines the data matrices from different groups and
@@ -48,7 +44,7 @@
 #'
 
 Limma_bootstrap <-
-    function(x, group.name, meta.info, formula.str, trend, robust) {
+    function(x, group.name, meta.info, formula.str) {
         data <- x
         combined_data <- data.frame(
             check.rows = FALSE,
@@ -92,7 +88,7 @@ Limma_bootstrap <-
                     design.matrix
             )
             fit2 <- contrasts.fit(fit, cont_matrix)
-            fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
+            fit.ebayes <- eBayes(fit2, trend = FALSE, robust = FALSE)
             d_values <- topTable(
                 fit.ebayes,
                 coef = pairwise_contrasts,
@@ -116,7 +112,7 @@ Limma_bootstrap <-
                     levels = design.matrix
                 )
             fit2 <- contrasts.fit(fit, cont_matrix)
-            fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
+            fit.ebayes <- eBayes(fit2, trend = FALSE, robust = FALSE)
             msr <- fit.ebayes$F * fit.ebayes$s2.post
             return(list(d = msr, s = fit.ebayes$s2.post))
         }

### 2024-12-08 R/Limma_permutating.R

@@ -14,10 +14,6 @@
 #' @param formula.str A string specifying the formula to be used in model
 #' fitting. It should follow the standard R formula syntax
 #' (e.g., `~ covariate1 + covariate2`).
-#' @param trend A logical value indicating whether to allow for an
-#' intensity-dependent trend in the prior variance.
-#' @param robust A logical value indicating whether to use a robust
-#' fitting procedure to protect against outliers.
 #'
 #' @details
 #' This function combines the data matrices from different groups and permutes
@@ -47,8 +43,8 @@
 #'
 #'
 #'
-Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
-    robust) {
+Limma_permutating <- function(x, group.name, meta.info, formula.str
+                                        ) {
     combined_data <- x
     covariates.p <- meta.info
     covariates.p$sample.id <- NULL
@@ -69,7 +65,7 @@ Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
             levels = design.matrix
         )
         fit2 <- contrasts.fit(fit, cont_matrix)
-        fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
+        fit.ebayes <- eBayes(fit2, trend = FALSE, robust = FALSE)
         d_values <- topTable(fit.ebayes,
             coef = pairwise_contrasts,
             number = "Inf", sort.by = "none"
@@ -90,7 +86,7 @@ Limma_permutating <- function(x, group.name, meta.info, formula.str, trend,
             levels = design.matrix
         )
         fit2 <- contrasts.fit(fit, cont_matrix)
-        fit.ebayes <- eBayes(fit2, trend = trend, robust = robust)
+        fit.ebayes <- eBayes(fit2, trend = FALSE, robust = FALSE)
         msr <- fit.ebayes$F * fit.ebayes$s2.post
         return(list(d = msr, s = fit.ebayes$s2.post))
     }

### 2024-12-08 tests/testthat/testPermutatingLim.R

@@ -12,8 +12,7 @@ test_that("Limma_permutating returns expected output structure", {
     formula.str <- "~ 0 + group"
 
     # Call the function
-    result <- Limma_permutating(data, "group", meta.info, formula.str,
-                                trend = FALSE, robust = FALSE)
+    result <- Limma_permutating(data, "group", meta.info, formula.str)
 
     # Check output structure
     expect_type(result, "list")

### 2025-05-14 R/LimROTS.r
Update documentation

@@ -21,7 +21,8 @@
 #' @param meta.info a character vector of the metadata needed for the
 #' model to run and can be retrieved using \code{colData()}.
 #' @param group.name A string specifying the column in \code{meta.info} that
-#' represents the groups or conditions for comparison.
+#' represents the groups or conditions for comparison. group.name should be 
+#' retrieved using \code{colData()} as factor.
 #' @param BPPARAM   A \code{BiocParallelParam} object specifying the
 #' parallelization backend (e.g., \code{MulticoreParam}, \code{SnowParam}).
 #' The default depends on the operating system: if the user is on Windows,
@@ -31,9 +32,9 @@
 #' It should include "~ 0 + ..." to exclude the intercept from the model.
 #' All the model parameters must be present in \code{meta.info}.
 #' @param robust indicating whether robust fitting should be used.
-#' Default is TRUE, see \link{eBayes}.
+#' Default is TRUE, see \link[limma]{eBayes}.
 #' @param trend indicating whether to include trend fitting in the
-#' differential expression analysis. Default is TRUE. see \link{eBayes}.
+#' differential expression analysis. Default is TRUE. see \link[limma]{eBayes}.
 #' @param permutating.group Logical, If \code{TRUE}, the permutation for
 #' calculating the null distribution is performed by permuting the target group
 #' only specified in \code{group.name} Preserving all the other sample
@@ -54,7 +55,7 @@
 #' \item{k}{Top list size used for ranking.}
 #' \item{corrected.logfc}{estimate of the log2-fold-change
 #' corresponding to the effect corrected by the s model
-#' see \link{topTable}.}
+#' see \link[limma]{topTable}.}
 #' \item{q_values}{Estimated q-values using the `qvalue` package.}
 #' \item{BH.pvalue}{Benjamini-Hochberg adjusted p-values.}
 #' \item{null.statistics}{The optimized null statistics for each feature.}
@@ -85,7 +86,7 @@
 #' @importFrom S4Vectors DataFrame metadata
 #'
 #' @details The **LimROTS** approach initially uses
-#' \link{limma} package functionality to simulate the intensity data of
+#' \pkg{limma} package functionality to simulate the intensity data of
 #' proteins and
 #' metabolites. A linear model is subsequently fitted using the design matrix.
 #' Empirical Bayes variance shrinking is then implemented. To obtain the
@@ -107,10 +108,10 @@
 #'          is the the adjusted
 #'          standard error. LimROTS generates p-values from permutation samples
[…]

### 2025-06-16 R/SanityChecK.R
define the condition column as a factor

@@ -67,13 +67,18 @@ SanityChecK <- function(x, niter = 1000, K = NULL,
     meta.info$temp <- NULL
     if (inherits(meta.info[, group.name], "character")) {
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), collapse = " & "))
         meta.info[, group.name] <- as.numeric(meta.info[, group.name])
         groups <- as.numeric(meta.info[, group.name])
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
     } else if (inherits(meta.info[, group.name], "factor")) {
         groups <- as.numeric(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), collapse = " & "))
+        meta.info[, group.name] <- as.numeric(meta.info[, group.name])
+        meta.info[, group.name] <- as.factor(meta.info[, group.name])
     } else {
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), collapse = " & "))
         meta.info[, group.name] <- as.numeric(meta.info[, group.name])
         groups <- as.numeric(meta.info[, group.name])
         meta.info[, group.name] <- as.factor(meta.info[, group.name])

### 2025-06-16 R/SanityChecK.R
define the condition column as a factor

@@ -67,13 +67,21 @@ SanityChecK <- function(x, niter = 1000, K = NULL,
     meta.info$temp <- NULL
     if (inherits(meta.info[, group.name], "character")) {
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), 
+                      collapse = " & "))
         meta.info[, group.name] <- as.numeric(meta.info[, group.name])
         groups <- as.numeric(meta.info[, group.name])
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
     } else if (inherits(meta.info[, group.name], "factor")) {
         groups <- as.numeric(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), 
+                      collapse = " & "))
+        meta.info[, group.name] <- as.numeric(meta.info[, group.name])
+        meta.info[, group.name] <- as.factor(meta.info[, group.name])
     } else {
         meta.info[, group.name] <- as.factor(meta.info[, group.name])
+        message(paste("Group Level: ", levels(meta.info[, group.name]), 
+                      collapse = " & "))
         meta.info[, group.name] <- as.numeric(meta.info[, group.name])
         groups <- as.numeric(meta.info[, group.name])
         meta.info[, group.name] <- as.factor(meta.info[, group.name])

### 2025-11-16 R/LimROTS.r
Adding citation information

@@ -108,11 +108,10 @@
 #'          is the the adjusted
 #'          standard error. LimROTS generates p-values from permutation samples
 #'          using the implementation available in
-#'          \link[qvalue]{qvalue} package, along with internal implementation of FDR
-#'          adapted from ROTS package. Additionally, the qvalue package is used
-#'          to calculate q-values, were the proportion of true null p-values is
-#'          set to the bootstrap method \link[qvalue]{pi0est}. We recommend using
-#'          permutation-derived p-values and qvalues.
+#'          \link[qvalue]{qvalue} package, along with internal implementation 
+#'          of FDR adapted from ROTS package. Additionally, the qvalue package 
+#'          is used to calculate q-values, were the proportion of true null 
+#'          p-values is set to the bootstrap method \link[qvalue]{pi0est}. 
 #'
 #' This function processes a dataset using parallel computation. It
 #' leverages the \pkg{BiocParallel} framework to distribute tasks
